### 2018-01-15 paper.tex
Added tex file as is when downloaded

%%%%%%%%%%%%%%%%%%%%%%%%%%  ltexpprt.tex  %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is ltexpprt.tex, an example file for use with the SIAM LaTeX2E
% Preprint Series macros. It is designed to provide double-column output.
% Please take the time to read the following comments, as they document
% how to use these macros. This file can be composed and printed out for
% use as sample output.

% Any comments or questions regarding these macros should be directed to:
%
%                 Donna Witzleben
%                 SIAM
%                 3600 University City Science Center
%                 Philadelphia, PA 19104-2688
%                 USA
%                 Telephone: (215) 382-9800
%                 Fax: (215) 386-7999
%                 e-mail: witzleben@siam.org

% This file is to be used as an example for style only. It should not be read
% for content.

%%%%%%%%%%%%%%% PLEASE NOTE THE FOLLOWING STYLE RESTRICTIONS %%%%%%%%%%%%%%%

%%  1. There are no new tags.  Existing LaTeX tags have been formatted to match
%%     the Preprint series style.
%%
%%  2. You must use \cite in the text to mark your reference citations and
%%     \bibitem in the listing of references at the end of your chapter. See
%%     the examples in the following file. If you are using BibTeX, please
%%     supply the bst file with the manuscript file.
%%
%%  3. This macro is set up for two levels of headings (\section and
%%     \subsection). The macro will automatically number the headings for you.
%%
%%  5. No running heads are to be used for this volume.
%%
%%  6. Theorems, Lemmas, Definitions, etc. are to be double numbered,
%%     indicating the section and the occurence of that element
%%     within that section. (For example, the first theorem in the second
%%     section would be numbered 2.1. The macro will
%%     automatically do the numbering for you.
%%
%%  7. Figures, equations, and tables must be single-numbered.
%%     Use existing LaTeX tags for these elements.
%%     Numbering will be done automatically.
%%
%%  8. Page numbering is no longer included in this macro.
%%     Pagination will be set by the program committee.
%%
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[twoside,leqno,twocolumn]{article}
\usepackage{ltexpprt}

\begin{document}

%\setcounter{chapter}{2} % If you are doing your chapter as chapter one,
%\setcounter{section}{3} % comment these two lines out.

\title{\Large SIAM/ACM Preprint Series Macros for
Use With LaTeX\thanks{Supported by GSF grants ABC123, DEF456, and GHI789.}}
\author{Corey Gray\thanks{Society for Industrial and Applied Mathematics.} \\
\and
Tricia Manning\thanks{Society for Industrial and Applied Mathematics.}}
\date{}

\maketitle

% Copyright Statement
% When submitting your final paper to a SIAM proceedings, it is requested that you include
% the appropriate copyright in the footer of the paper.  The copyright added should be
% consistent with the copyright selected on the copyright form submitted with the paper.
% Please note that "20XX" should be changed to the year of the meeting.

% Default Copyright Statement
\fancyfoot[R]{\footnotesize{\textbf{Copyright \textcopyright\ 20XX by SIAM\\
Unauthorized reproduction of this article is prohibited}}}

% Depending on which copyright you agree to when you sign the copyright form, the copyright
% can be changed to one of the following after commenting out the default copyright statement
% above.

%\fancyfoot[R]{\footnotesize{\textbf{Copyright \textcopyright\ 20XX\\
%Copyright for this paper is retained by authors}}}

%\fancyfoot[R]{\footnotesize{\textbf{Copyright \textcopyright\ 20XX\\
%Copyright retained by principal author's organization}}}

%\pagenumbering{arabic}
%\setcounter{page}{1}%Leave this line commented out.

\begin{abstract} \small\baselineskip=9pt This is the text of my abstract. It is a brief
description of my
paper, outlining the purposes and goals I am trying to address.\end{abstract}

\section{Problem Specification.}In this paper, we consider the solution of the $N \times
N$ linear
system
\begin{equation} \label{e1.1}
A x = b
\end{equation}
where $A$ is large, sparse, symmetric, and positive definite.  We consider
the direct solution of (\ref{e1.1}) by means of general sparse Gaussian
elimination.  In such a procedure, we find a permutation matrix $P$, and
compute the decomposition
\[
P A P^{t} = L D L^{t}
\]
where $L$ is unit lower triangular and $D$ is diagonal.

\section{Design Considerations.}Several good ordering algorithms (nested dissection and
minimum degree)
are available for computing $P$  \cite{GEORGELIU}, \cite{ROSE72}.
Since our interest here does not
focus directly on the ordering, we assume for convenience that $P=I$,
or that $A$ has been preordered to reflect an appropriate choice of $P$.

Our purpose here is to examine the nonnumerical complexity of the
sparse elimination algorithm given in  \cite{BANKSMITH}.
As was shown there, a general sparse elimination scheme based on the
bordering algorithm requires less storage for pointers and
row/column indices than more traditional implementations of general
sparse elimination.  This is accomplished by exploiting the m-tree,
a particular spanning tree for the graph of the filled-in matrix.

\begin{theorem} The method  was extended to three
dimensions. For the standard multigrid
coarsening
(in which, for a given grid, the next coarser grid has $1/8$
as many points), anisotropic problems require plane
relaxation to
obtain a good smoothing factor.\end{theorem}

Our purpose here is to examine the nonnumerical complexity of the
sparse elimination algorithm given in  \cite{BANKSMITH}.
As was shown there, a general sparse elimination scheme based on the
bordering algorithm requires less storage for pointers and
row/column indices than more traditional implementations of general
sparse elimination.  This is accomplished by exploiting the m-tree,
a particular spanning tree for the graph of the filled-in matrix.
Several good ordering algorithms (nested dissection and minimum degree)
are available for computing $P$  \cite{GEORGELIU}, \cite{ROSE72}.
Since our interest here does not
focus directly on the ordering, we assume for convenience that $P=I$,
or that $A$ has been preordered to reflect an appropriate choice of $P$.

\begin{proof} In this paper we consider two methods. The first method
is
basically the method considered with two differences:
first, we perform plane relaxation by a two-dimensional
multigrid method, and second, we use a slightly different
choice of
interpolation operator, which improves performance
for nearly singular problems. In the second method coarsening
is done by successively coarsening in each of the three
independent variables and then ignoring the intermediate
grids; this artifice simplifies coding considerably.
\end{proof}

Our purpose here is to examine the nonnumerical complexity of the
sparse elimination algorithm given in  \cite{BANKSMITH}.
As was shown there, a general sparse elimination scheme based on the
bordering algorithm requires less storage for pointers and
row/column indices than more traditional implementations of general
sparse elimination.  This is accomplished by exploiting the m-tree,
a particular spanning tree for the graph of the filled-in matrix.

\begin{Definition}{\rm We describe the two methods in \S 1.2. In \S\ 1.3. we
discuss
some remaining details.}
\end{Definition}

Our purpose here is to examine the nonnumerical complexity of the
sparse elimination algorithm given in  \cite{BANKSMITH}.
As was shown there, a general sparse elimination scheme based on the
bordering algorithm requires less storage for pointers and
row/column indices than more traditional implementations of general
sparse elimination.  This is accomplished by exploiting the m-tree,
a particular spanning tree for the graph of the filled-in matrix.
Several good ordering algorithms (nested dissection and minimum degree)
are available for computing $P$  \cite{GEORGELIU}, \cite{ROSE72}.
Since our interest here does not
focus directly on the ordering, we assume for convenience that $P=I$,
or that $A$ has been preordered to reflect an appropriate choice of $P$.

Our purpose here is to examine the nonnumerical complexity of the
sparse elimination algorithm given in  \cite{BANKSMITH}.
As was shown there, a general sparse elimination scheme based on the
bordering algorithm requires less storage for pointers and
row/column indices than more traditional implementations of general
sparse elimination.

\begin{lemma} We discuss first the choice for $I_{k-1}^k$
which is a generalization. We assume that $G^{k-1}$ is
obtained
from $G^k$
by standard coarsening; that is, if $G^k$ is a tensor product
grid $G_{x}^k \times G_{y}^k \times G_{z}^k$,
$G^{k-1}=G_{x}^{k-1} \times G_{y}^{k-1} \times G_{z}^{k-1}$,
where $G_{x}^{k-1}$ is obtained by deleting every other grid
point of $G_x^k$ and similarly for $G_{y}^k$ and $G_{z}^k$.
\end{lemma}

To our knowledge, the m-tree previously has not been applied in this
fashion to the numerical factorization, but it has been used,
directly or indirectly, in several optimal order algorithms for
computing the fill-in during the symbolic factorization phase
[4] - [10], [5], [6]. In \S 1.3., we analyze the complexity of the old and new
approaches to the intersection problem for the special case of
an $n \times n$ grid ordered by nested dissection. The special
structure of this problem allows us to make exact estimates of
the complexity. To our knowledge, the m-tree previously has not been applied in this
fashion to the numerical factorization, but it has been used,
directly or indirectly, in several optimal order algorithms for
computing the fill-in during the symbolic factorization phase
[4] - [10], [5], [6].

In \S 1.2, we review the bordering algorithm, and introduce
the sorting and intersection problems that arise in the
sparse formulation of the algorithm.
In \S 1.3., we analyze the complexity of the old and new
approaches to the intersection problem for the special case of
an $n \times n$ grid ordered by nested dissection. The special
structure of this problem allows us to make exact estimates of
the complexity. To our knowledge, the m-tree previously has not been applied in this
fashion to the numerical factorization, but it has been used,
directly or indirectly, in several optimal order algorithms for
computing the fill-in during the symbolic factorization phase
[4] - [10], [5], [6].

For the old approach, we show that the
complexity of the intersection problem is $O(n^{3})$, the same
as the complexity of the numerical computations.  For the
new approach, the complexity of the second part is reduced to
$O(n^{2} (\log n)^{2})$.

To our knowledge, the m-tree previously has not been applied in this
fashion to the numerical factorization, but it has been used,
directly or indirectly, in several optimal order algorithms for
computing the fill-in during the symbolic factorization phase
[4] - [10], [5], [6]. In \S 1.3., we analyze the complexity of the old and new
approaches to the intersection problem for the special case of
an $n \times n$ grid ordered by nested dissection. The special
structure of this problem allows us to make exact estimates of
the complexity. To our knowledge, the m-tree previously has not been applied in this
fashion to the numerical factorization, but it has been used,
directly or indirectly, in several optimal order algorithms for
computing the fill-in during the symbolic factorization phase
[4] - [10], [5], [6].
This is accomplished by exploiting the m-tree,
a particular spanning tree for the graph of the filled-in matrix.
To our knowledge, the m-tree previously has not been applied in this
fashion to the numerical factorization, but it has been used,
directly or indirectly, in several optimal order algorithms for
computing the fill-in during the symbolic factorization phase
\cite{EISENSTAT} - \cite{LIU2}, \cite{ROSE76},  \cite{SCHREIBER}.

\subsection{Robustness.}We do not
attempt to present an overview
here, but rather attempt to focus on those results that
are relevant to our particular algorithm.
This section assumes prior knowledge of the role of graph theory
in sparse Gaussian elimination; surveys of this role are
available in \cite{ROSE72} and \cite{GEORGELIU}. More general
discussions of elimination trees are given in
\cite{LAW} - \cite{LIU2}, \cite{SCHREIBER}.
Thus, at the $k$th stage, the bordering algorithm consists of
solving the lower triangular system
\begin{equation} \label{1.2}
 L_{k-1}v = c
\end{equation}
and setting
\begin{eqnarray}
\ell &=& D^{-1}_{k-1}v , \\
\delta &=& \alpha - \ell^{t} v .
\end{eqnarray}

\begin{figure}
\vspace{14pc}
\caption{This is a figure 1.1.}
\end{figure}

\section{Robustness.} We do not
attempt to present an overview
here, but rather attempt to focus on those results that
are relevant to our particular algorithm.

\subsection{Versatility.}The special
structure of this problem allows us to make exact estimates of
the complexity.  For the old approach, we show that the
complexity of the intersection problem is $O(n^{3})$, the same
as the complexity of the numerical computations
\cite{GEORGELIU}, \cite{ROSEWHITTEN}.  For the
new approach, the complexity of the second part is reduced to
$O(n^{2} (\log n)^{2})$.

To our knowledge, the m-tree previously has not been applied in this
fashion to the numerical factorization, but it has been used,
directly or indirectly, in several optimal order algorithms for
computing the fill-in during the symbolic factorization phase
[4] - [10], [5], [6]. In \S 1.3., we analyze the complexity of the old and new
approaches to the intersection problem for the special case of
an $n \times n$ grid ordered by nested dissection. The special
structure of this problem allows us to make exact estimates of
the complexity. To our knowledge, the m-tree previously has not been applied in this
fashion to the numerical factorization, but it has been used,
directly or indirectly, in several optimal order algorithms for
computing the fill-in during the symbolic factorization phase
[4] - [10], [5], [6].

In \S 1.2, we review the bordering algorithm, and introduce
the sorting and intersection problems that arise in the
sparse formulation of the algorithm.
In \S 1.3., we analyze the complexity of the old and new
approaches to the intersection problem for the special case of
an $n \times n$ grid ordered by nested dissection. The special
structure of this problem allows us to make exact estimates of
the complexity. To our knowledge, the m-tree previously has not been applied in this
fashion to the numerical factorization, but it has been used,
directly or indirectly, in several optimal order algorithms for
computing the fill-in during the symbolic factorization phase
[4] - [10], [5], [6].

For the old approach, we show that the
complexity of the intersection problem is $O(n^{3})$, the same
as the complexity of the numerical computations.  For the
new approach, the complexity of the second part is reduced to
$O(n^{2} (\log n)^{2})$.

To our knowledge, the m-tree previously has not been applied in this
fashion to the numerical factorization, but it has been used,
directly or indirectly, in several optimal order algorithms for
computing the fill-in during the symbolic factorization phase
[4] - [10], [5], [6]. In \S 1.3., we analyze the complexity of the old and new
approaches to the intersection problem for the special case of
an $n \times n$ grid ordered by nested dissection. The special
structure of this problem allows us to make exact estimates of
the complexity. To our knowledge, the m-tree previously has not been applied in this
fashion to the numerical factorization, but it has been used,
directly or indirectly, in several optimal order algorithms for
computing the fill-in during the symbolic factorization phase
[4] - [10], [5], [6].
This is accomplished by exploiting the m-tree,
a particular spanning tree for the graph of the filled-in matrix.
To our knowledge, the m-tree previously has not been applied in this
fashion to the numerical factorization, but it has been used,
directly or indirectly, in several optimal order algorithms for
computing the fill-in during the symbolic factorization phase
\cite{EISENSTAT} - \cite{LIU2}, \cite{ROSE76},  \cite{SCHREIBER}.

\begin{thebibliography}{99}

%\bibitem{GUIDE}
%R.~E. Bank, {\em PLTMG  users' guide, edition 5.0}, tech. report,
%  Department of Mathematics, University of California, San Diego, CA, 1988.

%\bibitem{HBMG}
%R.~E. Bank, T.~F. Dupont, and H.~Yserentant, {\em The hierarchical basis
%  multigrid method}, Numer. Math., 52 (1988), pp.~427--458.

\bibitem{BANKSMITH}
R.~E. Bank and R.~K. Smith, {\em General sparse elimination requires no
  permanent integer storage}, SIAM J. Sci. Stat. Comput., 8 (1987),
  pp.~574--584.

\bibitem{EISENSTAT}
S.~C. Eisenstat, M.~C. Gursky, M.~Schultz, and A.~Sherman, {\em
  Algorithms and data structures for sparse symmetric gaussian elimination},
  SIAM J. Sci. Stat. Comput., 2 (1982), pp.~225--237.

\bibitem{GEORGELIU}
A.~George and J.~Liu, {\em Computer Solution of Large Sparse Positive
  Definite Systems}, Prentice Hall, Englewood Cliffs, NJ, 1981.

\bibitem{LAW}
K.~H. Law and S.~J. Fenves, {\em A node addition model for symbolic
  factorization}, ACM TOMS, 12 (1986), pp.~37--50.

\bibitem{LIU}
J.~W.~H. Liu, {\em A compact row storage scheme for cholesky factors
  using elimination trees}, ACM TOMS, 12 (1986), pp.~127--148.

\bibitem{LIU2}
\sameauthor , {\em The role of
  elimination trees in sparse factorization}, Tech. Report CS-87-12,Department
  of Computer Science, York University, Ontario, Canada, 1987.

\bibitem{ROSE72}
D.~J. Rose, {\em A graph theoretic study of the numeric solution of
  sparse positive definite systems}, in Graph Theory and Computing, Academic  Press, New
York, 1972.

\bibitem{ROSE76}
D.~J. Rose, R.~E. Tarjan, and G.~S. Lueker, {\em Algorithmic aspects of
  vertex elimination on graphs}, SIAM J. Comput., 5 (1976), pp.~226--283.

\bibitem{ROSEWHITTEN}
D.~J. Rose and G.~F. Whitten, {\em A recursive analysis of disection
  strategies}, in Sparse Matrix Computations, Academic Press, New York, 1976.

\bibitem{SCHREIBER}
R.~Schrieber, {\em A new implementation of sparse gaussian elimination},
  ACM TOMS, 8 (1982), pp.~256--276.

\end{thebibliography}
\end{document}

% End of ltexpprt.tex

### 2018-01-25 paper.tex
Restructured into several files. Default text and bibliography entries removed.

@@
 
-\title{\Large SIAM/ACM Preprint Series Macros for
-Use With LaTeX\thanks{Supported by GSF grants ABC123, DEF456, and GHI789.}}
-\author{Corey Gray\thanks{Society for Industrial and Applied Mathematics.} \\
+\title{\Large Vehicle Classification From Automatic Number \\ Plate Recognition Camera Scans}
+\author{Pedro M. Pinto Silva \thanks{School of Computing Science, Newcastle University, United Kingdom}
 \and
-Tricia Manning\thanks{Society for Industrial and Applied Mathematics.}}
+Matthew Forshaw\footnotemark[1]
+\and
+Stephen McGough\footnotemark[1]}
 \date{}
@@
 % Default Copyright Statement
-\fancyfoot[R]{\footnotesize{\textbf{Copyright \textcopyright\ 20XX by SIAM\\
+\fancyfoot[R]{\footnotesize{\textbf{Copyright \textcopyright\ 2018 by BigTraffic\\
 Unauthorized reproduction of this article is prohibited}}}
@@
 
-\section{Problem Specification.}In this paper, we consider the solution of the $N \times
-N$ linear
-system
-\begin{equation} \label{e1.1}
-A x = b
-\end{equation}
-where $A$ is large, sparse, symmetric, and positive definite.  We consider
-the direct solution of (\ref{e1.1}) by means of general sparse Gaussian
-elimination.  In such a procedure, we find a permutation matrix $P$, and
-compute the decomposition
-\[
-P A P^{t} = L D L^{t}
-\]
-where $L$ is unit lower triangular and $D$ is diagonal.
+\input{intro.tex}
 
-\section{Design Considerations.}Several good ordering algorithms (nested dissection and
-minimum degree)
-are available for computing $P$  \cite{GEORGELIU}, \cite{ROSE72}.
-Since our interest here does not
-focus directly on the ordering, we assume for convenience that $P=I$,
-or that $A$ has been preordered to reflect an appropriate choice of $P$.
+\input{trips.tex}
 
-Our purpose here is to examine the nonnumerical complexity of the
-sparse elimination algorithm given in  \cite{BANKSMITH}.
-As was shown there, a general sparse elimination scheme based on the
-bordering algorithm requires less storage for pointers and
-row/column indices than more traditional implementations of general
-sparse elimination.  This is accomplished by exploiting the m-tree,
-a particular spanning tree for the graph of the filled-in matrix.
+\input{classification.tex}
 
-\begin{theorem} The method  was extended to three
-dimensions. For the standard multigrid
-coarsening
-(in which, for a given grid, the next coarser grid has $1/8$
-as many points), anisotropic problems require plane
-relaxation to
-obtain a good smoothing factor.\end{theorem}
-
-Our purpose here is to examine the nonnumerical complexity of the
-sparse elimination algorithm given in  \cite{BANKSMITH}.
-As was shown there, a general sparse elimination scheme based on the
-bordering algorithm requires less storage for pointers and
-row/column indices than more traditional implementations of general
-sparse elimination.  This is accomplished by exploiting the m-tree,
-a particular spanning tree for the graph of the filled-in matrix.
-Several good ordering algorithms (nested dissection and minimum degree)
-are available for computing $P$  \cite{GEORGELIU}, \cite{ROSE72}.
-Since our interest here does not
-focus directly on the ordering, we assume for convenience that $P=I$,
-or that $A$ has been preordered to reflect an appropriate choice of $P$.
-
-\begin{proof} In this paper we consider two methods. The first method
-is
-basically the method considered with two differences:
-first, we perform plane relaxation by a two-dimensional
-multigrid method, and second, we use a slightly different
-choice of
-interpolation operator, which improves performance
-for nearly singular problems. In the second method coarsening
-is done by successively coarsening in each of the three
-independent variables and then ignoring the intermediate
-grids; this artifice simplifies coding considerably.
-\end{proof}
-
-Our purpose here is to examine the nonnumerical complexity of the
-sparse elimination algorithm given in  \cite{BANKSMITH}.
-As was shown there, a general sparse elimination scheme based on the
-bordering algorithm requires less storage for pointers and
-row/column indices than more traditional implementations of general
-sparse elimination.  This is accomplished by exploiting the m-tree,
-a particular spanning tree for the graph of the filled-in matrix.
-
-\begin{Definition}{\rm We describe the two methods in \S 1.2. In \S\ 1.3. we
-discuss
-some remaining details.}
-\end{Definition}
-
-Our purpose here is to examine the nonnumerical complexity of the
-sparse elimination algorithm given in  \cite{BANKSMITH}.
-As was shown there, a general sparse elimination scheme based on the
-bordering algorithm requires less storage for pointers and
-row/column indices than more traditional implementations of general
-sparse elimination.  This is accomplished by exploiting the m-tree,
-a particular spanning tree for the graph of the filled-in matrix.
-Several good ordering algorithms (nested dissection and minimum degree)
-are available for computing $P$  \cite{GEORGELIU}, \cite{ROSE72}.
-Since our interest here does not
-focus directly on the ordering, we assume for convenience that $P=I$,
-or that $A$ has been preordered to reflect an appropriate choice of $P$.
-
-Our purpose here is to examine the nonnumerical complexity of the
-sparse elimination algorithm given in  \cite{BANKSMITH}.
-As was shown there, a general sparse elimination scheme based on the
-bordering algorithm requires less storage for pointers and
-row/column indices than more traditional implementations of general
-sparse elimination.
-
-\begin{lemma} We discuss first the choice for $I_{k-1}^k$
-which is a generalization. We assume that $G^{k-1}$ is
-obtained
-from $G^k$
-by standard coarsening; that is, if $G^k$ is a tensor product
-grid $G_{x}^k \times G_{y}^k \times G_{z}^k$,
-$G^{k-1}=G_{x}^{k-1} \times G_{y}^{k-1} \times G_{z}^{k-1}$,
-where $G_{x}^{k-1}$ is obtained by deleting every other grid
-point of $G_x^k$ and similarly for $G_{y}^k$ and $G_{z}^k$.
-\end{lemma}
-
-To our knowledge, the m-tree previously has not been applied in this
-fashion to the numerical factorization, but it has been used,
-directly or indirectly, in several optimal order algorithms for
-computing the fill-in during the symbolic factorization phase
-[4] - [10], [5], [6]. In \S 1.3., we analyze the complexity of the old and new
-approaches to the intersection problem for the special case of
-an $n \times n$ grid ordered by nested dissection. The special
-structure of this problem allows us to make exact estimates of
-the complexity. To our knowledge, the m-tree previously has not been applied in this
-fashion to the numerical factorization, but it has been used,
-directly or indirectly, in several optimal order algorithms for
-computing the fill-in during the symbolic factorization phase
-[4] - [10], [5], [6].
-
-In \S 1.2, we review the bordering algorithm, and introduce
-the sorting and intersection problems that arise in the
-sparse formulation of the algorithm.
-In \S 1.3., we analyze the complexity of the old and new
-approaches to the intersection problem for the special case of
-an $n \times n$ grid ordered by nested dissection. The special
-structure of this problem allows us to make exact estimates of
-the complexity. To our knowledge, the m-tree previously has not been applied in this
-fashion to the numerical factorization, but it has been used,
-directly or indirectly, in several optimal order algorithms for
-computing the fill-in during the symbolic factorization phase
-[4] - [10], [5], [6].
-
-For the old approach, we show that the
-complexity of the intersection problem is $O(n^{3})$, the same
-as the complexity of the numerical computations.  For the
-new approach, the complexity of the second part is reduced to
-$O(n^{2} (\log n)^{2})$.
-
-To our knowledge, the m-tree previously has not been applied in this
-fashion to the numerical factorization, but it has been used,
-directly or indirectly, in several optimal order algorithms for
-computing the fill-in during the symbolic factorization phase
-[4] - [10], [5], [6]. In \S 1.3., we analyze the complexity of the old and new
-approaches to the intersection problem for the special case of
-an $n \times n$ grid ordered by nested dissection. The special
-structure of this problem allows us to make exact estimates of
-the complexity. To our knowledge, the m-tree previously has not been applied in this
-fashion to the numerical factorization, but it has been used,
-directly or indirectly, in several optimal order algorithms for
-computing the fill-in during the symbolic factorization phase
-[4] - [10], [5], [6].
-This is accomplished by exploiting the m-tree,
-a particular spanning tree for the graph of the filled-in matrix.
-To our knowledge, the m-tree previously has not been applied in this
-fashion to the numerical factorization, but it has been used,
-directly or indirectly, in several optimal order algorithms for
-computing the fill-in during the symbolic factorization phase
-\cite{EISENSTAT} - \cite{LIU2}, \cite{ROSE76},  \cite{SCHREIBER}.
-
-\subsection{Robustness.}We do not
-attempt to present an overview
-here, but rather attempt to focus on those results that
-are relevant to our particular algorithm.
-This section assumes prior knowledge of the role of graph theory
-in sparse Gaussian elimination; surveys of this role are
-available in \cite{ROSE72} and \cite{GEORGELIU}. More general
-discussions of elimination trees are given in
-\cite{LAW} - \cite{LIU2}, \cite{SCHREIBER}.
-Thus, at the $k$th stage, the bordering algorithm consists of
-solving the lower triangular system
-\begin{equation} \label{1.2}
- L_{k-1}v = c
-\end{equation}
-and setting
-\begin{eqnarray}
-\ell &=& D^{-1}_{k-1}v , \\
-\delta &=& \alpha - \ell^{t} v .
-\end{eqnarray}
-
-\begin{figure}
-\vspace{14pc}
-\caption{This is a figure 1.1.}
-\end{figure}
-
-\section{Robustness.} We do not
-attempt to present an overview
-here, but rather attempt to focus on those results that
-are relevant to our particular algorithm.
-
-\subsection{Versatility.}The special
-structure of this problem allows us to make exact estimates of
-the complexity.  For the old approach, we show that the
-complexity of the intersection problem is $O(n^{3})$, the same
-as the complexity of the numerical computations
-\cite{GEORGELIU}, \cite{ROSEWHITTEN}.  For the
-new approach, the complexity of the second part is reduced to
-$O(n^{2} (\log n)^{2})$.
-
-To our knowledge, the m-tree previously has not been applied in this
-fashion to the numerical factorization, but it has been used,
-directly or indirectly, in several optimal order algorithms for
-computing the fill-in during the symbolic factorization phase
-[4] - [10], [5], [6]. In \S 1.3., we analyze the complexity of the old and new
-approaches to the intersection problem for the special case of
-an $n \times n$ grid ordered by nested dissection. The special
-structure of this problem allows us to make exact estimates of
-the complexity. To our knowledge, the m-tree previously has not been applied in this
-fashion to the numerical factorization, but it has been used,
-directly or indirectly, in several optimal order algorithms for
-computing the fill-in during the symbolic factorization phase
-[4] - [10], [5], [6].
-
-In \S 1.2, we review the bordering algorithm, and introduce
-the sorting and intersection problems that arise in the
-sparse formulation of the algorithm.
-In \S 1.3., we analyze the complexity of the old and new
-approaches to the intersection problem for the special case of
-an $n \times n$ grid ordered by nested dissection. The special
-structure of this problem allows us to make exact estimates of
-the complexity. To our knowledge, the m-tree previously has not been applied in this
-fashion to the numerical factorization, but it has been used,
-directly or indirectly, in several optimal order algorithms for
-computing the fill-in during the symbolic factorization phase
-[4] - [10], [5], [6].
-
-For the old approach, we show that the
-complexity of the intersection problem is $O(n^{3})$, the same
-as the complexity of the numerical computations.  For the
-new approach, the complexity of the second part is reduced to
-$O(n^{2} (\log n)^{2})$.
-
-To our knowledge, the m-tree previously has not been applied in this
-fashion to the numerical factorization, but it has been used,
-directly or indirectly, in several optimal order algorithms for
-computing the fill-in during the symbolic factorization phase
-[4] - [10], [5], [6]. In \S 1.3., we analyze the complexity of the old and new
-approaches to the intersection problem for the special case of
-an $n \times n$ grid ordered by nested dissection. The special
-structure of this problem allows us to make exact estimates of
-the complexity. To our knowledge, the m-tree previously has not been applied in this
-fashion to the numerical factorization, but it has been used,
-directly or indirectly, in several optimal order algorithms for
-computing the fill-in during the symbolic factorization phase
-[4] - [10], [5], [6].
-This is accomplished by exploiting the m-tree,
-a particular spanning tree for the graph of the filled-in matrix.
-To our knowledge, the m-tree previously has not been applied in this
-fashion to the numerical factorization, but it has been used,
-directly or indirectly, in several optimal order algorithms for
-computing the fill-in during the symbolic factorization phase
-\cite{EISENSTAT} - \cite{LIU2}, \cite{ROSE76},  \cite{SCHREIBER}.
+\input{discussion.tex}
 
@@
 
-%\bibitem{GUIDE}
-%R.~E. Bank, {\em PLTMG  users' guide, edition 5.0}, tech. report,
-%  Department of Mathematics, University of California, San Diego, CA, 1988.
+\bibitem{ReviewTransportation2010}
+Chen, B.
+and Cheng, H.H., 2010.
+{\em A review of the applications of agent technology in traffic and transportation systems}.
+IEEE Transactions on Intelligent Transportation Systems, 11(2), pp.485-497.
 
-%\bibitem{HBMG}
-%R.~E. Bank, T.~F. Dupont, and H.~Yserentant, {\em The hierarchical basis
-%  multigrid method}, Numer. Math., 52 (1988), pp.~427--458.
+\bibitem{EvolutionUTMC2013}
+Hamilton, A., Waterson, B., Cherrett, T., Robinson, A.
+and Snell, I., 2013.
+{\em The evolution of urban traffic control: changing policy and technology}.
+Transportation planning and technology, 36(1), pp.24-43.
 
-\bibitem{BANKSMITH}
-R.~E. Bank and R.~K. Smith, {\em General sparse elimination requires no
-  permanent integer storage}, SIAM J. Sci. Stat. Comput., 8 (1987),
-  pp.~574--584.
-
-\bibitem{EISENSTAT}
-S.~C. Eisenstat, M.~C. Gursky, M.~Schultz, and A.~Sherman, {\em
-  Algorithms and data structures for sparse symmetric gaussian elimination},
-  SIAM J. Sci. Stat. Comput., 2 (1982), pp.~225--237.
-
-\bibitem{GEORGELIU}
-A.~George and J.~Liu, {\em Computer Solution of Large Sparse Positive
-  Definite Systems}, Prentice Hall, Englewood Cliffs, NJ, 1981.
-
-\bibitem{LAW}
-K.~H. Law and S.~J. Fenves, {\em A node addition model for symbolic
-  factorization}, ACM TOMS, 12 (1986), pp.~37--50.
-
-\bibitem{LIU}
-J.~W.~H. Liu, {\em A compact row storage scheme for cholesky factors
-  using elimination trees}, ACM TOMS, 12 (1986), pp.~127--148.
-
-\bibitem{LIU2}
-\sameauthor , {\em The role of
-  elimination trees in sparse factorization}, Tech. Report CS-87-12,Department
-  of Computer Science, York University, Ontario, Canada, 1987.
-
-\bibitem{ROSE72}
-D.~J. Rose, {\em A graph theoretic study of the numeric solution of
-  sparse positive definite systems}, in Graph Theory and Computing, Academic  Press, New
-York, 1972.
-
-\bibitem{ROSE76}
-D.~J. Rose, R.~E. Tarjan, and G.~S. Lueker, {\em Algorithmic aspects of
-  vertex elimination on graphs}, SIAM J. Comput., 5 (1976), pp.~226--283.
-
-\bibitem{ROSEWHITTEN}
-D.~J. Rose and G.~F. Whitten, {\em A recursive analysis of disection
-  strategies}, in Sparse Matrix Computations, Academic Press, New York, 1976.
-
-\bibitem{SCHREIBER}
-R.~Schrieber, {\em A new implementation of sparse gaussian elimination},
-  ACM TOMS, 8 (1982), pp.~256--276.
+% \bibitem{BANKSMITH}
+% R.~E. Bank and R.~K. Smith, {\em General sparse elimination requires no
+%   permanent integer storage}, SIAM J. Sci. Stat. Comput., 8 (1987),
+%   pp.~574--584.
+%
+% \bibitem{EISENSTAT}
+% S.~C. Eisenstat, M.~C. Gursky, M.~Schultz, and A.~Sherman, {\em
+%   Algorithms and data structures for sparse symmetric gaussian elimination},
+%   SIAM J. Sci. Stat. Comput., 2 (1982), pp.~225--237.
 

### 2018-01-25 intro.tex
First introduction draft

\section{Introduction}

% Automatic number plate recognition (ANPR) cameras are used thou component of current intelligent transportation systems.

Automatic number plate recognition (ANPR) cameras actively employed in cities and urban traffic environments as part of recent advancements in intelligent transportation systems. They can be used in traffic control by Urban Traffic Control and Management (UTMC) entities; in electronic toll collection by comissioned highway agencies; or by law enforcement agencies in the detection of speeding vehicles and validation of number plate registrations. The wide diversity of applications paired with improved performance of ANPR hardware and software systems, has caused an increase in the number of ANPR cameras installed across the UK~\cite{EvolutionUTMC2013}.

There are 267 ANPR cameras in the area of Tyne and Wear, United Kingdom, . Scans made by these cameras are stored in a central database by the UTMC of Tyne and Wear and then used to compute travel times across particular links of interest in the road network. These can be major roads that see high volumes of traffic, or road segments that can be more prone to traffic jams. Average journey times can be then conveyed back to the drivers by the way of Variable Message Signs (VMS) or through . Even though the cameras only cover a very small subset of the entire road network, it provides valuable.

However, Furthermore, ANPR cameras can either be fixed or

Uses of trip data for understandint travel demands / traffic prediction

Studies about best placement of cameras

Number plate data is not as popular, for urban traffic monitoring and prediction , as other sources of data, namely GPS traces or floating vehicle data.

but, to the best of our knowledge, little research has gone into doing so with number plate data. Therefore, the contributions of this paper are twofold:

\begin{itemize}
  \item
\end{itemize}

% In this paper, we consider the solution of the $N \times
% N$ linear
% system
% \begin{equation} \label{e1.1}
% A x = b
% \end{equation}
% where $A$ is large, sparse, symmetric, and positive definite.  We consider
% the direct solution of (\ref{e1.1}) by means of general sparse Gaussian
% elimination.  In such a procedure, we find a permutation matrix $P$, and
% compute the decomposition
% \[
% P A P^{t} = L D L^{t}
% \]
% where $L$ is unit lower triangular and $D$ is diagonal.

### 2018-01-25 classification.tex
Added further sections

\section{Vehicle Classification Using Summarised Trip Information}

### 2018-01-25 trips.tex
Added further sections

\section{Identifying Trips In Number Plate Data}

\subsection{Definitions}

The \emph{ith} sighting of vehicle \emph{k} can be defined as:

\begin{equation} \label{e2.1.1}
s^{k}_{i} = \left(camera, time\right)
\end{equation}

Trip
\begin{equation} \label{e2.1.2}
trip^{k}_{u} = \left(s^{k}_{(1)}, s^{k}_{(2)}, \dots , s^{k}_{(n)}\right)
\end{equation}

where \( n \) is the number of sightings, i.e.\ length of the trip.

Trips by the same vehicle can be distinguished

From the \( uth \) trip of vehicle \( k \), we can calculate the corresponding journey time sequence, of length \(n-1\):

\begin{equation} \label{e2.1.3}
journey\ time^{k}_{u} = \left(time^{k}_{2} - time^{k}_{1}, \ldots, time^{k}_{n} - time^{k}_{n-1} \right)
\end{equation}

A trip of length \( \ge 2 \) is valid if all partial journey times are lower than a threshold \( \tau \):

\begin{equation} \label{e2.1.4}
journey\ time^{k}_{u(i)} < \tau \ , \ \forall i \in journey\ time^{k}_{u}
\end{equation}

Realistically there should also be a lower bound for each journey time.

Clock Synchronisation

\subsection{Considerations}

% \begin{Definition}{\rm We describe the two methods in \S 1.2. In \S\ 1.3. we
% discuss
% some remaining details.}
% \end{Definition}
%
% Our purpose here is to examine the nonnumerical complexity of the
% sparse elimination algorithm given in  \cite{BANKSMITH}.
% As was shown there, a general sparse elimination scheme based on the
% bordering algorithm requires less storage for pointers and
% row/column indices than more traditional implementations of general
% sparse elimination.  This is accomplished by exploiting the m-tree,
% a particular spanning tree for the graph of the filled-in matrix.
% Several good ordering algorithms (nested dissection and minimum degree)
% are available for computing $P$  \cite{GEORGELIU}, \cite{ROSE72}.
% Since our interest here does not
% focus directly on the ordering, we assume for convenience that $P=I$,
% or that $A$ has been preordered to reflect an appropriate choice of $P$.
%
% \begin{equation} \label{1.2}
%  L_{k-1}v = c
% \end{equation}
% and setting
% \begin{eqnarray}
% \ell &=& D^{-1}_{k-1}v , \\
% \delta &=& \alpha - \ell^{t} v .
% \end{eqnarray}
%
% \begin{figure}
% \vspace{14pc}
% \caption{This is a figure 1.1.}
% \end{figure}

### 2018-01-25 paper.tex
Added some references

@@
 \usepackage{ltexpprt}
+\usepackage{parskip}
 
@@
 
-\title{\Large Vehicle Classification From Automatic Number \\ Plate Recognition Camera Scans}
+\title{\Large Classifying Vehicles From Automatic Number \\ Plate Recognition Camera Scans}
 \author{Pedro M. Pinto Silva \thanks{School of Computing Science, Newcastle University, United Kingdom}
@@
 
-\bibitem{ReviewTransportation2010}
-Chen, B.
-and Cheng, H.H., 2010.
-{\em A review of the applications of agent technology in traffic and transportation systems}.
-IEEE Transactions on Intelligent Transportation Systems, 11(2), pp.485-497.
+\bibitem{SurveyITS2011}
+Zhang, Junping, et al. {\em Data-driven intelligent transportation systems: A survey}. IEEE Transactions on Intelligent Transportation Systems 12.4 (2011): 1624-1639.
 
 \bibitem{EvolutionUTMC2013}
-Hamilton, A., Waterson, B., Cherrett, T., Robinson, A.
-and Snell, I., 2013.
-{\em The evolution of urban traffic control: changing policy and technology}.
-Transportation planning and technology, 36(1), pp.24-43.
+Hamilton, Andrew, et al. {\em The evolution of urban traffic control: changing policy and technology}. Transportation planning and technology 36.1 (2013): 24-43.
+
+\bibitem{Castillo2008}
+Castillo, Enrique, José María Menéndez, and Pilar Jiménez.
+{\em Trip matrix and path flow reconstruction and estimation based on plate scanning and link observations}.
+Transportation Research Part B: Methodological 42.5 (2008): 455-481.
+
+\bibitem{Hazelton2012}
+Parry, Katharina, and Martin L.
+Hazelton.
+{\em Estimation of origin-–destination matrices from link counts and sporadic routing data}.
+Transportation Research Part B: Methodological 46.1 (2012): 175-188.
 

### 2018-01-25 intro.tex
Added some more text to introduction

@@
 
-% Automatic number plate recognition (ANPR) cameras are used thou component of current intelligent transportation systems.
+Automatic number plate recognition (ANPR) cameras are actively employed in urban traffic environments and play an important role in day-to-day intelligent transportation systems.
+They can be used by Urban Traffic Control and Management (UTMC) entities; by comissioned highway agencies in electronic toll collection; or by law enforcement organisations in the detection of speeding vehicles and validation of number plate registrations.
+The wide diversity of applications, paired with the large improvement in price-to-performance ratio of ANPR hardware and software systems, has resulted in the increase of installed ANPR cameras across the UK~\cite{SurveyITS2011, EvolutionUTMC2013}.
 
-Automatic number plate recognition (ANPR) cameras actively employed in cities and urban traffic environments as part of recent advancements in intelligent transportation systems.
-They can be used in traffic control by Urban Traffic Control and Management (UTMC) entities; in electronic toll collection by comissioned highway agencies; or by law enforcement agencies in the detection of speeding vehicles and validation of number plate registrations.
-The wide diversity of applications paired with improved performance of ANPR hardware and software systems, has caused an increase in the number of ANPR cameras installed across the UK~\cite{EvolutionUTMC2013}.
+In the area of Tyne and Wear, United Kingdom, there are over 250 active ANPR cameras.
+Scans made by these cameras are stored in a central database by the UTMC of Tyne and Wear and then used to compute travel times across particular links of interest in the road network.
+These are usually major roads that see high volumes of traffic, or road segments more prone to traffic jams.
+Average journey times can then be conveyed back to the drivers by the way of Variable Message Signs (VMS) or web based applications.
+ANPR cameras provide additional trip information, besides simple link counts, that has contibuted towards  estimating travel demands and traffic states.
+extended traditional methods of  travel demands
 
-There are 267 ANPR cameras in the area of Tyne and Wear, United Kingdom, .
-Scans made by these cameras are stored in a central database by the UTMC of Tyne and Wear and then used to compute travel times across particular links of interest in the road network.
-These can be major roads that see high volumes of traffic, or road segments that can be more prone to traffic jams.
-Average journey times can be then conveyed back to the drivers by the way of Variable Message Signs (VMS) or through .
-Even though the cameras only cover a very small subset of the entire road network, it provides valuable.
+Even though the cameras only cover a very small subset of the entire road network, the network of cameras provides a valuable time-varying picture of travel demands and traffic state.
 
-However, Furthermore, ANPR cameras can either be fixed or
+In ~\cite{Castillo2008}, the authors used partial trip information from plate number scans to compute travel demands, in the form of a trip matrix, and estimate link flows based on conservatin laws, prior and path flow disaggregation techniques.
+
+% However, Furthermore, ANPR cameras can either be fixed or
 
@@
 
+Although a lot of research has gone into 
 but, to the best of our knowledge, little research has gone into doing so with number plate data. Therefore, the contributions of this paper are twofold:

### 2018-01-25 trips.tex
Added mathematical definitions and concepts.

@@
-\section{Identifying Trips In Number Plate Data}
+\section{Identifying trips in number plate data}
 
-\subsection{Definitions}
+Trip data, identified from number plate data, has the potential to unlock a number of new applications for plate scanned data.
+In this section we are concerned with the methodology that allow us to group multiple camera observations of the same vehicle into one or several trips performed by that vehicle.
 
-The \emph{ith} sighting of vehicle \emph{k} can be defined as:
+The \emph{ith} sighting of vehicle \emph{k} can be defined as a tuple pair:
 
-\begin{equation} \label{e2.1.1}
-s^{k}_{i} = \left(camera, time\right)
+\begin{equation} \label{e.sighting}
+s^{k}_{i} = \left(c, t\right)
 \end{equation}
 
-Trip
-\begin{equation} \label{e2.1.2}
-trip^{k}_{u} = \left(s^{k}_{(1)}, s^{k}_{(2)}, \dots , s^{k}_{(n)}\right)
+where \emph{c} is an integer that uniquely identifies a camera, and \emph{t} is a scalar representing a point in time (e.g.\ a timestamp).
+
+An ordered sequence of sightings of vehicle \emph{k} defines the \emph{uth} trip, or journey, of \emph{k}:
+
+\begin{equation} \label{e.trip}
+w^{k}_{u} = \left(s^{k}_{(1)}, s^{k}_{(2)}, \dots , s^{k}_{(n)}\right)
 \end{equation}
 
-where \( n \) is the number of sightings, i.e.\ length of the trip.
+where \( n \) is the number of sightings, i.e.\ length of the trip. A trip of \emph{k} is valid if the following conditions are met:
 
-Trips by the same vehicle can be distinguished
+\begin{eqnarray}
+n &\ge& 1 , \\ \label{e.trip.constraints.1}
+t_{(i)}^{k} - t_{(i-1)}^{k} &>& \tau \label{e.trip.constraints.2}
+\end{eqnarray}
 
+Condition \ref{e.trip.constraints.1} is straightforward and tells us that every trip should have at least one sighting.
+The second condition \ref{e.trip.constraints.1} states that every two consecutive sightings in a trip should occur within a maximum time interval $\tau$.
+It's purpose is to differentiate trips made by the same vehicle and is essentially defining a maximum travel time between each observation.
+Evidently, this condition only makes sense for trips of length 2 or greater.
 From the \( uth \) trip of vehicle \( k \), we can calculate the corresponding journey time sequence, of length \(n-1\):
 
-\begin{equation} \label{e2.1.3}
-journey\ time^{k}_{u} = \left(time^{k}_{2} - time^{k}_{1}, \ldots, time^{k}_{n} - time^{k}_{n-1} \right)
+\begin{equation} \label{e.journeytime}
+jt^{k}_{u} = \left(t^{k}_{2} - t^{k}_{1}, \ldots, t^{k}_{n} - t^{k}_{n-1} \right)
 \end{equation}
 
-A trip of length \( \ge 2 \) is valid if all partial journey times are lower than a threshold \( \tau \):
+We can therefore re-write constraint~\ref{e.trip.constraints.2} as:
 
-\begin{equation} \label{e2.1.4}
-journey\ time^{k}_{u(i)} < \tau \ , \ \forall i \in journey\ time^{k}_{u}
+\begin{equation} \label{e.valid.trip.jt}
+jt^{k}_{u(i)} < \tau \ , \ \forall i \in jt^{k}_{u(i)}
 \end{equation}
 
-Realistically there should also be a lower bound for each journey time.
+However, realistically, there should also be a lower bound for each travel time, which would help identify unplausible trips in the database (vehicles travelling long distances in very short amounts of time), that may be caused by innacurate plate scans or vehicles travelling with cloned plate numbers.
 
-Clock Synchronisation
+\subsection{Choosing $\tau$}
 
-\subsection{Considerations}
+\subsection{Errors in plate scanning}
+
+\subsection{Duplicate scannings}
+
+\subsection{Clock synchronisation}
 

### 2018-01-26 trips.tex
Changed beggining of section: definitions

@@
 
-Trip data, identified from number plate data, has the potential to unlock a number of new applications for plate scanned data.
-In this section we are concerned with the methodology that allow us to group multiple camera observations of the same vehicle into one or several trips performed by that vehicle.
+Trip data, identified from number plate data, has the potential to unlock a number of new applications for urban traffic monitoring and control.
+In this section we are concerned with the methodology that allows multiple camera observations of the same vehicle to be grouped into one or several trips of that vehicle.
 
@@
 
-where \( n \) is the number of sightings, i.e.\ length of the trip. A trip of \emph{k} is valid if the following conditions are met:
-
-\begin{eqnarray}
-n &\ge& 1 , \\ \label{e.trip.constraints.1}
-t_{(i)}^{k} - t_{(i-1)}^{k} &>& \tau \label{e.trip.constraints.2}
-\end{eqnarray}
-
-Condition \ref{e.trip.constraints.1} is straightforward and tells us that every trip should have at least one sighting.
-The second condition \ref{e.trip.constraints.1} states that every two consecutive sightings in a trip should occur within a maximum time interval $\tau$.
-It's purpose is to differentiate trips made by the same vehicle and is essentially defining a maximum travel time between each observation.
-Evidently, this condition only makes sense for trips of length 2 or greater.
-From the \( uth \) trip of vehicle \( k \), we can calculate the corresponding journey time sequence, of length \(n-1\):
+where \( n \) is the number of sightings, i.e.\ length of the trip. Moreover, for a given trip, we can calculate the corresponding journey time sequence, of length \(n-1\):
 
@@
 
-We can therefore re-write constraint~\ref{e.trip.constraints.2} as:
+We consider a trip of \emph{k} valid if the following conditions are met:
 
-\begin{equation} \label{e.valid.trip.jt}
-jt^{k}_{u(i)} < \tau \ , \ \forall i \in jt^{k}_{u(i)}
-\end{equation}
+\begin{align}
+n &\ge 1 , \label{e.trip.constraints.1} \\
+\tau_{(i)} &< jt^{k}_{u(i)} < \Tau_{(i)} \ , \ \forall i \in jt^{k}_{u(i)} \label{e.trip.constraints.2}
+%\tau_{(i)} &>& 0
+%\tau &<& t_{(i)}^{k} - t_{(i-1)}^{k} < \Tau \label{e.trip.constraints.2}
+\end{align}
 
-However, realistically, there should also be a lower bound for each travel time, which would help identify unplausible trips in the database (vehicles travelling long distances in very short amounts of time), that may be caused by innacurate plate scans or vehicles travelling with cloned plate numbers.
+The first condition~\ref{e.trip.constraints.1} is straightforward and specifies that every trip should have at least one sighting.
+The second condition~\ref{e.trip.constraints.2} defines a minimum and maximum travel times between consecutive observations.
+Its purpose is twofold:
+\begin{enumerate*}[label=(\roman*)]
+  \item first, to allow trips made by the same vehicle to be differentiated.
+Given two sightings of \emph{k}, one occurring now and another 6 hours ago, we can interpret them as belonging to different trips of \emph{k};%every two consecutive sightings in a trip should occur within a maximum time interval $\Tau$.
+This allows trips made by the same vehicle to be differentiate.
+  \item second, it allows unplausible trips to be identified.
+An unplausible trip can result from observing \emph{k} at a given camera and then a few seconds later at a second camera, several miles apart.
+Two explanations are common, either one of the cameras made a detection error, or there is another vehicle with a cloned plate number travelling in the road network.
+\end{enumerate*}
 
-\subsection{Choosing $\tau$}
+Evidently, condition~\ref{e.trip.constraints.2} is only valid for trips of length 2 or greater. Furthermore,
+
+How to estimate the values of $\Tau$ and $\tau$ remains
+
+% This would help identify unplausible trips that otherwise resulted from vehicles travelling long distances in very short amounts of time, but that were instead caused by innacurate plate scans or vehicles with cloned plate numbers.
+%
+% We can re-write constraint~\ref{e.trip.constraints.2} using the definition of journey time:
+%
+% \begin{equation} \label{e.valid.trip.jt}
+%
+% \end{equation}
 

### 2018-01-26 trips.tex
A few improvements in maths and text

@@
 Trip data, identified from number plate data, has the potential to unlock a number of new applications for urban traffic monitoring and control.
-In this section we are concerned with the methodology that allows multiple camera observations of the same vehicle to be grouped into one or several trips of that vehicle.
+In this section we are concerned with the methodology for grouping multiple camera observations of the same vehicle into one or several trips of that vehicle.
 
-The \emph{ith} sighting of vehicle \emph{k} can be defined as a tuple pair:
+The \emph{ith} sighting of vehicle \emph{k} can be defined as an unordered pair:
 
 \begin{equation} \label{e.sighting}
-s^{k}_{i} = \left(c, t\right)
+s^{k}_{i} = \{ c, t \}
 \end{equation}
@@
 
-An ordered sequence of sightings of vehicle \emph{k} defines the \emph{uth} trip, or journey, of \emph{k}:
+An ordered sequence of sightings of vehicle \emph{k} defines the \emph{uth} trip of \emph{k}:
 
@@
 
-where \( n \) is the number of sightings, i.e.\ length of the trip. Moreover, for a given trip, we can calculate the corresponding journey time sequence, of length \(n-1\):
+where \( n \) is the number of sightings, i.e.\ length of the trip.
+Moreover, for a given trip, we can calculate the corresponding journey time sequence, of length \(n-1\), as the time differences of consecutive sightings:
 
 \begin{equation} \label{e.journeytime}
-jt^{k}_{u} = \left(t^{k}_{2} - t^{k}_{1}, \ldots, t^{k}_{n} - t^{k}_{n-1} \right)
+jt^{k}_{u} = \left(t^{k}_{(2)} - t^{k}_{(1)}, \ldots, t^{k}_{(n)} - t^{k}_{(n-1)} \right)
 \end{equation}
@@
 \tau_{(i)} &< jt^{k}_{u(i)} < \Tau_{(i)} \ , \ \forall i \in jt^{k}_{u(i)} \label{e.trip.constraints.2}
-%\tau_{(i)} &>& 0
-%\tau &<& t_{(i)}^{k} - t_{(i-1)}^{k} < \Tau \label{e.trip.constraints.2}
 \end{align}
@@
   \item first, to allow trips made by the same vehicle to be differentiated.
-Given two sightings of \emph{k}, one occurring now and another 6 hours ago, we can interpret them as belonging to different trips of \emph{k};%every two consecutive sightings in a trip should occur within a maximum time interval $\Tau$.
-This allows trips made by the same vehicle to be differentiate.
+For instance, given two consecutive sightings of \emph{k} three hours apart, we want to interpret them as belonging to different trips of \emph{k};
   \item second, it allows unplausible trips to be identified.
-An unplausible trip can result from observing \emph{k} at a given camera and then a few seconds later at a second camera, several miles apart.
+For example, an unplausible trip can result from observing \emph{k} at a given camera and then a few seconds later at a second camera, several miles apart.
 Two explanations are common, either one of the cameras made a detection error, or there is another vehicle with a cloned plate number travelling in the road network.
-\end{enumerate*}
+\end{enumerate*} Evidently, condition~\ref{e.trip.constraints.2} is only valid for trips of length two or greater.
 
-Evidently, condition~\ref{e.trip.constraints.2} is only valid for trips of length 2 or greater. Furthermore,
+Furthermore, condition ~\ref{e.trip.constraints.1} implies that no two sightings in a trip can be the values of $\tau$ and $\Tau$
 
 How to estimate the values of $\Tau$ and $\tau$ remains
-
-% This would help identify unplausible trips that otherwise resulted from vehicles travelling long distances in very short amounts of time, but that were instead caused by innacurate plate scans or vehicles with cloned plate numbers.
-%
-% We can re-write constraint~\ref{e.trip.constraints.2} using the definition of journey time:
-%
-% \begin{equation} \label{e.valid.trip.jt}
-%
-% \end{equation}
 
@@
 \subsection{Clock synchronisation}
-
-% \begin{Definition}{\rm We describe the two methods in \S 1.2. In \S\ 1.3. we
-% discuss
-% some remaining details.}
-% \end{Definition}
-%
-% Our purpose here is to examine the nonnumerical complexity of the
-% sparse elimination algorithm given in  \cite{BANKSMITH}.
-% As was shown there, a general sparse elimination scheme based on the
-% bordering algorithm requires less storage for pointers and
-% row/column indices than more traditional implementations of general
-% sparse elimination.  This is accomplished by exploiting the m-tree,
-% a particular spanning tree for the graph of the filled-in matrix.
-% Several good ordering algorithms (nested dissection and minimum degree)
-% are available for computing $P$  \cite{GEORGELIU}, \cite{ROSE72}.
-% Since our interest here does not
-% focus directly on the ordering, we assume for convenience that $P=I$,
-% or that $A$ has been preordered to reflect an appropriate choice of $P$.
-%
-% \begin{equation} \label{1.2}
-%  L_{k-1}v = c
-% \end{equation}
-% and setting
-% \begin{eqnarray}
-% \ell &=& D^{-1}_{k-1}v , \\
-% \delta &=& \alpha - \ell^{t} v .
-% \end{eqnarray}
-%
-% \begin{figure}
-% \vspace{14pc}
-% \caption{This is a figure 1.1.}
-% \end{figure}

### 2018-01-26 trips.tex
Improved maths, final draft for now. Needs revision/feedback by other authors.

@@
 
-Furthermore, condition ~\ref{e.trip.constraints.1} implies that no two sightings in a trip can be the values of $\tau$ and $\Tau$
+Furthermore, if we consider all valid trips of vehicle \emph{k}:
 
-How to estimate the values of $\Tau$ and $\tau$ remains
+\begin{align}
+W^{k} = \left( w^{k}_{(1)}, w^{k}_{(2)}, \ldots, w^{k}_{(N)} \right) \label{e.trip.history}
+\end{align}
+
+where \emph{N} is the number of trips of \emph{k}, then there should be no two trips containing the same sighting:
+
+\begin{align}
+s^{k}_{(u(i))} \neq s^{k}_{(v(j))}, \forall u,v &= 1, 2, \ldots, N \ , \ u \neq v,  \label{e.trip.history.constraint} \\
+\forall i,j &= 1, 2, \ldots, n \ , \ i \neq j \nonumber
+\end{align}
+
+where, we consider two sightings of vehicle \emph{k} to be different simply if they were observed at two different points in time:
+
+\begin{equation} \label{e.sighting}
+t^{k}_{i} \ne t^{k}_{j} \Rightarrow s^{k}_{i} \ne s^{k}_{j} \ , \ i \ne j
+\end{equation}
+
+When identifying trips from GPS traces, . It is easy to see that both $\tau$ and $\Tau$ should be functions of the
+
+but, how to estimate the values of $\Tau$ and $\tau$ remains a. For the rest of this work, we will consider these parameters to be fixed.
 

### 2018-01-29 intro.tex
Added new intro.

@@
 
-Automatic number plate recognition (ANPR) cameras are actively employed in urban traffic environments and play an important role in day-to-day intelligent transportation systems.
-They can be used by Urban Traffic Control and Management (UTMC) entities; by comissioned highway agencies in electronic toll collection; or by law enforcement organisations in the detection of speeding vehicles and validation of number plate registrations.
-The wide diversity of applications, paired with the large improvement in price-to-performance ratio of ANPR hardware and software systems, has resulted in the increase of installed ANPR cameras across the UK~\cite{SurveyITS2011, EvolutionUTMC2013}.
+The volume of traffic on our roads has been growing steadily for over 25 years, both in terms of the number of vehicles on the road -- increasing by 40.6\% in the UK~\cite{noVehiocles} -- and the distances covered -- 325.5 billion miles driven in the UK in the year ending September 2017 which is up nearly 30\% in the last 25 years~\cite{distance}.
+This is placing ever more burden on the road infrastructure along with those who police and manage it.
+In order to better understand how we can deal with this increase in demand we need to better understand how the road network is being used.
+By understanding road usage we can better deal with congestion, handle traffic incidents, plan road modifications and deal with illegal acts on the roads.
 
-In the area of Tyne and Wear, United Kingdom, there are over 250 active ANPR cameras.
-Scans made by these cameras are stored in a central database by the UTMC of Tyne and Wear and then used to compute travel times across particular links of interest in the road network.
-These are usually major roads that see high volumes of traffic, or road segments more prone to traffic jams.
-Average journey times can then be conveyed back to the drivers by the way of Variable Message Signs (VMS) or web based applications.
-ANPR cameras provide additional trip information, besides simple link counts, that has contibuted towards  estimating travel demands and traffic states.
-extended traditional methods of  travel demands
+In a utopian model we would have full disclosure of all journeys made by all vehicles on the road infrastructure.
+However, this has numerous ethical and technical issues.
+From an ethical standpoint should we be allowed to know where all vehicles are at any given point in time.
+From a technical point of view, although every vehicle could be fitted with a GPS tracker -- costly in its own right -- there would still exist the issue of how we would collect and stream all of this data for future processing.
+Alternatively one can view the problem the other way around and rather than tracking individual vehicles look at collecting information by observing vehicles passing points within the road networks.
+A prime example of this approach are Automatic Number-plate Recognition (ANPR) cameras.
+These cameras are a combination of digital camera coupled with Artificial Intelligence to identify number-plates within the image and convert these into strings of characters.
+ANPR cameras are normally fixed in location\footnote{Although cameras can be in a vehicle and moved from location to location.} able to view all vehicles passing that location.
 
-Even though the cameras only cover a very small subset of the entire road network, the network of cameras provides a valuable time-varying picture of travel demands and traffic state.
+For ANPR the problem now becomes that of recovering as much information about vehicle's journey as possible from the limited number of observations.
+ANPR cameras are normally located on major roads and interchanges, however, this only covers a tiny fraction of the road network.
+We can, though, estimate routes between cameras by understanding the distances between cameras and the most ``sensible'' routes between them.
+This allows us, given a set of ANPR sightings of the same vehicle, to produce a ``most likely'' route for that journey.
+It should be noted that we cannot determine the actual start and end of the journey as these will happen in areas not covered by ANPR.
+It should also be noted that for ethical reasons it is not normal to obtain actual number-plates, but rather the hash of these.
+Though, for most situations this will suffice.
+
+Once we have a set of sightings of a vehicle using ANPR, we now need to convert these into actual journeys.
+The first requirement is to identify individual journeys.
+Although this can't be done with certainty we can apply general rules to distinguish one journey from the next.
+For example if two sightings are made from ANPR cameras which are connectable by a ``sensible'' route\footnote{Here ``sensible'' implies that a route between cameras A and B would not need to go through a third camera C.} in a time interval which is ``sensible'' then these can be determined to be part of the same journey.
+However, if the timings between two sightings is significantly longer than what would be expected then this would imply that the vehicle stopped between these two cameras and that the latte sighting is part of a new journey.
+
+The process of journey identification needs to be performed on dirty data which contains numerous impurities which need to be handled. These include:
+
+\begin{itemize}
+	\item {\bf Number-plate miss-reads:} Although ANPR cameras have accuracies of around {\color{red}99\%}, miss-reads are possible.
+This can lead to sightings being missed or vehicles being wrongly sighted in locations.
+	\item {\bf Timing errors:} The time-stamps of sightings could be erroneous.
+The minor side of this is implausible journey times, though, more seriously, this can lead to reordering the set of cameras on a particular journey.
+	\item {\bf Clones number-plates:} For various reasons a number-plate may be cloned and used on a different vehicle.
+This can lead to impossible journeys and journeys that the real vehicle did not make.
+\end{itemize}
+
+Once journeys have been identified from the sightings we can then progress by using these journeys to identify higher-order issues within the road network.
+In this paper we demonstrate how we can use this journey information in order to identify the most likely class each vehicle is a member of.
+By clustering over such characteristics as how many journeys are made each day, average length of journeys, the number of different ANPR cameras seen in a day and the times when journeys are made we can cluster vehicles into buses, taxis, commuters and delivery vehicles.
+
+The rest of this paper is presented as follows.
+In Section \ref{s.related} we discuss related work.
+Section \ref{s.ncl} we presents the ANPR data for the Newcastle area.
+Our process for identifying individual journeys is presented in Section \ref{s.trips} while Section \ref{s.classification} presents our classification approach.
+We present results in Section \ref{s.results} before offering conclusions and future directions in Section \ref{s.conclusions}.
+
+\section{Related Work}
+\label{s.related}
 
@@
 
+\section{Tyne and Wear ANPR Data}
+\label{s.ncl}
+Automatic number plate recognition (ANPR) cameras are actively employed in urban traffic environments and play an important role in day-to-day intelligent transportation systems.
+They can be used by Urban Traffic Control and Management (UTMC) entities; by comissioned highway agencies in electronic toll collection; or by law enforcement organisations in the detection of speeding vehicles and validation of number plate registrations.
+The wide diversity of applications, paired with the large improvement in price-to-performance ratio of ANPR hardware and software systems, has resulted in the increase of installed ANPR cameras across the UK~\cite{SurveyITS2011, EvolutionUTMC2013}.
+
+In the area of Tyne and Wear, United Kingdom, there are over 250 active ANPR cameras.
+Scans made by these cameras are stored in a central database by the UTMC of Tyne and Wear and then used to compute travel times across particular links of interest in the road network.
+These are usually major roads that see high volumes of traffic, or road segments more prone to traffic jams.
+Average journey times can then be conveyed back to the drivers by the way of Variable Message Signs (VMS) or web based applications.
+ANPR cameras provide additional trip information, besides simple link counts, that has contibuted towards  estimating travel demands and traffic states.
+extended traditional methods of  travel demands
+
+Even though the cameras only cover a very small subset of the entire road network, the network of cameras provides a valuable time-varying picture of travel demands and traffic state.
+
 % In this paper, we consider the solution of the $N \times

### 2018-01-29 paper.tex
Added new packages
Added new intro.

@@
 \usepackage{parskip}
+\usepackage[inline]{enumitem}
+\usepackage{amsmath}
+\usepackage{color}
+
+\newcommand{\Tau}{\mathrm{T}}
 
@@
 
+\bibitem{noVehiocles}
+Department for Transport {\em Provisional Road Traffic Estimates Great Britain: October 2016-September 2017}.
+Department for Transport WebSite: https://www.gov.uk/government/statistics/provisional-road-traffic-estimates-great-britain-october-2016-to-september-2017.
+
+\bibitem{distance}
+Department for Transport {\em Vehicle Licensing Statistics: Quarter 3 (Jul-Sep) 2017}.
+Department for Transport WebSite: https://www.gov.uk/government/statistics/vehicle-licensing-statistics-july-to-september-2017.
+
 \bibitem{SurveyITS2011}

### 2018-01-29 trips.tex
Added new intro.

@@
 \section{Identifying trips in number plate data}
-
+\label{s.trips}
 Trip data, identified from number plate data, has the potential to unlock a number of new applications for urban traffic monitoring and control.

### 2018-01-29 src/as_trips.R
Added src code for gathering results

@@ -1,0 +1,98 @@
+#!/usr/bin/env Rscript
+suppressWarnings(suppressMessages(library(tidyverse)))
+
+args = commandArgs(trailingOnly=TRUE)
+
+if (length(args)!=1) {
+  stop("One argument only must be supplied: threshold value", call.=FALSE)
+} else if (length(args)==1) {
+  # default output file
+  threshold = as.integer(args[1])
+}
+
+start.time = Sys.time()
+
+cat("Threshold = ", threshold, "\n")
+
+cat("\nLoading data ..", "\n")
+
+tmpenv <- new.env()
+
+load(file = "data/np_data.RData", envir = tmpenv)
+
+attach(tmpenv)
+
+#print(head(np_data))
+
+as.trips = function(rawdata, threshold) {
+  start.time.1 = Sys.time()
+  
+  tmp = 
+    rawdata %>%
+    group_by(Tag) %>%
+    mutate(JourneyTime = difftime(Timestamp, lag(Timestamp, 1)))
+    
+  cat("Journey Time - elapsed time:", format(difftime(Sys.time(), start.time.1)), "\n", sep = " ")
+  
+  start.time.2 = Sys.time()
+  
+  tmp = 
+    tmp %>%
+    mutate(Trip = ifelse(JourneyTime > threshold | is.na(JourneyTime), TRUE, FALSE))
+  
+  cat("Trip (ifelse) - elapsed time:", format(difftime(Sys.time(), start.time.2)), "\n", sep = " ")
+  
+  start.time.3 = Sys.time()
+  
+  tmp = 
+    tmp %>%
+    mutate(Trip = cumsum(Trip))
+  
+ # print(head(tmp))
+  
+ # print(nrow(tmp))
+  
+  cat("Trip (cumsum) - elapsed time:", format(difftime(Sys.time(), start.time.3)), "\n", sep = " ")
+  
+  start.time.4 = Sys.time()
+  
+  tmp = 
+    tmp %>%
+    group_by(Tag, Trip) %>%
+    mutate(isDuplicate = (CameraId == lag(CameraId))) %>%
+    mutate(isDuplicate = ifelse(is.na(isDuplicate), FALSE, isDuplicate))  
+    
+  cat("Group, 2 x mutate isDuplicate - elapsed time:", format(difftime(Sys.time(), start.time.4)), "\n", sep = " ")
+  
+  start.time.5 = Sys.time()
+  
+  tmp = 
+    tmp %>%
+    filter(!isDuplicate) %>%
[…]

### 2018-01-29 src/get_features.R

@@ -1,0 +1,141 @@
+#!/usr/bin/env Rscript
+suppressWarnings(suppressMessages(library(tidyverse)))
+
+args = commandArgs(trailingOnly=TRUE)
+
+if (length(args)!=1) {
+  stop("One argument only must be supplied: threshold value", call.=FALSE)
+} else if (length(args)==1) {
+  # default output file
+  threshold = as.integer(args[1])
+}
+
+start.time = Sys.time()
+
+cat("Threshold = ", threshold, "\n")
+
+cat("\nLoading data ..", "\n")
+
+tmpenv <- new.env()
+
+load(file = paste("data/trips", threshold, "RData", sep ="."), envir = tmpenv)
+
+attach(tmpenv)
+
+print(head(trips))
+
+
+as.features = function(rawdata, threshold) {
+  # ---
+  # ---
+  # 1
+  # ---
+  # ---
+  
+  cat("\nComputing features (step 1 - mutate)", "\n")
+  start.time.1 = Sys.time()
+  
+  features =
+    rawdata %>%
+    group_by(Tag, Trip) %>%
+    mutate(Origin = CameraId[1]) %>%
+    mutate(Destination = ifelse(Length > 1, CameraId[Length], NA)) %>%
+    summarise(
+      Day = lubridate::day(Timestamp[1]),
+      Origin = Origin[1],
+      Destination = Destination[1],
+      StartTime = Timestamp[1],
+      EndTime = Timestamp[Length[1]],
+      Length = Length[1],
+      Route = paste(CameraId, collapse = "-"),
+      OD = paste(Origin[1], Destination[1], collapse = "-")
+      )
+  #unite("OD", c(Origin, Destination), remove = FALSE) %>%
+  
+  cat("Elapsed time:", format(difftime(Sys.time(), start.time.1)), "\n", sep = " ")
+  
+  # ---
+  # ---
+  # 2
+  # ---
+  # ---
+  
+  cat("\nComputing features (step 2 - mutate RestTime)", "\n")
+  start.time.2 = Sys.time()
+  
+  features =
+    features %>%
+    group_by(Tag, Day) %>%
+    mutate(RestTime = as.numeric(difftime(StartTime, lag(EndTime, 1), units = "hours")))
+  
+  cat("Elapsed time:", format(difftime(Sys.time(), start.time.2)), "\n", sep = " ")
+  
+
+  # ---
+  # ---
+  # 3
+  # ---
+  # ---
+  
[…]

### 2018-01-29 intro.tex
Rephrased a few sentences

--- 2018-01-25 intro.tex
@@
+% !TEX root = paper.tex
 \section{Introduction}
@@
 They can be used by Urban Traffic Control and Management (UTMC) entities; by comissioned highway agencies in electronic toll collection; or by law enforcement organisations in the detection of speeding vehicles and validation of number plate registrations.
-The wide diversity of applications, paired with the large improvement in price-to-performance ratio of ANPR hardware and software systems, has resulted in the increase of installed ANPR cameras across the UK~\cite{SurveyITS2011, EvolutionUTMC2013}.
+The wide diversity of applications, paired with the large improvements in price-to-performance ratios of ANPR hardware and software systems, has resulted in the increase of installed ANPR cameras across the UK~\cite{EvolutionUTMC2013, SurveyITS2011}.
 
@@
 Average journey times can then be conveyed back to the drivers by the way of Variable Message Signs (VMS) or web based applications.
-ANPR cameras provide additional trip information, besides simple link counts, that has contibuted towards  estimating travel demands and traffic states.
-extended traditional methods of  travel demands
+ANPR cameras provide additional trip information, besides simple link counts, that has
 
-Even though the cameras only cover a very small subset of the entire road network, the network of cameras provides a valuable time-varying picture of travel demands and traffic state.
+ estimating travel demands and traffic states. extended traditional methods of  travel demands
 
-In ~\cite{Castillo2008}, the authors used partial trip information from plate number scans to compute travel demands, in the form of a trip matrix, and estimate link flows based on conservatin laws, prior and path flow disaggregation techniques.
+Even though the cameras only cover a very small subset of the entire road network, the network of cameras provides a valuable time-varying picture of travel demands and traffic states.
+
+In ~\cite{Castillo2008}, the authors used partial trip information from plate number scans to compute travel demands, in the form of a trip matrix, and estimate link flows from a set of constraints composed of traffic conservation laws, prior knowledge and path flow disaggregation techniques, such as Stochastic User Equilibrium (SUE).
 
@@
 Number plate data is not as popular, for urban traffic monitoring and prediction , as other sources of data, namely GPS traces or floating vehicle data.
+Previous works trips from GPS traces, .
 
-Although a lot of research has gone into 
-but, to the best of our knowledge, little research has gone into doing so with number plate data. Therefore, the contributions of this paper are twofold:
+Although significant research has gone into identifying trips from GPS traces, to the best of our knowledge, little research has gone into doing so with number plate data.
+Therefore, the contributions of this paper are twofold:
 
-\begin{itemize}
+\begin{enumerate}
   \item
-\end{itemize}
-
-% In this paper, we consider the solution of the $N \times
-% N$ linear
-% system
-% \begin{equation} \label{e1.1}
-% A x = b
-% \end{equation}
-% where $A$ is large, sparse, symmetric, and positive definite.  We consider
-% the direct solution of (\ref{e1.1}) by means of general sparse Gaussian
-% elimination.  In such a procedure, we find a permutation matrix $P$, and
-% compute the decomposition
-% \[
-% P A P^{t} = L D L^{t}
-% \]
-% where $L$ is unit lower triangular and $D$ is diagonal.
+\end{enumerate}

### 2018-01-29 paper.tex
Ordered bibliography by name, added new input file

--- 2018-01-25 paper.tex
@@
 \usepackage{parskip}
+\usepackage[inline]{enumitem}
+\usepackage{amsmath}
+\usepackage{multicol}
+\usepackage[many]{tcolorbox}
+\usepackage{makecell}
+\usepackage{booktabs}
+\usepackage{graphicx}
+
+\newcommand{\Tau}{\mathrm{T}}
 
@@
 
+\input{dataset.tex}
+
 \input{trips.tex}
@@
 
-\bibitem{SurveyITS2011}
-Zhang, Junping, et al. {\em Data-driven intelligent transportation systems: A survey}. IEEE Transactions on Intelligent Transportation Systems 12.4 (2011): 1624-1639.
-
-\bibitem{EvolutionUTMC2013}
-Hamilton, Andrew, et al. {\em The evolution of urban traffic control: changing policy and technology}. Transportation planning and technology 36.1 (2013): 24-43.
+\bibitem{Castillo2010}
+Castillo, Enrique, et al. {\em Optimal use of plate-scanning resources for route flow estimation in traffic networks}. IEEE Transactions on Intelligent Transportation Systems 11.2 (2010): 380-391.
 
@@
 
+\bibitem{EvolutionUTMC2013}
+Hamilton, Andrew, et al. {\em The evolution of urban traffic control: changing policy and technology}. Transportation planning and technology 36.1 (2013): 24-43.
+
 \bibitem{Hazelton2012}
@@
 Transportation Research Part B: Methodological 46.1 (2012): 175-188.
+
+\bibitem{SurveyITS2011}
+Zhang, Junping, et al. {\em Data-driven intelligent transportation systems: A survey}. IEEE Transactions on Intelligent Transportation Systems 12.4 (2011): 1624-1639.
+
+\bibitem{ShortestPath?2015}
+Zhu, Shanjiang, and David Levinson. {\em Do people use the shortest path? An empirical test of Wardrop’s first principle}. PloS one 10.8 (2015): e0134322.
 

### 2018-01-29 trips.tex
Added a few corrections

--- 2018-01-26 trips.tex
@@
-\section{Identifying trips in number plate data}
+% !TEX root = paper.tex
+\section{Trip identification}
 
@@
 
-Furthermore, if we consider all valid trips of vehicle \emph{k}:
+Additionally, if we consider all valid trips of vehicle \emph{k}:
 
@@
 
-where, we consider two sightings of vehicle \emph{k} to be different simply if they were observed at two different points in time:
+where two sightings of vehicle \emph{k} are different simply if they were observed at two different points in time:
 
-\begin{equation} \label{e.sighting}
+\begin{equation} \label{e.sighting.different}
 t^{k}_{i} \ne t^{k}_{j} \Rightarrow s^{k}_{i} \ne s^{k}_{j} \ , \ i \ne j
@@
 
-When identifying trips from GPS traces, . It is easy to see that both $\tau$ and $\Tau$ should be functions of the
+It is easy to see that both $\tau$ and $\Tau$ should be functions of the
 

### 2018-01-29 discussion.tex

% !TEX root = paper.tex
\section{Discussion}\label{s.conclusions}

### 2018-01-29 intro.tex

--- 2018-01-29 intro.tex
@@
 
-In ~\cite{Castillo2008}, the authors used partial trip information from plate number scans to compute travel demands, in the form of a trip matrix, and estimate link flows based on conservatin laws, prior and path flow disaggregation techniques.
+%Even though the cameras only cover a very small subset of the entire road network, the network of cameras provides a valuable time-varying picture of travel demands and traffic states.
+
+In ~\cite{Castillo2008}, the authors used partial trip information from plate number scans to compute travel demands, in the form of a trip matrix, and estimate link flows from a set of constraints composed of traffic conservation laws, prior knowledge and path flow disaggregation techniques, such as Stochastic User Equilibrium (SUE).
 
@@
 Number plate data is not as popular, for urban traffic monitoring and prediction , as other sources of data, namely GPS traces or floating vehicle data.
+Previous works trips from GPS traces, .
 
-Although a lot of research has gone into 
-but, to the best of our knowledge, little research has gone into doing so with number plate data. Therefore, the contributions of this paper are twofold:
+Although significant research has gone into identifying trips from GPS traces, to the best of our knowledge, little research has gone into doing so with number plate data.
+Therefore, the contributions of this paper are twofold:
 
-\begin{itemize}
-  \item
-\end{itemize}
+% \begin{enumerate}
+%   \item
+% \end{itemize}
 
@@
 Average journey times can then be conveyed back to the drivers by the way of Variable Message Signs (VMS) or web based applications.
-ANPR cameras provide additional trip information, besides simple link counts, that has contibuted towards  estimating travel demands and traffic states.
-extended traditional methods of  travel demands
+ANPR cameras provide additional trip information, besides simple link counts, that has
 
-Even though the cameras only cover a very small subset of the entire road network, the network of cameras provides a valuable time-varying picture of travel demands and traffic state.
-
-% In this paper, we consider the solution of the $N \times
-% N$ linear
-% system
-% \begin{equation} \label{e1.1}
-% A x = b
-% \end{equation}
-% where $A$ is large, sparse, symmetric, and positive definite.  We consider
-% the direct solution of (\ref{e1.1}) by means of general sparse Gaussian
-% elimination.  In such a procedure, we find a permutation matrix $P$, and
-% compute the decomposition
-% \[
-% P A P^{t} = L D L^{t}
-% \]
-% where $L$ is unit lower triangular and $D$ is diagonal.
+ estimating travel demands and traffic states. extended traditional methods of  travel demands

### 2018-01-29 paper.tex

--- 2018-01-29 paper.tex
@@
 \usepackage{amsmath}
-\usepackage{color}
+\usepackage[many]{tcolorbox}
+\usepackage{graphicx}
 

### 2018-01-30 intro.tex
Moved text to dataset.tex

@@
 
+%Therefore, number plate scans provide additional trip information, besides simple link counts, that has
+
+ %estimating travel demands and traffic states. extended traditional methods of  travel demands
+
 Although significant research has gone into identifying trips from GPS traces, to the best of our knowledge, little research has gone into doing so with number plate data.
@@
 % \end{itemize}
-
-\section{Tyne and Wear ANPR Data}
-\label{s.ncl}
-Automatic number plate recognition (ANPR) cameras are actively employed in urban traffic environments and play an important role in day-to-day intelligent transportation systems.
-They can be used by Urban Traffic Control and Management (UTMC) entities; by comissioned highway agencies in electronic toll collection; or by law enforcement organisations in the detection of speeding vehicles and validation of number plate registrations.
-The wide diversity of applications, paired with the large improvements in price-to-performance ratios of ANPR hardware and software systems, has resulted in the increase of installed ANPR cameras across the UK~\cite{EvolutionUTMC2013, SurveyITS2011}.
-
-In the area of Tyne and Wear, United Kingdom, there are over 250 active ANPR cameras.
-Scans made by these cameras are stored in a central database by the UTMC of Tyne and Wear and then used to compute travel times across particular links of interest in the road network.
-These are usually major roads that see high volumes of traffic, or road segments more prone to traffic jams.
-Average journey times can then be conveyed back to the drivers by the way of Variable Message Signs (VMS) or web based applications.
-ANPR cameras provide additional trip information, besides simple link counts, that has
-
- estimating travel demands and traffic states. extended traditional methods of  travel demands

### 2018-01-30 dataset.tex
Added figures and text to section about data - unfinished

% !TEX root = paper.tex
\section{Tyne and Wear ANPR Data}\label{s.ncl}

Automatic number plate recognition (ANPR) cameras are actively employed in urban traffic environments and play an important role in day-to-day intelligent transportation systems. They can be used by government subsidised entities in urban traffic management and control; by comissioned highway agencies in electronic toll collection; or by law enforcement organisations in detecting speeding vehicles and validating number plate registrations. The wide diversity of applications, paired with the large improvements in price-to-performance ratios of ANPR hardware and software systems, has resulted in increased investments of ANPR cameras for urban environments~\cite{EvolutionUTMC2013, SurveyITS2011}.

\begin{figure}[t]
\centering
\includegraphics[width=.95\linewidth]{ANPR-overview.pdf}
\caption{}
\label{fig:anpr-overview}
\end{figure}

In the region of Tyne and Wear, United Kingdom, there are over 250 active ANPR cameras. This network of cameras records over 1 million license plate detections every day. Figure~\ref{fig:time-series} shows how the number of sightings fluctuates over a single month (February, 2017). Furtheremore, scans made by these cameras are stored in a central database by the Urban Traffic Management and Control (UTMC) of Tyne and Wear and are then used to compute travel times across particular links of interest in the road network. These are usually major roads that see high volumes of traffic, or road segments more prone to traffic jams. Average journey times can then be conveyed back to the drivers by the way of Variable Message Signs (VMS) or web based applications. Figure~\ref{fig:anpr-overview} represents this interaction.

\begin{figure*}[t]
\centering
\begin{subfigure}[t]{.48\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{observations_per_day.pdf}
  \caption{Total number of scans recorded per day in Tyne and Wear. There is a clear seasonal effect caused by decreasing traffic demands in weekends and increasing traffic volume in weekdays.}
  \label{fig:observations-per-day}
\end{subfigure}\hfill
\begin{subfigure}[t]{.48\textwidth}
  \centering
  \includegraphics[width=.9\linewidth]{observations_per_day_camera.pdf}
  \caption{Number of scans recorded per ANPR camera and day in Tyne and Wear. Inter-camera variability is observed, as some cameras are located in more traffic intensive road sections than others. Decomissioned or temporarily unavailable cameras (due to loss of power, faulty camera, road closed, etc) are depicted at the bottom of the graph.}
  \label{fig:observations-per-camera-day}
\end{subfigure}
\caption{License plate scans recorded by ANPR cameras during February 2017, in the region of Tyne and Wear, United Kingdom.}
\label{fig:time-series}
\end{figure*}

### 2018-01-30 dataset.tex
Finished section about ANPR data

@@
 \section{Tyne and Wear ANPR Data}\label{s.ncl}
-
-Automatic number plate recognition (ANPR) cameras are actively employed in urban traffic environments and play an important role in day-to-day intelligent transportation systems.
-They can be used by government subsidised entities in urban traffic management and control; by comissioned highway agencies in electronic toll collection; or by law enforcement organisations in detecting speeding vehicles and validating number plate registrations.
-The wide diversity of applications, paired with the large improvements in price-to-performance ratios of ANPR hardware and software systems, has resulted in increased investments of ANPR cameras for urban environments~\cite{EvolutionUTMC2013, SurveyITS2011}.
 
@@
 
-In the region of Tyne and Wear, United Kingdom, there are over 250 active ANPR cameras.
-This network of cameras records over 1 million license plate detections every day.
-Figure~\ref{fig:time-series} shows how the number of sightings fluctuates over a single month (February, 2017).
-Furtheremore, scans made by these cameras are stored in a central database by the Urban Traffic Management and Control (UTMC) of Tyne and Wear and are then used to compute travel times across particular links of interest in the road network.
-These are usually major roads that see high volumes of traffic, or road segments more prone to traffic jams.
-Average journey times can then be conveyed back to the drivers by the way of Variable Message Signs (VMS) or web based applications.
-Figure~\ref{fig:anpr-overview} represents this interaction.
+Automatic number plate recognition (ANPR) cameras are actively employed in urban traffic environments and play an important role in day-to-day intelligent transportation systems.
+They can be used by government subsidised entities in urban traffic management and control; by comissioned highway agencies in electronic toll collection; or by law enforcement organisations in detecting speeding vehicles and validating number plate registrations.
+The wide diversity of applications, paired with the large improvements in price-to-performance ratios of ANPR hardware and software systems, has resulted in increased investments of ANPR cameras for urban environments~\cite{EvolutionUTMC2013, SurveyITS2011}.
 
@@
 \end{figure*}
+
+In the region of Tyne and Wear, United Kingdom, there are over 250 active ANPR cameras.
+Over 1 million license plate detections are recorded by these cameras every day.
+Figure~\ref{fig:time-series} shows the number of daily scans recorded over a month (February, 2017).
+Furthermore, every scan is stored in a central database managed by the Urban Traffic Management and Control (UTMC) of Tyne and Wear, and used to compute travel times across particular links of interest in the road network.
+These are usually major roads that see high volumes of traffic, or road segments more prone to traffic jams.
+Average journey times can then be conveyed back to the drivers by the way of Variable Message Signs (VMS) or web based applications.
+Figure~\ref{fig:anpr-overview} represents this interaction.
+
+Number plate data, in its essence, is a stream of events, each representing a vehicle observed by one camera at a specific point in time.
+An excerpt of the data can be found in table~\ref{t:np_data_example}.
+All plate numbers were anonymised by the UTMC through a hashing algorithm before the data was shared.
+Cameras are uniquely identified by an integer and timestamps are relative to each camera's clock.
+However, the cameras are connected through a private network which provides clock synchronisation using the Network Time Protocol (NTP).
+Therefore, the timestamps can be used directly if the synchronisation error is negligible.
+The following additional information is also captured and provided by each camera:
+\begin{enumerate*}[label=(\roman*)]
+  \item the clock synchronisation error (milliseconds);
+  \item the camera's confidence that the identified number plate is the true number plate (percentage);
+  \item the direction of travel, away or towards the camera.
+\end{enumerate*}
+The confidence in the observation is especially useful as it helps diagnosing license plate recognition errors.
+On the other hand, the direction of travel is dependent upon the orientation of the camera, which is not provided.
+Hence, we chose to ignore the latter in this work, and aim to explore this information in future works.
+
+\begin{table}[ht]
+\centering
+\tabcolsep=0.17cm
+\small
+\begin{tabular}{c c c c c}
+  \hline
+  Vehicle & Camera & Timestamp & \thead{Clock \\Error} & Confidence \\
+  \hline
+  169239 & 1031 & 1454284800.26 &   0 & 100 \\
+  12862943 & 18 & 1454284800.97 &   8 &  61 \\
+  16243894 & 22 & 1454284801.46 &   6 &  86 \\
+  4817789 & 52 & 1454284803.43 &  13 &  94 \\
+  5503486 & 110 & 1454284802.19 &  22 &  91 \\
+  15244177 & 115 & 1454284802.83 &  18 &  87 \\
+  6756787 & 146 & 1454284801.53 &  22 &  99 \\
+  8487265 & 2 & 1454284803.88 &  10 &  93 \\
+   \hline
+\end{tabular}
+\caption{Sample of number plate data. Clock error is given in milliseconds and confidence as a percentual value.}
+\label{t:np_data_example}
+\vspace{-0.5cm}
+\end{table}
+
+The main use of ANPR data for the UTMC is average journey time estimation for selected links in the road network.
+Furthermore, several authors have extensively researched how to use used number plate data as an extension on link counts for estimating origin-destination matrices and link flows~\cite{Castillo2010, Castillo2008, Hazelton2012}.
+However, very few works have focused on analysing individual or collective travel patterns from number plate data, particularly across extended periods of time.
+Moreover, there is no consistent conceptual and analytical framework for transforming number plate data into a historical sequence of trips for each vehicle.
+Finally, we believe that trip data, properly identified from number plate data, has the potential to unlock a number of new applications for urban traffic control and law enforcement.
+Thus, in the following section we present a conceptual methodology for grouping multiple camera observations of the same vehicle into one or several trips of that vehicle.

### 2018-01-30 trips.tex
Finished section about ANPR data

@@
 % !TEX root = paper.tex
-\section{Trip identification}
+\section{Trip Identification}\label{s.trips}
 
-Trip data, identified from number plate data, has the potential to unlock a number of new applications for urban traffic monitoring and control.
-In this section we are concerned with the methodology for grouping multiple camera observations of the same vehicle into one or several trips of that vehicle.
-
-The \emph{ith} sighting of vehicle \emph{k} can be defined as an unordered pair:
+Let the \emph{ith} sighting of vehicle \emph{k} be defined as the unordered pair:
 
@@
 
-An ordered sequence of sightings of vehicle \emph{k} defines the \emph{uth} trip of \emph{k}:
+Let an ordered sequence of sightings of vehicle \emph{k} define the \emph{uth} trip of \emph{k}:
 

### 2018-01-30 trips.tex
Added images and text to section trip identification - unfinished

@@
 
-where \( n \) is the number of sightings, i.e.\ length of the trip.
-Moreover, for a given trip, we can calculate the corresponding journey time sequence, of length \(n-1\), as the time differences of consecutive sightings:
+where \( n \) is the length of the trip, i.e.\ the number of sightings.
+Moreover, let the corresponding journey time sequence, of length \(n-1\), be defined as the time difference of consecutive sightings:
 
@@
 \end{enumerate*} Evidently, condition~\ref{e.trip.constraints.2} is only valid for trips of length two or greater.
+Nevertheless, trips can easily be differentiated by first sorting sightings by time of occurrence, then calculating the journey time sequence for the entire sequence and finally comparing each element against $\Tau$.
+An example of a trip identified this way can be seen in Figure~\ref{fig:trip-example}.
 
-Additionally, if we consider all valid trips of vehicle \emph{k}:
+\begin{figure*}[!ht]%
+  \centering
+  \begin{subfigure}[c]{.5\textwidth}
+    \small
+    \tabcolsep=0.09cm
+    \begin{tabular}{c c c c c c c}
+      \hline
+      Vehicle & Camera & Timestamp & Trip & Sighting & \thead{Journey \\Time} & \thead{Trip \\Id}\\
+      \hline
+      2362920 & 1014 & \makecell{2017-02-01 \\ 00:00:06} &   1 &   1 & NA & 21 \\
+      2362920 & 1044 & \makecell{2017-02-01 \\ 00:01:28} &   1 &   2 & 82.38 & 21 \\
+      2362920 &  35 & \makecell{2017-02-01 \\ 00:02:32} &   1 &   3 & 63.50 & 21\\
+      2362920 &  32 & \makecell{2017-02-01 \\ 00:04:38} &   1 &   4 & 125.95 & 21\\
+       \hline
+    \end{tabular}
+    \label{fig:trip-example-table}
+  \end{subfigure}\hfill
+  %\qquad
+  \begin{subfigure}[c]{.48\textwidth}
+    \centering
+    \includegraphics[width=1\linewidth]{trip-example.pdf}
+    \label{fig:trip-example-map}
+  \end{subfigure}\hfill
+  \caption{Example of a trip of length 4. On the left side, w}%
+  \label{fig:trip-example}%
+\end{figure*}
+
+The simplest approaching to choosing the value of $\Tau$ is to pick a fixed empirical value, such as 5 or 10 minutes.
+However, if the distance between two cameras is greater than another origin-destination (od) pair of cameras, then it makes sense that $\Tau$ is relaxed.
+Similarly, if there is an anomaly in the road network, such as a traffic jam, and the routes connecting the two cameras are affected, then the value of $\Tau$ should also be adapted.
+Hence, $\Tau_{(i)}$ should be a function of the distance between the two cameras (or, more accurately, of the top n-routes between these) and the distribution of observed journey times.
+The same rationale can be applied to estimating $\tau$.
+However, whilst devising methods to estimate $\Tau$ and $\tau$ under varying circumstances remains an open problem, it is something that we plan to work in the future.
+Thus, for the purposes of this work, we will consider these parameters to be fixed.
+Nevertheless, the importance of estimating these parameters should not be understated, as they are responsible for differentiating trips and identifying and filtering unplausible trips.
+Any errors that result from will propagate to the results and affect the outcome of the research.
+Figure~\ref{fig:length-dist} provides
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=1\linewidth]{length-dist.pdf}
+  \caption{A}
+  \label{fig:length-dist}
+\end{figure}
+
+\subsection{Duplicate scannings}
+
+To ensure that every trip of vehicle \emph{k} is unique, the sequence of valid trips of vehicle \emph{k}:
 
@@
 
-where \emph{N} is the number of trips of \emph{k}, then there should be no two trips containing the same sighting:
+where \emph{N} is the number of trips of \emph{k}, is such that there should be no two trips containing the same sighting:
 
@@
 
-It is easy to see that both $\tau$ and $\Tau$ should be functions of the
+However, if these
 
-but, how to estimate the values of $\Tau$ and $\tau$ remains a. For the rest of this work, we will consider these parameters to be fixed.
+then the two sightings are deemed as duplicates.
 
 \subsection{Errors in plate scanning}
-
-\subsection{Duplicate scannings}
-
-\subsection{Clock synchronisation}

### 2018-01-31 trips.tex
Added caption to trip example. Almost finished subsection about duplicates (finish math formula). Started last subsection about camera errors

@@
 
-Let the \emph{ith} sighting of vehicle \emph{k} be defined as the unordered pair:
+Let the $i_{th}$ sighting of vehicle \emph{k} be defined as the unordered pair:
 
@@
 
-Let an ordered sequence of sightings of vehicle \emph{k} define the \emph{uth} trip of \emph{k}:
+Let an ordered sequence of sightings of vehicle \emph{k} define the $u_{th}$ trip of \emph{k}:
 
@@
 
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=1\linewidth]{length-dist.pdf}
+  \caption{Distribution of trips per length of trip.}
+  \label{fig:length-dist}
+\end{figure}
+
 \begin{figure*}[!ht]%
@@
   \end{subfigure}\hfill
-  \caption{Example of a trip of length 4. On the left side, w}%
+  \caption{Example of a trip of length 4.
+On the left side the correspoding data table is shown.
+The $u_{th}$ trip of each vehicle is given by the variable \emph{Trip}, whereas the $i_{th}$ sighting is given by variable \emph{Sighting}.
+The variable \emph{JourneyTime} gives the travel time from the previous to current sighting.
+Lastly, the variable \emph{TripId} represents the unique sequence of cameras that describes the trip.
+This allows trips to be grouped and summarised not only in terms of their origins and destination, but also routes.
+On the right side, the same trip is plotted on a map.
+However, the lines do not represent the true route taken by the vehicle, but instead the fastest driving route between sightings.
+Even though no routing information is available for each consecutive pair of sightings, the observed journey times can be compared against the distribution of collective journey times to rank the set of most likely routes chosen (which can be determined for instance by Stochastic User Equilibrium~\cite{Castillo2008}).}%
   \label{fig:trip-example}%
@@
 Similarly, if there is an anomaly in the road network, such as a traffic jam, and the routes connecting the two cameras are affected, then the value of $\Tau$ should also be adapted.
-Hence, $\Tau_{(i)}$ should be a function of the distance between the two cameras (or, more accurately, of the top n-routes between these) and the distribution of observed journey times.
+Hence, $\Tau$ should be a function of the distance between the two cameras (or, more accurately, of the top n-routes between these) and the distribution of observed journey times.
 The same rationale can be applied to estimating $\tau$.
-However, whilst devising methods to estimate $\Tau$ and $\tau$ under varying circumstances remains an open problem, it is something that we plan to work in the future.
+
+On the other hand, a consistent set of models and methodologies for estimating $\Tau$ and $\tau$, under a variety of circumstances and road anomalies, remains an open research problem.
 Thus, for the purposes of this work, we will consider these parameters to be fixed.
-Nevertheless, the importance of estimating these parameters should not be understated, as they are responsible for differentiating trips and identifying and filtering unplausible trips.
-Any errors that result from will propagate to the results and affect the outcome of the research.
-Figure~\ref{fig:length-dist} provides
-
-\begin{figure}[t]
-  \centering
-  \includegraphics[width=1\linewidth]{length-dist.pdf}
-  \caption{A}
-  \label{fig:length-dist}
-\end{figure}
+Nevertheless, the importance of estimating these parameters should not be understated.
+Any errors in trip identification that occur due to poor estimation and filtering methods will propagate and be amplified in posterior analysis done using trip data.
+Therefore, $\Tau$ and $\tau$ estimation is something that needs to be carefully addressed and studied in the future in order to fully understand its impact in the outcome of the research.
+To briefly exemplify this, figure~\ref{fig:length-dist} shows how the number of trips per length of trip varies by fixing $\Tau$ at different empirical values ($\tau$ is ignored).
 
@@
 
-To ensure that every trip of vehicle \emph{k} is unique, the sequence of valid trips of vehicle \emph{k}:
+To ensure that every trip of vehicle \emph{k} is unique in the sequence of all valid trips of vehicle \emph{k}:
 
@@
 
-where \emph{N} is the number of trips of \emph{k}, is such that there should be no two trips containing the same sighting:
+where \emph{N} is the number of trips of \emph{k}, then there should be no two trips containing the same sighting:
 
@@
 
-where two sightings of vehicle \emph{k} are different simply if they were observed at two different points in time:
+However, ANPR can identify the same vehicle multiple times in the same run, if for instance the vehicle is stopped at a junction or traffic light.
+Hence, if two sightings occurred at the same location in a very short period of time, then there is a strong possibility that these are duplicate observations.
+As a simplification, we can assume that a trip should not contain cycles and that no camera should appear twice in the same trip.
+Yet, this assumption ignores cases where a vehicle is required to correct its route by passing through the same location as one previously observed in the same trip.
+Thus, we affirm that two sightings of vehicle \emph{k} are different if they were observed at two different points in time at different locations, or, if observed at the same location, then the time interval must be greater than a parameter $\gamma$, otherwise the two sightings are deemed as duplicates:
 
@@
 
-However, if these
-
-then the two sightings are deemed as duplicates.
+ Although the estimation of $\gamma$ carries similar considerations and consequences as those of estimating $\Tau$ and $\tau$, most duplicates can be identified in consecutive sightings of the same camera within the same trip.
+Even though a poor estimation of $\gamma$ also has impact in error propogation, this decreases substantially after filtering duplicates according to the heuristic above, due to the low occurrence of cycles in trips.
+Nevertheless, this issue needs to be fully investigated in future works.
 
 \subsection{Errors in plate scanning}
+
+Even though number plate recognition cameras have accuracy rates of 99.9\% or higher, on average 1 observation out of 1000 is a misclassified plate number.
+We categorise the errors made by ANPR cameras into two types:
+\begin{enumerate}
+  \item a passing vehicle is not detected by the camera;
+  \item a different vehicle is detected instead.
+\end{enumerate}
+In the first case, there is no observed data.
+The corresponding trip sequence for the affected vehicle will be missing a sighting but we will have no indication of this.
+On the second case, there is a recorded sighting, but this will be assigned to the incorrect vehicle.

### 2018-01-31 classification.tex
Added subsections
Added new intro.
Renamed section and subsection
Renamed sections
Added todo

@@
-\section{Vehicle Classification Using Summarised Trip Information}
+% !TEX root = paper.tex
+\section{Clustering vehicles}\label{s.classification}
+
+One of the most straightfoward applications of trip data is unsupervised learning. Clustering trips to identify
+
+Usou se os seguintes thresholds ...
+
+Filtrou-se todos os confidence blabla..
+
+Filtrou-se duplicates desta forma
+
+Ignorou-se TimestampError porque..
+
+Calculou-se as seguintes features
+
+TABELA de features exemplo
+
+Correlation between features.. we removed three features because of that
+
+\section{Evaluation}\label{s.results}
+
+TABELA de MEANS para o best k e best (ou baseline) threshold?
+
+TABELA DE DIFERENCAS para o best k entre o baseline threshold e os outros thresholds
+
+TABELA de RESULTADOS POR k (e por threshold ou average dos thresholds?), para determinar o melhor k
+
+INTERPRETACAO DE RESULTADOS

### 2018-01-31 classification.tex
Added first results at the end - unformatted, just as a hack, will fix this later on

@@
 INTERPRETACAO DE RESULTADOS
+
+\clearpage
+
+\begin{table*}[t]
+\centering
+\begin{tabular}{rrrrrrr}
+  \hline
+ k & t300 & t450 & t600 & t900 & t1200 & t1800 \\
+  \hline
+1 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 & 0.00 \\
+  2 & 543907.35 & 534453.24 & 534706.04 & 561666.18 & 618943.98 & 737557.49 \\
+  3 & 822985.63 & 789214.80 & 760759.14 & 723082.06 & 719393.29 & 743001.88 \\
+  4 & 913945.86 & 876675.37 & 838741.42 & 781306.76 & 754241.10 & 711478.70 \\
+  5 & 1012979.13 & 943556.95 & 878417.35 & 786755.20 & 745483.33 & 683551.70 \\
+  6 & 1084917.22 & 989249.70 & 905072.13 & 716272.22 & 730215.42 & 651959.66 \\
+  7 & 1113954.86 & 1003744.40 & 852060.06 & 761738.57 & 713320.54 & 633101.34 \\
+  8 & 1061923.12 & 972081.71 & 892493.94 & 776092.12 & 708701.00 & 606811.46 \\
+   \hline
+\end{tabular}
+\caption{Calinski-Harabasz Criterion per k-cluster, per value of threshold}
+\end{table*}
+
+\begin{table}[!ht]
+\centering
+\begin{tabular}{rr}
+  \hline
+ threshold & best k \\
+  \hline
+t300 &   7 \\
+  t450 &   7 \\
+  t600 &   6 \\
+  t900 &   5 \\
+  t1200 &   4 \\
+  t1800 &   3 \\
+   \hline
+\end{tabular}
+\caption{Best k for each threshold}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rr}
+  \hline
+ threshold & value \\
+  \hline
+t300 & 0.91 \\
+  t450 & 0.90 \\
+  t600 & 0.88 \\
+  t900 & 0.84 \\
+  t1200 & 0.79 \\
+  t1800 & 0.71 \\
+   \hline
+\end{tabular}
+\caption{$BetweenSS / TotalSS$ -- Amount of variance in data explained}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrrrrr}
+  \hline
+ threshold & V1 & V2 & V3 & V4 & V5 & V6 & V7 \\
+  \hline
+t300 & 550 & 99091 & 7975 & 39032 & 2478 & 168677 & 338847 \\
+  t450 & 666 & 10380 & 2948 & 45414 & 108877 & 163985 & 309736 \\
+  t600 & 26568 & 865 & 97745 & 332060 & 171024 & 4408 &  \\
+  t900 & 352029 & 74821 & 181295 & 11200 & 1787 &  &  \\
+  t1200 & 189224 & 371318 & 49852 & 4593 &  &  &  \\
+  t1800 & 26484 & 191957 & 388959 &  &  &  &  \\
+   \hline
+\end{tabular}
+\caption{Sizes of each cluster for best k per threshold}
+\end{table}
+
+% latex table generated in R 3.3.1 by xtable 1.8-2 package
+% Wed Jan 31 03:12:43 2018
+\begin{table*}[ht]
+\centering
+\begin{tabular}{rrrrrrrrr}
+  \hline
+ & Trips & AvTrips & AvLength & AvDisOri & AvDisDes & AvFirstH & AvLastH & AvRest \\
+  \hline
+1 & 412.89 & 33.00 & 1.25 & 10.64 & 4.95 & 5.75 & 19.71 & 13.45 \\
+  2 & 36.34 & 3.41 & 1.39 & 3.08 & 1.06 & 9.64 & 16.21 & 6.50 \\
+  3 & 108.34 & 8.92 & 1.27 & 6.78 & 2.63 & 8.20 & 17.77 & 9.48 \\
+  4 & 60.61 & 5.10 & 1.34 & 4.37 & 1.59 & 8.89 & 16.98 & 8.00 \\
+  5 & 210.51 & 16.93 & 1.28 & 10.79 & 4.67 & 6.67 & 18.97 & 12.21 \\
+  6 & 19.75 & 2.56 & 1.34 & 2.37 & 0.73 & 10.95 & 15.17 & 4.17 \\
+  7 & 6.49 & 2.20 & 1.33 & 2.10 & 0.64 & 12.47 & 14.49 & 1.98 \\
+   \hline
+\end{tabular}
+\caption{Centers for threshold = 300}
+\end{table*}
+
+% latex table generated in R 3.3.1 by xtable 1.8-2 package
+% Wed Jan 31 03:12:43 2018
+\begin{table*}[ht]
+\centering
+\begin{tabular}{rrrrrrrrr}
+  \hline
+ & Trips & AvTrips & AvLength & AvDisOri & AvDisDes & AvFirstH & AvLastH & AvRest \\  \hline
+1 & 331.50 & 26.43 & 1.57 & 9.31 & 5.17 & 5.73 & 19.59 & 13.14 \\
+  2 & 86.99 & 7.16 & 1.43 & 5.55 & 2.49 & 8.41 & 17.60 & 9.03 \\
+  3 & 170.08 & 13.69 & 1.47 & 9.20 & 4.67 & 6.89 & 18.69 & 11.65 \\
+  4 & 50.30 & 4.26 & 1.52 & 3.67 & 1.57 & 9.04 & 16.86 & 7.69 \\
+  5 & 30.67 & 2.91 & 1.56 & 2.65 & 1.09 & 9.67 & 16.11 & 6.35 \\
+  6 & 16.99 & 2.30 & 1.46 & 2.13 & 0.77 & 11.12 & 14.99 & 3.81 \\
+  7 & 5.95 & 2.01 & 1.44 & 1.92 & 0.68 & 12.52 & 14.50 & 1.93 \\
+   \hline
+\end{tabular}
+\caption{Centers for threshold = 450}
+\end{table*}
+
+% latex table generated in R 3.3.1 by xtable 1.8-2 package
+% Wed Jan 31 03:12:43 2018
+\begin{table*}[ht]
+\centering
+\begin{tabular}{rrrrrrrrr}
+  \hline
+ & Trips & AvTrips & AvLength & AvDisOri & AvDisDes & AvFirstH & AvLastH & AvRest \\
+  \hline
+1 & 60.90 & 5.04 & 1.58 & 4.13 & 1.95 & 8.78 & 17.20 & 8.24 \\
+  2 & 271.76 & 21.64 & 1.79 & 8.77 & 5.37 & 5.72 & 19.42 & 12.80 \\
+  3 & 34.38 & 3.07 & 1.68 & 2.76 & 1.27 & 9.38 & 16.41 & 6.90 \\
+  4 & 6.39 & 1.92 & 1.52 & 1.83 & 0.69 & 12.45 & 14.51 & 1.99 \\
+  5 & 18.96 & 2.28 & 1.57 & 2.11 & 0.85 & 10.74 & 15.24 & 4.41 \\
+  6 & 131.86 & 10.66 & 1.61 & 7.53 & 4.09 & 7.32 & 18.27 & 10.74 \\
+   \hline
+\end{tabular}
+\caption{Centers for threshold = 600}
+\end{table*}
+
+% latex table generated in R 3.3.1 by xtable 1.8-2 package
+% Wed Jan 31 03:12:43 2018
+\begin{table*}[ht]
+\centering
+\begin{tabular}{rrrrrrrrr}
+  \hline
+ & Trips & AvTrips & AvLength & AvDisOri & AvDisDes & AvFirstH & AvLastH & AvRest \\
+  \hline
+1 & 6.80 & 1.81 & 1.62 & 1.72 & 0.70 & 12.37 & 14.53 & 2.07 \\
+  2 & 38.62 & 3.30 & 1.82 & 2.89 & 1.48 & 9.18 & 16.73 & 7.35 \\
+  3 & 20.89 & 2.25 & 1.73 & 2.09 & 0.96 & 10.31 & 15.52 & 5.08 \\
+  4 & 78.63 & 6.41 & 1.77 & 4.94 & 2.76 & 8.24 & 17.62 & 9.08 \\
+  5 & 180.14 & 14.33 & 2.06 & 7.52 & 4.99 & 6.11 & 18.89 & 11.78 \\
+   \hline
+\end{tabular}
+\caption{Centers for threshold = 900}
+\end{table*}
+
+% latex table generated in R 3.3.1 by xtable 1.8-2 package
+% Wed Jan 31 03:12:43 2018
+\begin{table*}[ht]
+\centering
+\begin{tabular}{rrrrrrrrr}
+  \hline
+ & Trips & AvTrips & AvLength & AvDisOri & AvDisDes & AvFirstH & AvLastH & AvRest \\
+  \hline
+1 & 22.82 & 2.29 & 1.84 & 2.11 & 1.06 & 10.00 & 15.76 & 5.60 \\
+  2 & 7.19 & 1.76 & 1.68 & 1.68 & 0.71 & 12.30 & 14.55 & 2.14 \\
+  3 & 44.85 & 3.73 & 1.90 & 3.16 & 1.73 & 8.98 & 17.02 & 7.76 \\
+  4 & 114.34 & 9.12 & 2.30 & 5.92 & 3.98 & 6.93 & 18.21 & 10.43 \\
+   \hline
+\end{tabular}
+\caption{Centers for threshold = 1200}
+\end{table*}
+
+% latex table generated in R 3.3.1 by xtable 1.8-2 package
+% Wed Jan 31 03:12:43 2018
+\begin{table*}[ht]
+\centering
+\begin{tabular}{rrrrrrrrr}
+  \hline
+ & Trips & AvTrips & AvLength & AvDisOri & AvDisDes & AvFirstH & AvLastH & AvRest \\
+  \hline
+1 & 54.75 & 4.42 & 2.33 & 3.52 & 2.21 & 8.45 & 17.40 & 8.35 \\
+  2 & 24.54 & 2.31 & 1.98 & 2.13 & 1.15 & 9.72 & 16.00 & 6.06 \\
+  3 & 7.52 & 1.69 & 1.77 & 1.61 & 0.72 & 12.23 & 14.58 & 2.22 \\
+   \hline
+\end{tabular}
+\caption{Centers for threshold = 1800}
+\end{table*}
+
+\clearpage

### 2018-01-31 paper.tex
Added figures and text to section about data - unfinished
Finished section about ANPR data
Changed title and added bibliography
Fixed affiliation, added biliography and package

@@
 \usepackage{graphicx}
+\usepackage{subcaption}
+\usepackage{makecell}
+\usepackage{float}
+\usepackage{placeins}
 
 \newcommand{\Tau}{\mathrm{T}}
+
+\graphicspath{{./images/}}
 
@@
 
-\title{\Large Classifying Vehicles From Automatic Number \\ Plate Recognition Camera Scans}
-\author{Pedro M. Pinto Silva \thanks{School of Computing Science, Newcastle University, United Kingdom}
+\title{\Large Clustering Vehicle Trips Identified from Automatic \\ Number Plate Recognition Camera Scans}
+\author{Pedro M. Pinto Silva \thanks{School of Computing, Newcastle University, United Kingdom}
 \and
@@
 
+\bibitem{Clustering}
+Chen, Huiyu, Chao Yang, and Xiangdong Xu. {\em Clustering Vehicle Temporal and Spatial Travel Behavior Using License Plate Recognition Data}. Journal of Advanced Transportation 2017 (2017).
+
 \bibitem{EvolutionUTMC2013}
@@
 Transportation Research Part B: Methodological 46.1 (2012): 175-188.
+
+\bibitem{Kmeans}
+Steinley, Douglas, and Michael J. Brusco. {\em Choosing the number of clusters in Κ-means clustering}. Psychological Methods 16.3 (2011): 285.
 

### 2018-01-31 classification.tex
Added some text and final (hopefully) tables.

--- 2018-01-25 classification.tex
@@
-\section{Vehicle Classification Using Summarised Trip Information}
+% !TEX root = paper.tex
+\section{Clustering vehicles}\label{s.classification}
+
+\begin{table*}[t]
+\centering
+\small
+\begin{tabular}{c c c c c c c c c c c}
+  \hline
+ \thead{Total\\Trips} & \thead{Average\\Trips} & \thead{Average\\Length} & \thead{Average\\Sightings} & \thead{Average\\Distinct\\Origins} & \thead{Average\\Distinct\\Destinations} & \thead{Average\\Distinct\\Routes} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} & \thead{Average\\Hour\\Difference} & \thead{Average\\Rest\\Time} \\
+  \hline
+41 & 3.42 & 1.25 & 5.75 & 2.75 & 1.00 & 3.00 & 15.33 & 19.25 & 3.80 & 3.70 \\
+3 & 1.50 & 1.00 & 1.50 & 1.50 & 0.00 & 1.50 & 14.00 & 14.50 & 0.73 & 0.73 \\
+7 & 2.33 & 1.33 & 3.33 & 2.33 & 0.67 & 2.33 & 11.00 & 13.00 & 2.44 & 2.41 \\
+12 & 2.40 & 1.10 & 3.60 & 2.40 & 0.60 & 2.40 & 14.40 & 16.40 & 2.00 & 1.95 \\
+   \hline
+\end{tabular}
+\caption{Sample of extracted features from trips taken from 15 weekdays of number plate data.}
+\label{t:features}
+\end{table*}
+
+One direct application of trip data is unsupervised learning.
+Previous works have focused on clustering trips in order to identify travel modes or purpose of trip ~\cite{Clustering}.
+Estimating a distribution of trip mode travel, is one of the fundamental steps in traffic modelling .
+In this work we focus on identifying groups of vehicles with similar trip patterns based on frequency and diversity of travel.
+For instance, taxis usually .
+%This in turn, may enable trip modes to be estimated
+However, due to its tiny coverage of the road network, it is not clear if ANPR data can capture the travel characteristics that describe the several groups.
+Thus, we aim with this work to evaluate how good ANPR data is for providing trip information and ultimately that of the underlyng road network.
+In the remaining of this section we provide detail on the choice of parameters and filtering methods used in trip identification, as well as in the choice of features.
+Then, we
+
+Due to distinct traffic behavior during weekends, we chose to consider only trips occurring during weekdays.
+We thus used all number plate data collected during between the 6th and 24th of February and excluded data from the 2 weekends in between.
+Furthermore, as mentioned in section~\ref{s.trips}, we used fixed empirical values for $\Tau$, mostly due to time constraints.
+As such, we ran the trip identification sequence for the following values of $\Tau$: $[5 , 7.5, 10, 15, 20, 30]$ minutes.
+The total number of trips identified during the 3 weeks, for each threshold, was approximately [1.36M, 1.24M, 1.17M, 1.08M, 1.03M, 0.97M] respectively.
+Moreover, we did not set a value for $\tau$, but instead handled unplausible trips by filtering all sightings with confidence below 85\%.
+Duplicates were filtered after identifying consecutive sightings by the same camera occurring within the same trip.
+Clock synchronisation errors were provided in milliseconds with none exceeding 5 seconds.
+These were therefore ignored.
+
+Transforming trips into features that could be used in clustering algorithms was a 3-step process:
+\begin{enumerate}
+  \item First, every trip was summarised as a single row of data. The following information was extracted: length of trip, origin, destination, route, start and end times.
+  \item Second, daily trip information was obtained for each vehicle: number of trips, median of trip lengths, number of sightings, distinct number of origins destinations, and routes, hour of first sighting, hour of last sighting and total rest time between trips.
+  \item Finaly, daily information per vehicle was collapsed into a single row by averaging this information across the 15 days.
+\end{enumerate}
+
+Table~/ref{t:features} depicts a sample of the resulting features vector.
+Each  The aim was to choose features that captured different aspects of trip frequency and diversity.
+However, because a high percentage of trips contain a single sighting.
+1034107 vehicles
+
+We identified correlated features and removed the following features .. because ..
+
+Filtered all cases where TotalTrips <= 3. (HOW MANY?)
+
+We ran the kmeans algorithm:
+ k = 2:8
+ iter.max = 200,
+ runs = 100
+
+\section{Results and Discussion}\label{s.results}
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{c c c c c}
+  \hline
+\thead{Tau\\(min)} & Best k & Betweenss & Whithinss & \thead{Calinski-\\Harabasz} \\
+  \hline
+5 &   8 & 0.92 & 0.09 & 1,116,962 \\
+  7.5 &   7 & 0.90 & 0.10 & 1,003,744 \\
+  10 &   7 & 0.90 & 0.18 &   911,044 \\
+  15 &   6 & 0.86 & 0.19 &   794,557 \\
+  20 &   4 & 0.79 & 0.19 &   754,241 \\
+  30 &   3 & 0.71 & 0.30 &   743,001 \\
+   \hline
+\end{tabular}
+\caption{Kmeans performance comparison for each best k }
+\label{t:tau_comparison}
+\end{table}
+
+\begin{table*}[t]
+\centering
+\small
+\begin{tabular}{c c c c c c c c c c}
+  \hline
+ Cluster &  Size & \thead{Total\\Trips} & \thead{Average\\Trips} & \thead{Average\\Length} & \thead{Average\\Distinct\\Origins} & \thead{Average\\Distinct\\Destinations} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} & \thead{Average\\Rest\\Time} \\
+  \hline
+  1 & 108868 & 30.67 & 2.91 & 1.56 & 2.65 & 1.09 & 9.67 & 16.11 & 6.35 \\
+  2 & 2948 & 170.08 & 13.69 & 1.47 & 9.20 & 4.67 & 6.89 & 18.69 & 11.65 \\
+  3 & 309748 & 5.95 & 2.01 & 1.44 & 1.92 & 0.68 & 12.52 & 14.50 & 1.93 \\
+  4 & 163982 & 16.99 & 2.30 & 1.46 & 2.13 & 0.77 & 11.12 & 14.99 & 3.81 \\
+  5 & 45414 & 50.30 & 4.26 & 1.52 & 3.67 & 1.57 & 9.04 & 16.86 & 7.69 \\
+  6 & 10380 & 86.99 & 7.16 & 1.43 & 5.55 & 2.49 & 8.41 & 17.60 & 9.03 \\
+  7 & 666 & 331.50 & 26.43 & 1.57 & 9.31 & 5.17 & 5.73 & 19.59 & 13.14 \\
+   \hline
+\end{tabular}
+\caption{Clusters sizes and mean centers for $\Tau = 7.5$ minutes}
+\label{t:kmeans_centers_450}
+\end{table*}

### 2018-01-31 classification.tex
Added draft of section 5: Clustering Vehicles

@@
 One direct application of trip data is unsupervised learning.
-Previous works have focused on clustering trips in order to identify travel modes or purpose of trip ~\cite{Clustering}.
-Estimating a distribution of trip mode travel, is one of the fundamental steps in traffic modelling .
-In this work we focus on identifying groups of vehicles with similar trip patterns based on frequency and diversity of travel.
-For instance, taxis usually .
+Previous works have focused on identifying travel modes or purpose of trip from different sources of data, most notably: survey data, GPS data and mobile phone data~\cite{ODMobileData, ClusteringGPS}.
+Determining the distribution of travel modes is one of the fundamental steps in the four-stage model: an essential traffic modelling methodology for transportation planning~\cite{FourStepModel}.
+Although number plate data has been used in~\cite{Clustering}, to identify different categories of trips, these do not differentiate between private or public travel modes and focus for the most part on when the trips occurred.
+Therefore, we aim to identify groups of vehicles with similar trip patterns based on frequency and diversity of travel.
+This can be private vehicles, such as work-home commuters, transit vehicles like buses and taxis, or other types of vehicles such as delivery trucks.
 %This in turn, may enable trip modes to be estimated
-However, due to its tiny coverage of the road network, it is not clear if ANPR data can capture the travel characteristics that describe the several groups.
-Thus, we aim with this work to evaluate how good ANPR data is for providing trip information and ultimately that of the underlyng road network.
-In the remaining of this section we provide detail on the choice of parameters and filtering methods used in trip identification, as well as in the choice of features.
-Then, we
+However, due to its tiny coverage of the road network, it is not clear how well ANPR data can capture much the travel characteristics necessary to accurately describe the different groups.
 
@@
 Table~/ref{t:features} depicts a sample of the resulting features vector.
-Each  The aim was to choose features that captured different aspects of trip frequency and diversity.
-However, because a high percentage of trips contain a single sighting.
-1034107 vehicles
+A total of 1034107 distinct vehicles were detected.
+However, because a high percentage of trips contains a single sighting, some of these features were highly correlated.
+We therefore, chose to remove 3 of the features represented in~/ref{t:features}: \emph{Average Sightings}, \emph{Average Distinct Routes} and \emph{Average Hour Difference}, to avoid the obfuscation of the natural clustering ~\cite{Kmeans}.
+Furthermore we considered that a trip of length 1 has no destination (which explains values of average distinct destination below zero) and we filtered all instances of vehicles where the total number of trips is lower than 3, resulting in 642006 unique vehicles.
 
-We identified correlated features and removed the following features .. because ..
-
-Filtered all cases where TotalTrips <= 3. (HOW MANY?)
-
-We ran the kmeans algorithm:
- k = 2:8
- iter.max = 200,
- runs = 100
+Clustering of vehicles was performed using the Hartigan and Wong k-means algorithm, for each value of $\Tau$.
+The number of clusters \emph{k} was varied between 2 and 8 and executed with 200 maximum iterations and 100 different starts of the algorithm.
+The Calinski-Harabasz criterion is used to pick the best value of $k$, the one that minimises the whithin-cluster and between-cluster errors and provides the more natural clusters ~/cite{Kmeans}.
+The results of vehicle clustering are presented and discussed in the section below.
 
@@
 \end{tabular}
-\caption{Clusters sizes and mean centers for $\Tau = 7.5$ minutes}
+\caption{Clusters sizes and mean centers for $\Tau = 7.5$ minutes. Averages are daily based.}
 \label{t:kmeans_centers_450}
 \end{table*}
+
+\begin{table*}[t]
+\centering
+\small
+\begin{tabular}{c c c c c c c c c c}
+  \hline
+ Cluster &  Size & \thead{Total\\Trips} & \thead{Average\\Trips} & \thead{Average\\Length} & \thead{Average\\Distinct\\Origins} & \thead{Average\\Distinct\\Destinations} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} & \thead{Average\\Rest\\Time} \\
+  \hline
+  1 & 371318 & 7.19 & 1.76 & 1.68 & 1.68 & 0.71 & 12.30 & 14.55 & 2.14 \\
+    2 & 49852 & 44.85 & 3.73 & 1.90 & 3.16 & 1.73 & 8.98 & 17.02 & 7.76 \\
+    3 & 189224 & 22.82 & 2.29 & 1.84 & 2.11 & 1.06 & 10.00 & 15.76 & 5.60 \\
+    4 & 4593 & 114.34 & 9.12 & 2.30 & 5.92 & 3.98 & 6.93 & 18.21 & 10.43 \\
+   \hline
+\end{tabular}
+\caption{Clusters sizes and mean centers for $\Tau = 20$ minutes. Averages are daily based.}
+\label{t:kmeans_centers_900}
+\end{table*}

### 2018-01-31 paper.tex
Added draft of section 5: Clustering Vehicles

@@
 
-\bibitem{noVehiocles}
-Department for Transport {\em Provisional Road Traffic Estimates Great Britain: October 2016-September 2017}.
-Department for Transport WebSite: https://www.gov.uk/government/statistics/provisional-road-traffic-estimates-great-britain-october-2016-to-september-2017.
-
-\bibitem{distance}
-Department for Transport {\em Vehicle Licensing Statistics: Quarter 3 (Jul-Sep) 2017}.
-Department for Transport WebSite: https://www.gov.uk/government/statistics/vehicle-licensing-statistics-july-to-september-2017.
+\bibitem{ODMobileData}
+Alexander, Lauren, et al. "Origin–-destination trips by purpose and time of day inferred from mobile phone data." Transportation research part c: emerging technologies 58 (2015): 240-250.
 
@@
 
+\bibitem{noVehiocles}
+Department for Transport {\em Provisional Road Traffic Estimates Great Britain: October 2016-September 2017}.
+Department for Transport WebSite: https://www.gov.uk/government/statistics/provisional-road-traffic-estimates-great-britain-october-2016-to-september-2017.
+
+\bibitem{distance}
+Department for Transport {\em Vehicle Licensing Statistics: Quarter 3 (Jul-Sep) 2017}.
+Department for Transport WebSite: https://www.gov.uk/government/statistics/vehicle-licensing-statistics-july-to-september-2017.
+
 \bibitem{EvolutionUTMC2013}
 Hamilton, Andrew, et al. {\em The evolution of urban traffic control: changing policy and technology}. Transportation planning and technology 36.1 (2013): 24-43.
+
+\bibitem{FourStepModel}
+McNally, Michael G. "The four-step model." Handbook of Transport Modelling: 2nd Edition. Emerald Group Publishing Limited, 2007. 35-53.
 
@@
 Transportation Research Part B: Methodological 46.1 (2012): 175-188.
+
+\bibitem{ClusteringGPS}
+Schüssler, Nadine, and Kay W.
+Axhausen.
+{\em Identifying trips and activities and their characteristics from GPS raw data without further information}.
+Arbeitsbericht Verkehrs-und Raumplanung 502 (2008).
 

### 2018-01-31 classification.tex
Added some text about results - continuing..

@@
 \centering
-\small
 \begin{tabular}{c c c c c c c c c c c}
@@
 Furthermore, as mentioned in section~\ref{s.trips}, we used fixed empirical values for $\Tau$, mostly due to time constraints.
-As such, we ran the trip identification sequence for the following values of $\Tau$: $[5 , 7.5, 10, 15, 20, 30]$ minutes.
-The total number of trips identified during the 3 weeks, for each threshold, was approximately [1.36M, 1.24M, 1.17M, 1.08M, 1.03M, 0.97M] respectively.
+As such, we ran the trip identification sequence for the following values of $\Tau$: [5 , 7.5, 10, 15, 20, 30] minutes.
+The corresponding number of identified trips and average trip length is respectively [1.36M, 1.24M, 1.17M, 1.08M, 1.03M, 0.97M] and [1.46, 1.60, 1.69, 1.83, 1.92, 2.04].
 Moreover, we did not set a value for $\tau$, but instead handled unplausible trips by filtering all sightings with confidence below 85\%.
@@
 
-Table~/ref{t:features} depicts a sample of the resulting features vector.
+Table~\ref{t:features} depicts a sample of the resulting features vector.
 A total of 1034107 distinct vehicles were detected.
 However, because a high percentage of trips contains a single sighting, some of these features were highly correlated.
-We therefore, chose to remove 3 of the features represented in~/ref{t:features}: \emph{Average Sightings}, \emph{Average Distinct Routes} and \emph{Average Hour Difference}, to avoid the obfuscation of the natural clustering ~\cite{Kmeans}.
+We therefore, chose to remove 3 of the features represented in~\ref{t:features}: \emph{Average Sightings}, \emph{Average Distinct Routes} and \emph{Average Hour Difference}, to avoid the obfuscation of the natural clustering ~\cite{Kmeans}.
 Furthermore we considered that a trip of length 1 has no destination (which explains values of average distinct destination below zero) and we filtered all instances of vehicles where the total number of trips is lower than 3, resulting in 642006 unique vehicles.
 
-Clustering of vehicles was performed using the Hartigan and Wong k-means algorithm, for each value of $\Tau$.
+Clustering of vehicles was performed using the Hartigan and Wong \emph{k-means} algorithm, for each value of $\Tau$.
 The number of clusters \emph{k} was varied between 2 and 8 and executed with 200 maximum iterations and 100 different starts of the algorithm.
-The Calinski-Harabasz criterion is used to pick the best value of $k$, the one that minimises the whithin-cluster and between-cluster errors and provides the more natural clusters ~/cite{Kmeans}.
+The Calinski-Harabasz criterion is used to pick the best value of $k$, the one that minimises the whithin-cluster and between-cluster errors and provides the more natural clusters ~\cite{Kmeans}.
 The results of vehicle clustering are presented and discussed in the section below.
@@
 
+Table~\ref{t:tau_comparison} provides a summary of multiple runs of \emph{k-means} for each value of $\Tau$.
+The optimal number of clusters, selected by maximising the Calinski-Harabasz criterion, is presented and is shown to decrease inversely with $\Tau$.
+This in part expected because as $\Tau$ increases, so does the average trip length, as opposed to the total number of trips, and trips per vehicle.
+The interp   ...
+The relative between and average whithin clusters sum of squares is also shown.
+
 \begin{table}[ht]
 \centering
+\tabcolsep=0.17cm
 \begin{tabular}{c c c c c}
   \hline
-\thead{Tau\\(min)} & Best k & Betweenss & Whithinss & \thead{Calinski-\\Harabasz} \\
+Tau & Best $k$ & Betweenss & \thead{Average\\Whithinss} & \thead{Calinski-\\Harabasz} \\
   \hline
-5 &   8 & 0.92 & 0.09 & 1,116,962 \\
-  7.5 &   7 & 0.90 & 0.10 & 1,003,744 \\
-  10 &   7 & 0.90 & 0.18 &   911,044 \\
-  15 &   6 & 0.86 & 0.19 &   794,557 \\
-  20 &   4 & 0.79 & 0.19 &   754,241 \\
-  30 &   3 & 0.71 & 0.30 &   743,001 \\
+5 min &   8 & 0.923 & 0.125 & 1,116,962 \\
+  7.5 min &   7 & 0.904 & 0.143 & 1,003,744 \\
+  10 min&   7 & 0.896 & 0.143 &   911,044 \\
+  15 min &   6 & 0.865 & 0.167 &   794,557 \\
+  20 min &   4 & 0.786 & 0.250 &   754,241 \\
+  30 min &   3 & 0.710 & 0.333 &   743,001 \\
    \hline
 \end{tabular}
-\caption{Kmeans performance comparison for each best k }
+\caption{\emph{k-means} performance for several values of $\Tau$.}
 \label{t:tau_comparison}
 \end{table}
+
+Tables~\ref{t:kmeans_centers_450} and~\ref{t:kmeans_centers_1200} depict the cluster centers for $\Tau$ equal to 7.5 minutes and 20 minutes respectively.
+The values for the cluster centers partially meet our expectations.
+We were expecting to find a relatively small cluster representing taxis with a high average number of trips per day, occurring over a variety of origins and destinations and over a large time frame.
+Clusters 2 and 7 for $\Tau = 7.5$ and cluster 4 for $\Tau = 20$ do indeed fit this profile.
+What differentiates cluster 2 from 7 in the first case is essentially the mean number of trips per day and the average time at which the first and last trips of the day occur.
+We could expect that buses could be separated from taxis by showing less diversity in the number of origins and destinations as these essentially do multiple runs of the same trip throughout the day.
+Still, this can be explained by the fact that buses take routes through main and secundary roads.
+As most cameras are placed in main roads, the one long bus trip can be perceived as multiple small trips as the bus alternates between arterial and main roads.
+
+On the other hand, we could expect that commuters display average and relatively few trips per day.
 
@@
 \end{tabular}
-\caption{Clusters sizes and mean centers for $\Tau = 7.5$ minutes. Averages are daily based.}
+\caption{Clusters sizes and mean centers for $\Tau = 7.5$ minutes.}
 \label{t:kmeans_centers_450}
@@
 \end{tabular}
-\caption{Clusters sizes and mean centers for $\Tau = 20$ minutes. Averages are daily based.}
-\label{t:kmeans_centers_900}
+\caption{Clusters sizes and mean centers for $\Tau = 20$ minutes.}
+\label{t:kmeans_centers_1200}
 \end{table*}
+
+The differences between expected and observed results are in great part due to the fact that a high percentage of trips contains a single sighting.
+..
+It may be the case that commuters choose routes that do not pass through ANPR cameras.
+Devising further models and methods that model and capture this uncertainty is fundamental to getting the most out of ANPR data.

### 2018-01-31 classification.tex
Added conclusion and future work

@@
 What differentiates cluster 2 from 7 in the first case is essentially the mean number of trips per day and the average time at which the first and last trips of the day occur.
-We could expect that buses could be separated from taxis by showing less diversity in the number of origins and destinations as these essentially do multiple runs of the same trip throughout the day.
+We expected that buses could be separated from taxis by showing less diversity in the number of origins and destinations as these essentially do multiple runs of the same trip throughout the day.
 Still, this can be explained by the fact that buses take routes through main and secundary roads.
@@
 
-On the other hand, we could expect that commuters display average and relatively few trips per day.
+On the other hand, we could expect that commuters display trips starting at around 8h in the morning and terminating at 17h-18h.
+Additionally, we would expect to see few trips per day as most of the day is spent in work, ocasionally .
 
@@
 \end{tabular}
-\caption{Clusters sizes and mean centers for $\Tau = 7.5$ minutes.}
+\caption{Clusters sizes and centers for $\Tau = 7.5$ minutes.}
 \label{t:kmeans_centers_450}
@@
 \end{tabular}
-\caption{Clusters sizes and mean centers for $\Tau = 20$ minutes.}
+\caption{Clusters sizes and  centers for $\Tau = 20$ minutes.}
 \label{t:kmeans_centers_1200}

### 2018-01-31 discussion.tex
Renamed section
Added conclusion and future work

@@
 % !TEX root = paper.tex
-\section{Discussion}\label{s.conclusions}
+\section{Conclusion and Future Work}\label{s.conclusions}
+
+Most urban cities in the world possess a network of ANPR cameras that is used in law enforcement and for traffic monitoring and control.
+Number plate data collected in the Tyne and Wear area is stored and leveraged by the UTMC for computing average journey times across a selection of sensitive roads.
+However, number plate data could be used more extensively to identify and study individual and collective travel patterns.
+In this paper, we've presented a set of definitions and constraints that establish a conceptual foundation for identifying vehicle trips from number plate detections.
+We've identified two parameters, $\tau$ and $\Tau$ as critical in the discrimination of plausible and unplausible trips.
+Hence, future work should first and foremost focus on developing formal methods to estimate these parameters from observed distributions of travel times and by applying knowledge about the structure of the road network.
+Moreover, methods for addressing issues concerning camera performance, namely wrong and duplicate scans, should be further developed and researched.
+
+Once trip data has been computed, a range of interesting applications is available.
+This work tries to identify groups of vehicles by clustering information about frequency and diversity of travel.
+By associating a vehicle with a cluster that represents taxis, or home-to-work commuters, one can begin estimating trip mode usage across the city.
+However, the results presented here could benefit from extra work and further validation.
+Finally, future work can focus on using trip data to solve interesting research problems such as:
+\begin{enumerate*}[label=(\roman*)]
+  \item real-time route recommendation using probabilistic graphical models;
+  \item detection of abnormal trip patterns for helping law enforcement in the identification of suspect vehicles or behavior;
+  \item modelling how drivers make routing choices in the presence of anomalies in the road network.
+\end{enumerate*}

### 2018-01-31 classification.tex
Fixed some mistakes, added table in vehicle clustering section

@@
 \section{Clustering vehicles}\label{s.classification}
+
+One direct application of trip data is unsupervised learning.
+Previous works have focused on identifying travel modes or purpose of trip from different sources of data, most notably: survey data, GPS data and mobile phone data~\cite{ODMobileData, ClusteringGPS}.
+Determining the distribution of travel modes is one of the fundamental steps in the four-stage model: an essential traffic modelling methodology for transportation planning~\cite{FourStepModel}.
+Although number plate data has been used in~\cite{Clustering}, to identify different categories of trips, these do not differentiate between private or public travel modes and focus for the most part on when the trips occurred.
+Therefore, we aim to identify groups of vehicles with similar trip patterns based on frequency and diversity of travel.
+This can be private vehicles, such as work-home commuters, transit vehicles like buses and taxis, or other types of vehicles such as delivery trucks.
+%This in turn, may enable trip modes to be estimated
+However, due to its limited coverage of the road network, it is not clear how well ANPR data can capture much the travel characteristics necessary to accurately describe the different groups.
+
+Due to distinct traffic behavior during weekends, we chose to consider only trips occurring during weekdays.
+We thus used all number plate data collected during between the 6th and 24th of February and excluded data from the 2 weekends in between.
+Furthermore, as mentioned in section~\ref{s.trips}, we used fixed empirical values for $\Tau$, mostly due to time constraints.
+Table~\ref{t:trips-tau} displays the number of trips, average trip length and the proportion of trips of length one, for varying values of $\Tau$.
+The effect of incrementing $\Tau$ on resulting trips is clear: increased trip lengths and less trips containing just a single sighting.
+On the other hand, we did not set a value for $\tau$, but instead handled unplausible trips by filtering all sightings with confidence below 85\%.
+Duplicates were filtered after identifying consecutive sightings by the same camera occurring within the same trip.
+Clock synchronisation errors were provided in milliseconds with none exceeding 5 seconds.
+These were therefore ignored.
+
+\begin{table}[t]
+\centering
+\tabcolsep=0.25cm
+\begin{tabular}{c c c c}
+  \hline
+Tau & Trips & \thead{Average\\Length} & \thead{Proportion of Trips \\ Length 1} \\
+  \hline
+5 min & 13,603,759 & 1.46 & 0.70 \\
+7.5 min & 12,394,709 & 1.60 & 0.64 \\
+10 min & 11,690,791 & 1.69 & 0.61 \\
+15 min & 10,823,333 & 1.83 & 0.57 \\
+20 min & 10,305,860 & 1.92 & 0.54 \\
+30 min &  9,653,499 & 2.04 & 0.52 \\
+   \hline
+\end{tabular}
+\caption{Overview of trip data for varying values of $\Tau$.}
+\label{t:trips-tau}
+\end{table}
+
+Transforming trips into features that could be used in clustering algorithms was a 3-step process:
+\begin{enumerate}
+  \item First, every trip was summarised as a single row of data. The following information was extracted: length of trip, origin, destination, route, start and end times.
+  \item Second, daily trip information was obtained for each vehicle: number of trips, median of trip lengths, number of sightings, distinct number of origins destinations, and routes, hour of first sighting, hour of last sighting and total rest time between trips.
+  \item Finaly, daily information per vehicle was collapsed into a single row by averaging this information across the 15 days.
+\end{enumerate}
+
+Table~\ref{t:features} depicts a sample of the resulting features vector.
+A total of 1034107 distinct vehicles were detected.
+However, because a high percentage of trips contains a single sighting, some of these features were highly correlated.
+We therefore, chose to remove 3 of the features represented in~\ref{t:features}: \emph{Average Sightings}, \emph{Average Distinct Routes} and \emph{Average Hour Difference}, to avoid the obfuscation of the natural clustering ~\cite{Kmeans}.
+Furthermore we considered that a trip of length 1 has no destination (which explains values of average distinct destination below zero) and we filtered all instances of vehicles where the total number of trips is lower than 3, resulting in 642006 unique vehicles.
 
@@
 
-One direct application of trip data is unsupervised learning.
-Previous works have focused on identifying travel modes or purpose of trip from different sources of data, most notably: survey data, GPS data and mobile phone data~\cite{ODMobileData, ClusteringGPS}.
-Determining the distribution of travel modes is one of the fundamental steps in the four-stage model: an essential traffic modelling methodology for transportation planning~\cite{FourStepModel}.
-Although number plate data has been used in~\cite{Clustering}, to identify different categories of trips, these do not differentiate between private or public travel modes and focus for the most part on when the trips occurred.
-Therefore, we aim to identify groups of vehicles with similar trip patterns based on frequency and diversity of travel.
-This can be private vehicles, such as work-home commuters, transit vehicles like buses and taxis, or other types of vehicles such as delivery trucks.
-%This in turn, may enable trip modes to be estimated
-However, due to its tiny coverage of the road network, it is not clear how well ANPR data can capture much the travel characteristics necessary to accurately describe the different groups.
-
-Due to distinct traffic behavior during weekends, we chose to consider only trips occurring during weekdays.
-We thus used all number plate data collected during between the 6th and 24th of February and excluded data from the 2 weekends in between.
-Furthermore, as mentioned in section~\ref{s.trips}, we used fixed empirical values for $\Tau$, mostly due to time constraints.
-As such, we ran the trip identification sequence for the following values of $\Tau$: [5 , 7.5, 10, 15, 20, 30] minutes.
-The corresponding number of identified trips and average trip length is respectively [1.36M, 1.24M, 1.17M, 1.08M, 1.03M, 0.97M] and [1.46, 1.60, 1.69, 1.83, 1.92, 2.04].
-Moreover, we did not set a value for $\tau$, but instead handled unplausible trips by filtering all sightings with confidence below 85\%.
-Duplicates were filtered after identifying consecutive sightings by the same camera occurring within the same trip.
-Clock synchronisation errors were provided in milliseconds with none exceeding 5 seconds.
-These were therefore ignored.
-
-Transforming trips into features that could be used in clustering algorithms was a 3-step process:
-\begin{enumerate}
-  \item First, every trip was summarised as a single row of data. The following information was extracted: length of trip, origin, destination, route, start and end times.
-  \item Second, daily trip information was obtained for each vehicle: number of trips, median of trip lengths, number of sightings, distinct number of origins destinations, and routes, hour of first sighting, hour of last sighting and total rest time between trips.
-  \item Finaly, daily information per vehicle was collapsed into a single row by averaging this information across the 15 days.
-\end{enumerate}
-
-Table~\ref{t:features} depicts a sample of the resulting features vector.
-A total of 1034107 distinct vehicles were detected.
-However, because a high percentage of trips contains a single sighting, some of these features were highly correlated.
-We therefore, chose to remove 3 of the features represented in~\ref{t:features}: \emph{Average Sightings}, \emph{Average Distinct Routes} and \emph{Average Hour Difference}, to avoid the obfuscation of the natural clustering ~\cite{Kmeans}.
-Furthermore we considered that a trip of length 1 has no destination (which explains values of average distinct destination below zero) and we filtered all instances of vehicles where the total number of trips is lower than 3, resulting in 642006 unique vehicles.
-
 Clustering of vehicles was performed using the Hartigan and Wong \emph{k-means} algorithm, for each value of $\Tau$.
 The number of clusters \emph{k} was varied between 2 and 8 and executed with 200 maximum iterations and 100 different starts of the algorithm.
-The Calinski-Harabasz criterion is used to pick the best value of $k$, the one that minimises the whithin-cluster and between-cluster errors and provides the more natural clusters ~\cite{Kmeans}.
+The Calinski-Harabasz criterion is used to pick the best value of $k$, the one that minimises the within-cluster and between-cluster errors and provides the more natural clusters ~\cite{Kmeans}.
 The results of vehicle clustering are presented and discussed in the section below.
@@
 
-\begin{table}[ht]
+Tables~\ref{t:kmeans_centers_450} and~\ref{t:kmeans_centers_1200} depict the cluster centres for $\Tau$ equal to 7.5 minutes and 20 minutes respectively.
+The values for the cluster centers partially meet our expectations.
+We were expecting to find a relatively small cluster representing taxis with a high average number of trips per day, occurring over a variety of origins and destinations and over a large time frame.
+Clusters 2 and 7 for $\Tau = 7.5$ and cluster 4 for $\Tau = 20$ do indeed fit this profile.
+What differentiates cluster 2 from 7 in the first case is essentially the mean number of trips per day and the average time at which the first and last trips of the day occur.
+We expected that buses could be separated from taxis by showing less diversity in the number of origins and destinations as these essentially do multiple runs of the same trip throughout the day.
+Still, this can be explained by the fact that buses take routes through main and secondary roads.
+As most cameras are placed in main roads, the one long bus trip can be perceived as multiple small trips as the bus alternates between arterial and main roads.
+
+On the other hand, we could expect that commuters display trips starting at around 8h in the morning and terminating at 17h-18h.
+Additionally, we would expect to see few trips per day as most of the day is spent in work, ocasionally .
+
+\begin{table}[t]
 \centering
@@
   \hline
-Tau & Best $k$ & Betweenss & \thead{Average\\Whithinss} & \thead{Calinski-\\Harabasz} \\
+Tau & Best $k$ & Betweenss & \thead{Average\\Withinss} & \thead{Calinski-\\Harabasz} \\
   \hline
@@
 \end{table}
-
-Tables~\ref{t:kmeans_centers_450} and~\ref{t:kmeans_centers_1200} depict the cluster centers for $\Tau$ equal to 7.5 minutes and 20 minutes respectively.
-The values for the cluster centers partially meet our expectations.
-We were expecting to find a relatively small cluster representing taxis with a high average number of trips per day, occurring over a variety of origins and destinations and over a large time frame.
-Clusters 2 and 7 for $\Tau = 7.5$ and cluster 4 for $\Tau = 20$ do indeed fit this profile.
-What differentiates cluster 2 from 7 in the first case is essentially the mean number of trips per day and the average time at which the first and last trips of the day occur.
-We expected that buses could be separated from taxis by showing less diversity in the number of origins and destinations as these essentially do multiple runs of the same trip throughout the day.
-Still, this can be explained by the fact that buses take routes through main and secundary roads.
-As most cameras are placed in main roads, the one long bus trip can be perceived as multiple small trips as the bus alternates between arterial and main roads.
-
-On the other hand, we could expect that commuters display trips starting at around 8h in the morning and terminating at 17h-18h.
-Additionally, we would expect to see few trips per day as most of the day is spent in work, ocasionally .
 
@@
 \end{tabular}
-\caption{Clusters sizes and centers for $\Tau = 7.5$ minutes.}
+\caption{Clusters sizes and centres for $\Tau = 7.5$ minutes.}
 \label{t:kmeans_centers_450}
@@
 \end{tabular}
-\caption{Clusters sizes and  centers for $\Tau = 20$ minutes.}
+\caption{Clusters sizes and centres for $\Tau = 20$ minutes.}
 \label{t:kmeans_centers_1200}

### 2018-01-31 intro.tex
Fixed some mistakes, added table in vehicle clustering section

@@
 For example if two sightings are made from ANPR cameras which are connectable by a ``sensible'' route\footnote{Here ``sensible'' implies that a route between cameras A and B would not need to go through a third camera C.} in a time interval which is ``sensible'' then these can be determined to be part of the same journey.
-However, if the timings between two sightings is significantly longer than what would be expected then this would imply that the vehicle stopped between these two cameras and that the latte sighting is part of a new journey.
+However, if the timings between two sightings is significantly longer than what would be expected then this would imply that the vehicle stopped between these two cameras and that the later sighting is part of a new journey.
 

### 2018-02-01 trips.tex
Fixed maths formula in duplicate scannings subsection
Finished text on errors scans

@@
 
-\begin{equation} \label{e.sighting.different}
-t^{k}_{i} \ne t^{k}_{j} \Rightarrow s^{k}_{i} \ne s^{k}_{j} \ , \ i \ne j
-\end{equation}
+\begin{align} \label{e.sighting.different.2}
+ (\ c^{k}_{i} &\ne c^{k}_{j} \, )\ \vee \\
+ (\ c^{k}_{i} &= c^{k}_{j} \wedge |t^{k}_{i} - t^{k}_{j}| < \gamma \, )\ \Rightarrow s^{k}_{i} \ne s^{k}_{j}  \ , \ i \ne j \nonumber
+\end{align}
 
@@
 
-Even though number plate recognition cameras have accuracy rates of 99.9\% or higher, on average 1 observation out of 1000 is a misclassified plate number.
+Number plate recognition cameras have accuracy rates of 99.9\% or higher.
+If we consider that on average 1 to 10 out of 10000 scans are misclassified plate number then approximately 200 to 2000 scans everyday are incorrect scans.
 We categorise the errors made by ANPR cameras into two types:
@@
 On the second case, there is a recorded sighting, but this will be assigned to the incorrect vehicle.
+Even though, the camera provides a value of \emph{confidence} which helps in diagnosing incorrect scans, errors still occur for high values of confidence.
+Even though that in this work we do not present a complex solution for this issue, methods to detect and address these two types of errors need to be designed and implemented in the future.

### 2018-02-01 paper.tex
A few fixes and title change

@@
 
-\title{\Large Clustering Vehicle Trips Identified from Automatic \\ Number Plate Recognition Camera Scans}
+\title{\Large Clustering Trips Identified from Automatic Number \\ Plate Recognition Camera Scans}
 \author{Pedro M. Pinto Silva \thanks{School of Computing, Newcastle University, United Kingdom}
@@
 
-\begin{abstract} \small\baselineskip=9pt This is the text of my abstract. It is a brief
-description of my
-paper, outlining the purposes and goals I am trying to address.\end{abstract}
+\begin{abstract} \small\baselineskip=9pt Increases in price and performance of automatic number plate recognition over the last two decades has led to  by law enforcement and urban traffic control.
+In this work we show how \end{abstract}
 

### 2018-02-01 dataset.tex
Removed vspace from table
A few fixes and title change
Edits to sections 2 and 3

@@
 \includegraphics[width=.95\linewidth]{ANPR-overview.pdf}
-\caption{}
+\caption{Automatic number plate recognition system overview.}
 \label{fig:anpr-overview}
@@
 
-Automatic number plate recognition (ANPR) cameras are actively employed in urban traffic environments and play an important role in day-to-day intelligent transportation systems.
-They can be used by government subsidised entities in urban traffic management and control; by comissioned highway agencies in electronic toll collection; or by law enforcement organisations in detecting speeding vehicles and validating number plate registrations.
+Automatic number-plate recognition (ANPR) cameras are actively employed in urban traffic environments and play an important role in day-to-day intelligent transportation systems.
+They can be used by government subsidised entities in urban traffic management and control; by commissioned highway agencies in electronic toll collection; or by law enforcement organisations in detecting speeding vehicles and validating number plate registrations.
 The wide diversity of applications, paired with the large improvements in price-to-performance ratios of ANPR hardware and software systems, has resulted in increased investments of ANPR cameras for urban environments~\cite{EvolutionUTMC2013, SurveyITS2011}.
@@
   \includegraphics[width=.9\linewidth]{observations_per_day.pdf}
-  \caption{Total number of scans recorded per day in Tyne and Wear. There is a clear seasonal effect caused by decreasing traffic demands in weekends and increasing traffic volume in weekdays.}
+  \caption{Total number of scans recorded per day in Tyne and Wear. There is a clear seasonal effect caused by decreasing traffic demands at weekends and increasing traffic volume during weekdays.}
   \label{fig:observations-per-day}
@@
 Inter-camera variability is observed, as some cameras are located in more traffic intensive road sections than others.
-Decomissioned or temporarily unavailable cameras (due to loss of power, faulty camera, road closed, etc) are depicted at the bottom of the graph.}
+Decommissioned or temporarily unavailable cameras (due to loss of power, faulty camera, road closed, etc) are depicted at the bottom of the graph.}
   \label{fig:observations-per-camera-day}
@@
 Figure~\ref{fig:time-series} shows the number of daily scans recorded over a month (February, 2017).
-Furthermore, every scan is stored in a central database managed by the Urban Traffic Management and Control (UTMC) of Tyne and Wear, and used to compute travel times across particular links of interest in the road network.
+Furthermore, every scan is stored in a central database managed by Urban Traffic Management and Control (UTMC) Tyne and Wear, and used to compute travel times across particular links of interest in the road network.
 These are usually major roads that see high volumes of traffic, or road segments more prone to traffic jams.
@@
 An excerpt of the data can be found in table~\ref{t:np_data_example}.
-All plate numbers were anonymised by the UTMC through a hashing algorithm before the data was shared.
+All number-plates were anonymised by the UTMC through a hashing algorithm before the data was shared.
 Cameras are uniquely identified by an integer and timestamps are relative to each camera's clock.
@@
 \end{tabular}
-\caption{Sample of number plate data. Clock error is given in milliseconds and confidence as a percentual value.}
+\caption{Sample of number plate data. Clock error is given in milliseconds and confidence as a percentile value.}
 \label{t:np_data_example}
-\vspace{-0.5cm}
 \end{table}
 
-The main use of ANPR data for the UTMC is average journey time estimation for selected links in the road network.
-Furthermore, several authors have extensively researched how to use used number plate data as an extension on link counts for estimating origin-destination matrices and link flows~\cite{Castillo2010, Castillo2008, Hazelton2012}.
-However, very few works have focused on analysing individual or collective travel patterns from number plate data, particularly across extended periods of time.
+{\color{red}{\color{blue}Much of this paragraph should be in related work}
+The main use of ANPR data for the UTMC is estimating average journey times for selected links in the road network.
+Furthermore, several authors have extensively researched how to use number-plate data as an extension to link counts for estimating origin-destination matrices and link flows~\cite{Castillo2010, Castillo2008, Hazelton2012}.
+However, very few works have focused on analysing individual or collective travel patterns from number-plate data, particularly across extended periods of time.
 Moreover, there is no consistent conceptual and analytical framework for transforming number plate data into a historical sequence of trips for each vehicle.
-Finally, we believe that trip data, properly identified from number plate data, has the potential to unlock a number of new applications for urban traffic control and law enforcement.
+Finally, we believe that trip data, properly identified from number-plate data, has the potential to unlock a number of new applications for urban traffic control and law enforcement.
 Thus, in the following section we present a conceptual methodology for grouping multiple camera observations of the same vehicle into one or several trips of that vehicle.
+}

### 2018-02-01 intro.tex
Edits to sections 2 and 3

@@
 
-In ~\cite{Castillo2008}, the authors used partial trip information from plate number scans to compute travel demands, in the form of a trip matrix, and estimate link flows from a set of constraints composed of traffic conservation laws, prior knowledge and path flow disaggregation techniques, such as Stochastic User Equilibrium (SUE).
+Castillo et al~\cite{Castillo2008} used partial trip information from number-plate scans to compute travel demands, constructing this in the form of a trip matrix.
+They estimated link flows from a set of constraints composed of traffic conservation laws, prior knowledge and path flow disaggregation techniques, such as Stochastic User Equilibrium (SUE).
 
@@
 
-Uses of trip data for understandint travel demands / traffic prediction
+{\color{red}Uses of trip data for understanding travel demands / traffic prediction
 
@@
 
-Number plate data is not as popular, for urban traffic monitoring and prediction , as other sources of data, namely GPS traces or floating vehicle data.
+Number plate data is not as popular, for urban traffic monitoring and prediction, as other sources of data, namely GPS traces or floating vehicle data.
 Previous works trips from GPS traces, .
@@
 % \end{itemize}
+}

### 2018-02-01 paper.tex
Fixed some grammar mistakes and added abstract

@@
 
-\begin{abstract} \small\baselineskip=9pt Increases in price and performance of automatic number plate recognition over the last two decades has led to  by law enforcement and urban traffic control.
-In this work we show how \end{abstract}
+\begin{abstract} \small\baselineskip=9pt In major cities, government agencies increasingly employ automatic number plate recognition (ANPR) technology in law enforcement and traffic control.
+In the region of Tyne and Wear (UK) the network of ANPR cameras is used to monitor travel times across sensitive roads.
+However, few works have explored the full potential of number plate scans for analysing individual and collective travel patterns.
+In this work we present a methodology for deriving trips from vehicle sightings at fixed camera locations.
+We illustrate the effect of parameters $\tau$ and $\Tau$ on trip discrimination and on the detection of implausible trips.
+To demonstrate the potential of trip data we apply \emph{k-means} clustering to trips identified from over 40 million plate scans recorded over 15 weekdays.
+Results show that whilst private and transit travel modes can begin to be inferred from the resulting clusters, further work needs to be put into developing a more consistent and integrated framework for trip identification in ANPR data.\end{abstract}
 

### 2018-02-01 classification.tex
Fixed some grammar mistakes and added abstract
Gone through classification

@@
 Determining the distribution of travel modes is one of the fundamental steps in the four-stage model: an essential traffic modelling methodology for transportation planning~\cite{FourStepModel}.
-Although number plate data has been used in~\cite{Clustering}, to identify different categories of trips, these do not differentiate between private or public travel modes and focus for the most part on when the trips occurred.
+Although number-plate data has been used in~\cite{Clustering}, to identify different categories of trips, these do not differentiate between private or public travel modes and focus for the most part on when the trips occurred.
 Therefore, we aim to identify groups of vehicles with similar trip patterns based on frequency and diversity of travel.
@@
 
-Due to distinct traffic behavior during weekends, we chose to consider only trips occurring during weekdays.
-We thus used all number plate data collected during between the 6th and 24th of February and excluded data from the 2 weekends in between.
-Furthermore, as mentioned in section~\ref{s.trips}, we used fixed empirical values for $\Tau$, mostly due to time constraints.
+Due to distinctly different traffic behaviour during weekends, we chose to consider only trips occurring during weekdays.
+We thus used all number-plate data collected between the 6th and 24th of February and exclude data from the 2 weekends in-between.
+Furthermore, as mentioned in section~\ref{s.trips}, we used fixed empirical values for $\Tau$.
 Table~\ref{t:trips-tau} displays the number of trips, average trip length and the proportion of trips of length one, for varying values of $\Tau$.
-The effect of incrementing $\Tau$ on resulting trips is clear: increased trip lengths and less trips containing just a single sighting.
-On the other hand, we did not set a value for $\tau$, but instead handled unplausible trips by filtering all sightings with confidence below 85\%.
+The effect of incrementing $\Tau$ on resulting trips is clear: increased trip lengths and decreased trips containing just a single sighting.
+On the other hand, we did not set a value for $\tau$, but instead handled implausible trips by filtering all sightings with confidence below 85\%.
 Duplicates were filtered after identifying consecutive sightings by the same camera occurring within the same trip.
@@
 
-Transforming trips into features that could be used in clustering algorithms was a 3-step process:
+Transforming trips into features which can be used in clustering algorithms was a 3-step process:
 \begin{enumerate}
@@
   \item Second, daily trip information was obtained for each vehicle: number of trips, median of trip lengths, number of sightings, distinct number of origins destinations, and routes, hour of first sighting, hour of last sighting and total rest time between trips.
-  \item Finaly, daily information per vehicle was collapsed into a single row by averaging this information across the 15 days.
+  \item Finally, daily information per vehicle was collapsed into a single row by averaging this information across the 15 days.
 \end{enumerate}
@@
 A total of 1034107 distinct vehicles were detected.
-However, because a high percentage of trips contains a single sighting, some of these features were highly correlated.
-We therefore, chose to remove 3 of the features represented in~\ref{t:features}: \emph{Average Sightings}, \emph{Average Distinct Routes} and \emph{Average Hour Difference}, to avoid the obfuscation of the natural clustering ~\cite{Kmeans}.
-Furthermore we considered that a trip of length 1 has no destination (which explains values of average distinct destination below zero) and we filtered all instances of vehicles where the total number of trips is lower than 3, resulting in 642006 unique vehicles.
+However, because a high percentage of trips containing a single sighting, some of these features were highly correlated.
+We therefore, chose to remove three of the features represented in~\ref{t:features}: \emph{Average Sightings}, \emph{Average Distinct Routes} and \emph{Average Hour Difference}, to avoid the obfuscation of the natural clustering ~\cite{Kmeans}.
+Furthermore we considered that a trip of length one has no destination (which explains values of average distinct destination below zero) and we filtered all instances of vehicles where the total number of trips is lower than 3, resulting in 642006 unique vehicles.
 
@@
 Clustering of vehicles was performed using the Hartigan and Wong \emph{k-means} algorithm, for each value of $\Tau$.
-The number of clusters \emph{k} was varied between 2 and 8 and executed with 200 maximum iterations and 100 different starts of the algorithm.
+The number of clusters \emph{k} was varied between 2 and 8 and executed with a maximum of 200 iterations and 100 different starting states of the algorithm.
 The Calinski-Harabasz criterion is used to pick the best value of $k$, the one that minimises the within-cluster and between-cluster errors and provides the more natural clusters ~\cite{Kmeans}.
@@
 This in part expected because as $\Tau$ increases, so does the average trip length, as opposed to the total number of trips, and trips per vehicle.
-The interp   ...
-The relative between and average whithin clusters sum of squares is also shown.
+{\color{blue}The interp   ...} The relative between and average within clusters sum of squares is also shown.
 
 Tables~\ref{t:kmeans_centers_450} and~\ref{t:kmeans_centers_1200} depict the cluster centres for $\Tau$ equal to 7.5 minutes and 20 minutes respectively.
-The values for the cluster centers partially meet our expectations.
+The values for the cluster centres partially meet our expectations.
 We were expecting to find a relatively small cluster representing taxis with a high average number of trips per day, occurring over a variety of origins and destinations and over a large time frame.
@@
 
-On the other hand, we could expect that commuters display trips starting at around 8h in the morning and terminating at 17h-18h.
-Additionally, we would expect to see few trips per day as most of the day is spent in work, ocasionally .
+On the other hand, we could expect that commuters display trips starting at around eight in the morning and a second trip terminating between five and six in the evening.
+Additionally, we would expect to see few trips per day as most of the day is spent in work.
 
@@
 The differences between expected and observed results are in great part due to the fact that a high percentage of trips contains a single sighting.
-..
 It may be the case that commuters choose routes that do not pass through ANPR cameras.

### 2018-02-01 classification.tex
Used bits from other sections as the related work. Had to move figure 1 around so that the structure of the paper would not be affected as a result.

--- 2018-01-31 classification.tex
@@
 One direct application of trip data is unsupervised learning.
-Previous works have focused on identifying travel modes or purpose of trip from different sources of data, most notably: survey data, GPS data and mobile phone data~\cite{ODMobileData, ClusteringGPS}.
-Determining the distribution of travel modes is one of the fundamental steps in the four-stage model: an essential traffic modelling methodology for transportation planning~\cite{FourStepModel}.
-Although number plate data has been used in~\cite{Clustering}, to identify different categories of trips, these do not differentiate between private or public travel modes and focus for the most part on when the trips occurred.
-Therefore, we aim to identify groups of vehicles with similar trip patterns based on frequency and diversity of travel.
-This can be private vehicles, such as work-home commuters, transit vehicles like buses and taxis, or other types of vehicles such as delivery trucks.
+In this section we identify groups of vehicles with similar trip patterns based on frequency and diversity of travel.
+Clusters can represent private vehicles, such as work-home commuters, transit vehicles like buses and taxis, or other types of vehicles such as delivery trucks.
 %This in turn, may enable trip modes to be estimated
-However, due to its limited coverage of the road network, it is not clear how well ANPR data can capture much the travel characteristics necessary to accurately describe the different groups.
+%However, due to its limited coverage of the road network, it is not clear how well ANPR data can capture much the travel characteristics necessary to accurately describe the different groups.
 
@@
 The effect of incrementing $\Tau$ on resulting trips is clear: increased trip lengths and less trips containing just a single sighting.
-On the other hand, we did not set a value for $\tau$, but instead handled unplausible trips by filtering all sightings with confidence below 85\%.
+On the other hand, we did not set a value for $\tau$, but instead handled implausible trips by filtering all sightings with confidence below 85\%.
 Duplicates were filtered after identifying consecutive sightings by the same camera occurring within the same trip.

### 2018-02-01 trips.tex
A few fixes and title change

@@
 \section{Trip Identification}\label{s.trips}
+
+In this section we explain our technique for identifying which sets of sightings can be combined in order to make a trip.
 
@@
 \begin{equation} \label{e.sighting}
-s^{k}_{i} = \{ c, t \}
+s^{k}_{i} = \{ c, t \},
 \end{equation}
 
-where \emph{c} is an integer that uniquely identifies a camera, and \emph{t} is a scalar representing a point in time (e.g.\ a timestamp).
+where \emph{c} uniquely identifies a camera, and \emph{t} is a scalar representing a point in time (e.g.\ a timestamp).
 
-Let an ordered sequence of sightings of vehicle \emph{k} define the $u_{th}$ trip of \emph{k}:
+Let an ordered sequence of sightings of vehicle \emph{k} define the $u_{th}$ trip of vehicle \emph{k}:
 
 \begin{equation} \label{e.trip}
-w^{k}_{u} = \left(s^{k}_{(1)}, s^{k}_{(2)}, \dots , s^{k}_{(n)}\right)
+w^{k}_{u} = \left(s^{k}_{(1)}, s^{k}_{(2)}, \dots , s^{k}_{(n)}\right),
 \end{equation}
@@
 \begin{equation} \label{e.journeytime}
-jt^{k}_{u} = \left(t^{k}_{(2)} - t^{k}_{(1)}, \ldots, t^{k}_{(n)} - t^{k}_{(n-1)} \right)
+jt^{k}_{u} = \left(t^{k}_{(1)} - t^{k}_{(2)}, \ldots, t^{k}_{(n)} - t^{k}_{(n-1)} \right).
 \end{equation}
 
-We consider a trip of \emph{k} valid if the following conditions are met:
+We consider a trip of \emph{$w^k_u$} valid if the following conditions hold:
 
@@
 n &\ge 1 , \label{e.trip.constraints.1} \\
-\tau_{(i)} &< jt^{k}_{u(i)} < \Tau_{(i)} \ , \ \forall i \in jt^{k}_{u(i)} \label{e.trip.constraints.2}
+\tau_{(i)} &< jt^{k}_{u(i)} < \Tau_{(i)} \ , \ \forall i \in jt^{k}_{u(i)} \label{e.trip.constraints.2},
 \end{align}
+{\color{blue}where $\tau$ is ... and $T_{(i)}$ ...}
 
 The first condition~\ref{e.trip.constraints.1} is straightforward and specifies that every trip should have at least one sighting.
+We acknowledge that thrips do exist which pass no ANPR cameras, and would thus have a value of $n$ equal to zero.
+However, for the purposes of the work here we cannot analyse such trips.
 The second condition~\ref{e.trip.constraints.2} defines a minimum and maximum travel times between consecutive observations.
@@
 \begin{enumerate*}[label=(\roman*)]
-  \item first, to allow trips made by the same vehicle to be differentiated.
-For instance, given two consecutive sightings of \emph{k} three hours apart, we want to interpret them as belonging to different trips of \emph{k};
-  \item second, it allows unplausible trips to be identified.
-For example, an unplausible trip can result from observing \emph{k} at a given camera and then a few seconds later at a second camera, several miles apart.
-Two explanations are common, either one of the cameras made a detection error, or there is another vehicle with a cloned plate number travelling in the road network.
-\end{enumerate*} Evidently, condition~\ref{e.trip.constraints.2} is only valid for trips of length two or greater.
+  \item first, to allow separate trips made by the same vehicle to be differentiated.
+For instance, given two consecutive sightings of \emph{k} three hours apart, we would want to interpret them as belonging to different trips of \emph{k};
+  \item second, it allows implausible trips to be identified.
+For example, an implausible trip can result from observing \emph{k} at a given camera and then a few seconds later at a second camera, several miles apart.
+Two explanations are common, either one of the cameras made a detection error, or there is another vehicle with a cloned number-plate travelling on the road network.
+\end{enumerate*} Evidently, condition~\ref{e.trip.constraints.2} is only valid for trips of length two sightings or greater.
 Nevertheless, trips can easily be differentiated by first sorting sightings by time of occurrence, then calculating the journey time sequence for the entire sequence and finally comparing each element against $\Tau$.
@@
   \caption{Example of a trip of length 4.
-On the left side the correspoding data table is shown.
+On the left side the corresponding data table is shown.
 The $u_{th}$ trip of each vehicle is given by the variable \emph{Trip}, whereas the $i_{th}$ sighting is given by variable \emph{Sighting}.
@@
 
-The simplest approaching to choosing the value of $\Tau$ is to pick a fixed empirical value, such as 5 or 10 minutes.
-However, if the distance between two cameras is greater than another origin-destination (od) pair of cameras, then it makes sense that $\Tau$ is relaxed.
+The simplest approach to choosing the value of $\Tau$ is to pick a fixed empirical value, such as 5 or 10 minutes.
+However, if the distance between two cameras is greater than another pair of cameras, then it makes sense that $\Tau$ is relaxed.
 Similarly, if there is an anomaly in the road network, such as a traffic jam, and the routes connecting the two cameras are affected, then the value of $\Tau$ should also be adapted.
 Hence, $\Tau$ should be a function of the distance between the two cameras (or, more accurately, of the top n-routes between these) and the distribution of observed journey times.
-The same rationale can be applied to estimating $\tau$.
+The same rationale can be applied to $\tau$.
 
@@
 Nevertheless, the importance of estimating these parameters should not be understated.
-Any errors in trip identification that occur due to poor estimation and filtering methods will propagate and be amplified in posterior analysis done using trip data.
+Any errors in trip identification that occur due to poor estimation and filtering methods will propagate and will be amplified in posterior analysis done using trip data.
 Therefore, $\Tau$ and $\tau$ estimation is something that needs to be carefully addressed and studied in the future in order to fully understand its impact in the outcome of the research.
@@
 
-To ensure that every trip of vehicle \emph{k} is unique in the sequence of all valid trips of vehicle \emph{k}:
+We need to ensure that every trip of vehicle \emph{k} is unique from all other trips of vehicle \emph{k}. I.e. given $W^k$ the set of all valid trips of $k$:
 
 \begin{align}
-W^{k} = \left( w^{k}_{(1)}, w^{k}_{(2)}, \ldots, w^{k}_{(N)} \right) \label{e.trip.history}
+W^{k} = \left( w^{k}_{(1)}, w^{k}_{(2)}, \ldots, w^{k}_{(N)} \right) \label{e.trip.history},
 \end{align}
@@
 \begin{align}
-s^{k}_{(u(i))} \neq s^{k}_{(v(j))}, \forall u,v &= 1, 2, \ldots, N \ , \ u \neq v,  \label{e.trip.history.constraint} \\
-\forall i,j &= 1, 2, \ldots, n \ , \ i \neq j \nonumber
+s^{k}_{(u,i)} \neq s^{k}_{(v,j)}, \forall u,v &= 1, 2, \ldots, N \ , \ u \neq v,  \label{e.trip.history.constraint} \\
+\forall i,j &= 1, 2, \ldots, n^u \ ,\nonumber \\
+\forall i,j &= 1, 2, \ldots, n^v \ , \nonumber
 \end{align}
+where $s_{(u,i)}^k$ is the $i^{th}$ sighting of vehicle $k$ during trip $u$, and $n^u$ is the number of sightings in trip $u$.
 
@@
 
+{\color{blue}where $c_i^k$ is ..., }
+
  Although the estimation of $\gamma$ carries similar considerations and consequences as those of estimating $\Tau$ and $\tau$, most duplicates can be identified in consecutive sightings of the same camera within the same trip.
-Even though a poor estimation of $\gamma$ also has impact in error propogation, this decreases substantially after filtering duplicates according to the heuristic above, due to the low occurrence of cycles in trips.
+Even though a poor estimation of $\gamma$ also has an impact on error propagation, this decreases substantially after filtering duplicates according to the heuristic above, due to the low occurrence of cycles in trips.
 Nevertheless, this issue needs to be fully investigated in future works.
@@
 Number plate recognition cameras have accuracy rates of 99.9\% or higher.
-If we consider that on average 1 to 10 out of 10000 scans are misclassified plate number then approximately 200 to 2000 scans everyday are incorrect scans.
+If we consider that on average 1 to 10 out of 10000 scans are misclassified number-plates then approximately 200 to 2000 scans everyday are incorrect.
 We categorise the errors made by ANPR cameras into two types:
@@
 The corresponding trip sequence for the affected vehicle will be missing a sighting but we will have no indication of this.
-On the second case, there is a recorded sighting, but this will be assigned to the incorrect vehicle.
+It should be noted that this could also happen in the case where a vehicle does not take the `normal' route and pass this intermediary camera.
+For the second case, there is a recorded sighting, but this will be assigned to the incorrect vehicle.
 Even though, the camera provides a value of \emph{confidence} which helps in diagnosing incorrect scans, errors still occur for high values of confidence.
-Even though that in this work we do not present a complex solution for this issue, methods to detect and address these two types of errors need to be designed and implemented in the future.
+Even so, in this work we do not present a complex solution for this issue.
+Methods to detect and address these two types of errors need to be designed and implemented in the future.

### 2018-02-01 discussion.tex
Fixed some grammar mistakes and added abstract
All sections now covered.

@@
 
-Most urban cities in the world possess a network of ANPR cameras that is used in law enforcement and for traffic monitoring and control.
+Most urban cities in the world possess a network of ANPR cameras that are used for law enforcement as well as traffic monitoring and control.
 Number plate data collected in the Tyne and Wear area is stored and leveraged by the UTMC for computing average journey times across a selection of sensitive roads.
 However, number plate data could be used more extensively to identify and study individual and collective travel patterns.
-In this paper, we've presented a set of definitions and constraints that establish a conceptual foundation for identifying vehicle trips from number plate detections.
-We've identified two parameters, $\tau$ and $\Tau$ as critical in the discrimination of plausible and unplausible trips.
+In this paper, we have presented a set of definitions and constraints that establish a conceptual foundation for identifying vehicle trips from number plate detections.
+We also identify two parameters, $\tau$ and $\Tau$ as critical in the discrimination of plausible and implausible trips.
 Hence, future work should first and foremost focus on developing formal methods to estimate these parameters from observed distributions of travel times and by applying knowledge about the structure of the road network.
@@
 
-Once trip data has been computed, a range of interesting applications is available.
+Once trip data has been computed, a range of interesting applications are available.
 This work tries to identify groups of vehicles by clustering information about frequency and diversity of travel.
@@
   \item real-time route recommendation using probabilistic graphical models;
-  \item detection of abnormal trip patterns for helping law enforcement in the identification of suspect vehicles or behavior;
+  \item detection of abnormal trip patterns for helping law enforcement in the identification of suspect vehicles or behaviour;
   \item modelling how drivers make routing choices in the presence of anomalies in the road network.

### 2018-02-01 paper.tex
Used bits from other sections as the related work. Had to move figure 1 around so that the structure of the paper would not be affected as a result.
All sections now covered.

@@
 
-\begin{abstract} \small\baselineskip=9pt In major cities, government agencies increasingly employ automatic number plate recognition (ANPR) technology in law enforcement and traffic control.
-In the region of Tyne and Wear (UK) the network of ANPR cameras is used to monitor travel times across sensitive roads.
-However, few works have explored the full potential of number plate scans for analysing individual and collective travel patterns.
+\begin{abstract} \small\baselineskip=9pt In major cities, government agencies increasingly employ automatic number-plate recognition (ANPR) technology in law enforcement and traffic control.
+In the Tyne and Wear region (UK) the network of ANPR cameras is used to monitor travel times across sensitive roads.
+However, few works have explored the full potential of number-plate scans for analysing individual and collective travel patterns.
 In this work we present a methodology for deriving trips from vehicle sightings at fixed camera locations.
 We illustrate the effect of parameters $\tau$ and $\Tau$ on trip discrimination and on the detection of implausible trips.
-To demonstrate the potential of trip data we apply \emph{k-means} clustering to trips identified from over 40 million plate scans recorded over 15 weekdays.
+To demonstrate the potential of trip data we apply \emph{k-means} clustering to trips identified from over 40 million plate scans recorded over fifteen weekdays.
 Results show that whilst private and transit travel modes can begin to be inferred from the resulting clusters, further work needs to be put into developing a more consistent and integrated framework for trip identification in ANPR data.\end{abstract}

### 2018-02-01 classification.tex

--- 2018-02-01 classification.tex
@@
 One direct application of trip data is unsupervised learning.
-Previous works have focused on identifying travel modes or purpose of trip from different sources of data, most notably: survey data, GPS data and mobile phone data~\cite{ODMobileData, ClusteringGPS}.
-Determining the distribution of travel modes is one of the fundamental steps in the four-stage model: an essential traffic modelling methodology for transportation planning~\cite{FourStepModel}.
-Although number-plate data has been used in~\cite{Clustering}, to identify different categories of trips, these do not differentiate between private or public travel modes and focus for the most part on when the trips occurred.
-Therefore, we aim to identify groups of vehicles with similar trip patterns based on frequency and diversity of travel.
-This can be private vehicles, such as work-home commuters, transit vehicles like buses and taxis, or other types of vehicles such as delivery trucks.
+In this section we identify groups of vehicles with similar trip patterns based on frequency and diversity of travel.
+Clusters can represent private vehicles, such as work-home commuters, transit vehicles like buses and taxis, or other types of vehicles such as delivery trucks.
 %This in turn, may enable trip modes to be estimated
-However, due to its limited coverage of the road network, it is not clear how well ANPR data can capture much the travel characteristics necessary to accurately describe the different groups.
+%However, due to its limited coverage of the road network, it is not clear how well ANPR data can capture much the travel characteristics necessary to accurately describe the different groups.
 

### 2018-02-01 dataset.tex
Fixed some grammar mistakes and added abstract

@@
 \section{Tyne and Wear ANPR Data}\label{s.ncl}
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=.95\linewidth]{ANPR-overview.pdf}
-\caption{Automatic number plate recognition system overview.}
-\label{fig:anpr-overview}
-\end{figure}
 
@@
   15244177 & 115 & 1454284802.83 &  18 &  87 \\
-  6756787 & 146 & 1454284801.53 &  22 &  99 \\
-  8487265 & 2 & 1454284803.88 &  10 &  93 \\
    \hline
@@
 \label{t:np_data_example}
+%\vspace{-2cm}
 \end{table}
-
-{\color{red}{\color{blue}Much of this paragraph should be in related work}
-The main use of ANPR data for the UTMC is estimating average journey times for selected links in the road network.
-Furthermore, several authors have extensively researched how to use number-plate data as an extension to link counts for estimating origin-destination matrices and link flows~\cite{Castillo2010, Castillo2008, Hazelton2012}.
-However, very few works have focused on analysing individual or collective travel patterns from number-plate data, particularly across extended periods of time.
-Moreover, there is no consistent conceptual and analytical framework for transforming number plate data into a historical sequence of trips for each vehicle.
-Finally, we believe that trip data, properly identified from number-plate data, has the potential to unlock a number of new applications for urban traffic control and law enforcement.
-Thus, in the following section we present a conceptual methodology for grouping multiple camera observations of the same vehicle into one or several trips of that vehicle.
-}

### 2018-02-01 intro.tex

--- 2018-01-31 intro.tex
@@
 
-%Even though the cameras only cover a very small subset of the entire road network, the network of cameras provides a valuable time-varying picture of travel demands and traffic states.
+\begin{figure}[t]
+\centering
+\includegraphics[width=.95\linewidth]{ANPR-overview.pdf}
+\caption{Automatic number plate recognition system overview.}
+\label{fig:anpr-overview}
+\end{figure}
 
-In ~\cite{Castillo2008}, the authors used partial trip information from plate number scans to compute travel demands, in the form of a trip matrix, and estimate link flows from a set of constraints composed of traffic conservation laws, prior knowledge and path flow disaggregation techniques, such as Stochastic User Equilibrium (SUE).
+The main use of ANPR data for the UTMC is estimating average journey times for selected or sensitive links in the road network.
+Furthermore, several authors have extensively researched how to use number-plate data as an extension to link counts for estimating origin-destination matrices and link flows~\cite{Castillo2010, Castillo2008, Hazelton2012}.
+However, very few works have focused on analysing individual or collective travel patterns from number-plate data, particularly across extended periods of time.
+Moreover, there is no consistent conceptual and analytical framework for transforming number plate data into a historical sequence of trips for each vehicle.
+Finally, we believe that trip data, properly identified from number-plate data, has the potential to unlock a number of new applications for urban traffic control and law enforcement.
+Thus, in section~\ref{s.trips} we present a conceptual methodology for grouping multiple camera observations of the same vehicle into one or several trips of that vehicle.
+
+Determining the distribution of travel modes is one of the fundamental steps in the four-stage model: an essential traffic modelling methodology for transportation planning~\cite{FourStepModel}.
+Previous works have used trip information derived from different sources of data to identify travel mode or purpose of trip.
+More notably, survey data, floating car data and mobile phone data have been used~\cite{ODMobileData, ClusteringGPS}.
+Although number plate data has been used in~\cite{Clustering}, to identify different categories of trips, the authors do not differentiate between private or public travel modes and focus instead on categorising trips by time of occurrence.
+Hence, in section~\ref{s.classification} we apply the \emph{k-means} clustering algorithm to derived trip data and based on the results, we discuss the limitations of proposed methods and that of ANPR data.
 
@@
 
-Uses of trip data for understandint travel demands / traffic prediction
-
-Studies about best placement of cameras
-
-Number plate data is not as popular, for urban traffic monitoring and prediction , as other sources of data, namely GPS traces or floating vehicle data.
-Previous works trips from GPS traces, .
-
-%Therefore, number plate scans provide additional trip information, besides simple link counts, that has
-
- %estimating travel demands and traffic states. extended traditional methods of  travel demands
-
-Although significant research has gone into identifying trips from GPS traces, to the best of our knowledge, little research has gone into doing so with number plate data.
+% {\color{red}Uses of trip data for understanding travel demands / traffic prediction
+%
+% Studies about best placement of cameras
+%
+% Number plate data is not as popular, for urban traffic monitoring and prediction, as other sources of data, namely GPS traces or floating vehicle data.
+% Previous works trips from GPS traces, .
+%
+% %Therefore, number plate scans provide additional trip information, besides simple link counts, that has
+%
+%  %estimating travel demands and traffic states. extended traditional methods of  travel demands
+%
+%
+% Although significant research has gone into identifying trips from GPS traces, to the best of our knowledge, little research has gone into doing so with number plate data.
 Therefore, the contributions of this paper are twofold:

### 2018-02-01 trips.tex
some squeezing latex

@@
 Let the $i_{th}$ sighting of vehicle \emph{k} be defined as the unordered pair:
-
 \begin{equation} \label{e.sighting}
@@
 \end{equation}
-
 where \emph{c} uniquely identifies a camera, and \emph{t} is a scalar representing a point in time (e.g.\ a timestamp).
@@
 Let an ordered sequence of sightings of vehicle \emph{k} define the $u_{th}$ trip of vehicle \emph{k}:
-
 \begin{equation} \label{e.trip}
@@
 \end{equation}
-
 where \( n \) is the length of the trip, i.e.\ the number of sightings.
 Moreover, let the corresponding journey time sequence, of length \(n-1\), be defined as the time difference of consecutive sightings:
-
 \begin{equation} \label{e.journeytime}
@@
 \end{equation}
-
 We consider a trip of \emph{$w^k_u$} valid if the following conditions hold:
-
 \begin{align}

### 2018-02-01 trips.tex
Added changes proposed in blue and simplified enumerations to save space

--- 2018-02-01 trips.tex
@@
 \end{align}
-{\color{blue}where $\tau$ is ... and $T_{(i)}$ ...}
+where $\tau_{(i)}$ and $\Tau_{(i)}$ are the lower and upper bound of the $i_{th}$ element of the journey time sequence.
 
@@
 
-{\color{blue}where $c_i^k$ is ..., }
+where $c_i^k$ and $t_i^k$ are the camera and timestamp of the $i_{th}$ sighting of vehicle $k$.
 
@@
 We categorise the errors made by ANPR cameras into two types:
-\begin{enumerate}
+\begin{enumerate*}[label=(\roman*)]
   \item a passing vehicle is not detected by the camera;
   \item a different vehicle is detected instead.
-\end{enumerate}
+\end{enumerate*}
 In the first case, there is no observed data.

### 2018-02-01 classification.tex
Added changes proposed in blue and simplified enumerations to save space
Final version finished. Finished results description and fixed bad boxes in bibliography. Finally managed to fit everything with a few vspaces here and there

@@
 \label{t:trips-tau}
+\vspace{-0.5cm}
 \end{table}
@@
 Transforming trips into features which can be used in clustering algorithms was a 3-step process:
-\begin{enumerate}
+\begin{enumerate*}[label=(\roman*)]
   \item First, every trip was summarised as a single row of data. The following information was extracted: length of trip, origin, destination, route, start and end times.
@@
   \item Finally, daily information per vehicle was collapsed into a single row by averaging this information across the 15 days.
-\end{enumerate}
+\end{enumerate*}
 
@@
 A total of 1034107 distinct vehicles were detected.
-However, because a high percentage of trips containing a single sighting, some of these features were highly correlated.
+However, because there is a high percentage of trips containing a single sighting, some of these features were highly correlated.
 We therefore, chose to remove three of the features represented in~\ref{t:features}: \emph{Average Sightings}, \emph{Average Distinct Routes} and \emph{Average Hour Difference}, to avoid the obfuscation of the natural clustering ~\cite{Kmeans}.
@@
 \caption{Sample of extracted features from trips taken from 15 weekdays of number plate data.}
+\vspace{-0.2cm}
 \label{t:features}
@@
 
-Table~\ref{t:tau_comparison} provides a summary of multiple runs of \emph{k-means} for each value of $\Tau$.
-The optimal number of clusters, selected by maximising the Calinski-Harabasz criterion, is presented and is shown to decrease inversely with $\Tau$.
-This in part expected because as $\Tau$ increases, so does the average trip length, as opposed to the total number of trips, and trips per vehicle.
-{\color{blue}The interp   ...} The relative between and average within clusters sum of squares is also shown.
+Table~\ref{t:tau_comparison} provides a summary of multiple runs of \emph{k-means}.
+For each value of $\Tau$, the optimal number of clusters is selected by picking the value of $k$ that maximises the Calinski-Harabasz criterion.
+Furthermore, the measures of inter-cluster (betweeness) and intra-cluster (withiness) are depicted relative to the corresponding total sum of squares (total betweeness and total withiness).
+It is noteworthy that the best value of $k$ increases inversely to $\Tau$.
+Although a higher value of $k$ seems to suggest that trips with smaller values of $\Tau$ are able capture the variance in the data better, we have to consider that varying $\Tau$ affects the average trip length in the same direction whilst affecting the total number of trips in the opposite direction (table~\ref{t:trips-tau}).
 
 Tables~\ref{t:kmeans_centers_450} and~\ref{t:kmeans_centers_1200} depict the cluster centres for $\Tau$ equal to 7.5 minutes and 20 minutes respectively.
-The values for the cluster centres partially meet our expectations.
-We were expecting to find a relatively small cluster representing taxis with a high average number of trips per day, occurring over a variety of origins and destinations and over a large time frame.
+These partially meet our expectations.
+For instance, we were expecting to find a relatively small cluster representing taxis with a high average number of trips per day, occurring over a variety of origins and destinations and over a large time frame.
 Clusters 2 and 7 for $\Tau = 7.5$ and cluster 4 for $\Tau = 20$ do indeed fit this profile.
-What differentiates cluster 2 from 7 in the first case is essentially the mean number of trips per day and the average time at which the first and last trips of the day occur.
-We expected that buses could be separated from taxis by showing less diversity in the number of origins and destinations as these essentially do multiple runs of the same trip throughout the day.
+What differentiates cluster 2 from 7 in the first case is essentially a lower mean number of trips per day and smaller average time at which the first and last trips of the day occur.
+Additionally, buses to some extent fit this category as well, however we expected that buses could be separated from taxis by showing less diversity in the number of origins and destinations as these essentially do multiple runs of the same trip throughout the day.
 Still, this can be explained by the fact that buses take routes through main and secondary roads.
 As most cameras are placed in main roads, the one long bus trip can be perceived as multiple small trips as the bus alternates between arterial and main roads.
+This is in fact, one of the downsides of ANPR data and is something that the methodology presented in section~\ref{s.trips} needs to develop in the future.
 
-On the other hand, we could expect that commuters display trips starting at around eight in the morning and a second trip terminating between five and six in the evening.
-Additionally, we would expect to see few trips per day as most of the day is spent in work.
+On the other hand, we expected to observe a group representing home to work commuters with the first trip of the day starting approximately at eight in the morning and a second trip terminating between five and six in the evening.
+Although we observe one or two groups with those characteristics, these contain a higher average of trips per day than expected, which may represent for example work to school trips.
+However, we observe at least one big group of trips occurring mostly during lunch hour.
+Several interpretations are possible but further work is needed to provide a more consistent interpretation and validation of these results.
+A big contributing factor however is the fact that a high proportion of trips contains only a single sighting (table~\ref{t:trips-tau}).
+It may be the case that commuters choose other routes than those going through ANPR cameras.
+Still, we need to devise further models and methods that are able to capture this uncertainty.
 
@@
 \label{t:tau_comparison}
+\vspace{-0.5cm}
 \end{table}
@@
 \end{table*}
-
-The differences between expected and observed results are in great part due to the fact that a high percentage of trips contains a single sighting.
-It may be the case that commuters choose routes that do not pass through ANPR cameras.
-Devising further models and methods that model and capture this uncertainty is fundamental to getting the most out of ANPR data.

### 2018-02-01 intro.tex
Final version finished. Finished results description and fixed bad boxes in bibliography. Finally managed to fit everything with a few vspaces here and there

@@
 In order to better understand how we can deal with this increase in demand we need to better understand how the road network is being used.
-By understanding road usage we can better deal with congestion, handle traffic incidents, plan road modifications and deal with illegal acts on the roads.
+By understanding road usage we can better deal with congestion, handle traffic incidents, plan road modifications and deal with illegal acts.
 
@@
 \centering
-\includegraphics[width=.95\linewidth]{ANPR-overview.pdf}
-\caption{Automatic number plate recognition system overview.}
+\includegraphics[width=.90\linewidth]{ANPR-overview.pdf}
+\caption{Overview of an ANPR system.}
 \label{fig:anpr-overview}
+\vspace{-0.6cm}
 \end{figure}

### 2018-02-01 paper.tex

@@
 \usepackage{placeins}
+\usepackage{microtype}
 
@@
 \and
-Stephen McGough\footnotemark[1]}
+A. Stephen McGough\footnotemark[1]}
 \date{}
@@
 Department for Transport {\em Provisional Road Traffic Estimates Great Britain: October 2016-September 2017}.
-Department for Transport WebSite: https://www.gov.uk/government/statistics/provisional-road-traffic-estimates-great-britain-october-2016-to-september-2017.
+url (visited January 2017): https://www.gov.uk\-/government/statistics/provisional-road-traffic-estima\-tes-great-britain-october-2016-to-september-2017.
 
@@
 Department for Transport {\em Vehicle Licensing Statistics: Quarter 3 (Jul-Sep) 2017}.
-Department for Transport WebSite: https://www.gov.uk/government/statistics/vehicle-licensing-statistics-july-to-september-2017.
+url (visited 31 January 2017): https://www.gov.uk/government/statistics\-/vehicle-licensing-statistics-july-to-september-2017.
 
@@
 
-\bibitem{ShortestPath?2015}
-Zhu, Shanjiang, and David Levinson. {\em Do people use the shortest path? An empirical test of Wardrop’s first principle}. PloS one 10.8 (2015): e0134322.
+% \bibitem{ShortestPath?2015}
+% Zhu, Shanjiang, and David Levinson. {\em Do people use the shortest path? An empirical test of Wardrop’s first principle}. PloS one 10.8 (2015): e0134322.
 

### 2018-02-21 trips.tex
Fixed the following issues: Font color red to black, formula 4.3 fixed, re-written paragraph about constraint 4.4

@@
 \begin{equation} \label{e.journeytime}
-jt^{k}_{u} = \left(t^{k}_{(1)} - t^{k}_{(2)}, \ldots, t^{k}_{(n)} - t^{k}_{(n-1)} \right).
+jt^{k}_{u} = \left(t^{k}_{(2)} - t^{k}_{(1)}, \ldots, t^{k}_{(n)} - t^{k}_{(n-1)} \right).
 \end{equation}
@@
 
-The first condition~\ref{e.trip.constraints.1} is straightforward and specifies that every trip should have at least one sighting.
-We acknowledge that thrips do exist which pass no ANPR cameras, and would thus have a value of $n$ equal to zero.
-However, for the purposes of the work here we cannot analyse such trips.
+The first condition~\ref{e.trip.constraints.1} is straightforward and specifies that every identified trip should have at least one sighting.
+Obviously, vehicles can make trips that do not pass through any ANPR cameras and thus have no associated sightings: $n = 0$.
+However, this work focus on trips that we can observe and hence we consider that $n > 0$.
 The second condition~\ref{e.trip.constraints.2} defines a minimum and maximum travel times between consecutive observations.

### 2018-02-21 classification.tex
Changed structure of the paper. Added main Methodology section with subsections
A few corrections and changes. Changed the name of length to degree, removed most references to future work. Rephrased a few sentences.

@@
 % !TEX root = paper.tex
-\section{Clustering vehicles}\label{s.classification}
+\subsection{Clustering trips}\label{s.classification}
 
@@
 Furthermore, as mentioned in section~\ref{s.trips}, we used fixed empirical values for $\Tau$.
-Table~\ref{t:trips-tau} displays the number of trips, average trip length and the proportion of trips of length one, for varying values of $\Tau$.
-The effect of incrementing $\Tau$ on resulting trips is clear: increased trip lengths and decreased trips containing just a single sighting.
+Table~\ref{t:trips-tau} displays the number of trips, average trip degree and the proportion of trips of degree one, for varying values of $\Tau$.
+The effect of incrementing $\Tau$ on resulting trips is clear: increased trip degrees and decreased trips containing just a single sighting.
 On the other hand, we did not set a value for $\tau$, but instead handled implausible trips by filtering all sightings with confidence below 85\%.
@@
   \hline
-Tau & Trips & \thead{Average\\Length} & \thead{Proportion of Trips \\ Length 1} \\
+Tau & Trips & \thead{Average\\Degree} & \thead{Proportion of Trips \\ Degree 1} \\
   \hline
@@
 \begin{enumerate*}[label=(\roman*)]
-  \item First, every trip was summarised as a single row of data. The following information was extracted: length of trip, origin, destination, route, start and end times.
-  \item Second, daily trip information was obtained for each vehicle: number of trips, median of trip lengths, number of sightings, distinct number of origins destinations, and routes, hour of first sighting, hour of last sighting and total rest time between trips.
+  \item First, every trip was summarised as a single row of data. The following information was extracted: degree of trip, origin, destination, route, start and end times.
+  \item Second, daily trip information was obtained for each vehicle: number of trips, median of trip degrees, number of sightings, distinct number of origins destinations, and routes, hour of first sighting, hour of last sighting and total rest time between trips.
   \item Finally, daily information per vehicle was collapsed into a single row by averaging this information across the 15 days.
@@
 We therefore, chose to remove three of the features represented in~\ref{t:features}: \emph{Average Sightings}, \emph{Average Distinct Routes} and \emph{Average Hour Difference}, to avoid the obfuscation of the natural clustering ~\cite{Kmeans}.
-Furthermore we considered that a trip of length one has no destination (which explains values of average distinct destination below zero) and we filtered all instances of vehicles where the total number of trips is lower than 3, resulting in 642006 unique vehicles.
+Furthermore we considered that a trip of degree one has no destination (which explains values of average distinct destination below zero) and we filtered all instances of vehicles where the total number of trips is lower than 3, resulting in 642006 unique vehicles.
 
@@
   \hline
- \thead{Total\\Trips} & \thead{Average\\Trips} & \thead{Average\\Length} & \thead{Average\\Sightings} & \thead{Average\\Distinct\\Origins} & \thead{Average\\Distinct\\Destinations} & \thead{Average\\Distinct\\Routes} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} & \thead{Average\\Hour\\Difference} & \thead{Average\\Rest\\Time} \\
+ \thead{Total\\Trips} & \thead{Average\\Trips} & \thead{Average\\Degree} & \thead{Average\\Sightings} & \thead{Average\\Distinct\\Origins} & \thead{Average\\Distinct\\Destinations} & \thead{Average\\Distinct\\Routes} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} & \thead{Average\\Hour\\Difference} & \thead{Average\\Rest\\Time} \\
   \hline
@@
 It is noteworthy that the best value of $k$ increases inversely to $\Tau$.
-Although a higher value of $k$ seems to suggest that trips with smaller values of $\Tau$ are able capture the variance in the data better, we have to consider that varying $\Tau$ affects the average trip length in the same direction whilst affecting the total number of trips in the opposite direction (table~\ref{t:trips-tau}).
+Although a higher value of $k$ seems to suggest that trips with smaller values of $\Tau$ are able capture the variance in the data better, we have to consider that varying $\Tau$ affects the average trip degree in the same direction whilst affecting the total number of trips in the opposite direction (table~\ref{t:trips-tau}).
 
@@
 \label{t:tau_comparison}
-\vspace{-0.5cm}
 \end{table}
 
-\begin{table*}[t]
+\begin{table}[t]
 \centering
-\small
-\begin{tabular}{c c c c c c c c c c}
+\tabcolsep=0.15cm
+\begin{tabular}{c c c c c c}
   \hline
- Cluster &  Size & \thead{Total\\Trips} & \thead{Average\\Trips} & \thead{Average\\Length} & \thead{Average\\Distinct\\Origins} & \thead{Average\\Distinct\\Destinations} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} & \thead{Average\\Rest\\Time} \\
+ Cluster & Size & \thead{Average\\Trips} & \thead{Average\\Distinct\\Origins} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} \\
   \hline
-  1 & 108868 & 30.67 & 2.91 & 1.56 & 2.65 & 1.09 & 9.67 & 16.11 & 6.35 \\
-  2 & 2948 & 170.08 & 13.69 & 1.47 & 9.20 & 4.67 & 6.89 & 18.69 & 11.65 \\
-  3 & 309748 & 5.95 & 2.01 & 1.44 & 1.92 & 0.68 & 12.52 & 14.50 & 1.93 \\
-  4 & 163982 & 16.99 & 2.30 & 1.46 & 2.13 & 0.77 & 11.12 & 14.99 & 3.81 \\
-  5 & 45414 & 50.30 & 4.26 & 1.52 & 3.67 & 1.57 & 9.04 & 16.86 & 7.69 \\
-  6 & 10380 & 86.99 & 7.16 & 1.43 & 5.55 & 2.49 & 8.41 & 17.60 & 9.03 \\
-  7 & 666 & 331.50 & 26.43 & 1.57 & 9.31 & 5.17 & 5.73 & 19.59 & 13.14 \\
+1 & 157962 & 2.45 & 2.27 & 11.38 & 14.93 \\
+  2 & 303513 & 2.18 & 2.09 & 12.53 & 14.47 \\
+  3 & 21549 & 6.02 & 5.00 & 8.68 & 17.23 \\
+  4 & 108094 & 2.99 & 2.73 & 9.99 & 15.86 \\
+  5 & 509 & 33.62 & 10.42 & 5.81 & 19.67 \\
+  6 & 1993 & 17.99 & 11.24 & 6.45 & 19.09 \\
+  7 & 4971 & 10.41 & 7.64 & 7.97 & 18.04 \\
+  8 & 58059 & 4.09 & 3.62 & 9.22 & 16.60 \\
    \hline
 \end{tabular}
-\caption{Clusters sizes and centres for $\Tau = 7.5$ minutes.}
-\label{t:kmeans_centers_450}
-\end{table*}
-
-\begin{table*}[t]
-\centering
-\small
-\begin{tabular}{c c c c c c c c c c}
-  \hline
- Cluster &  Size & \thead{Total\\Trips} & \thead{Average\\Trips} & \thead{Average\\Length} & \thead{Average\\Distinct\\Origins} & \thead{Average\\Distinct\\Destinations} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} & \thead{Average\\Rest\\Time} \\
-  \hline
-  1 & 371318 & 7.19 & 1.76 & 1.68 & 1.68 & 0.71 & 12.30 & 14.55 & 2.14 \\
-    2 & 49852 & 44.85 & 3.73 & 1.90 & 3.16 & 1.73 & 8.98 & 17.02 & 7.76 \\
-    3 & 189224 & 22.82 & 2.29 & 1.84 & 2.11 & 1.06 & 10.00 & 15.76 & 5.60 \\
-    4 & 4593 & 114.34 & 9.12 & 2.30 & 5.92 & 3.98 & 6.93 & 18.21 & 10.43 \\
-   \hline
-\end{tabular}
-\caption{Clusters sizes and centres for $\Tau = 20$ minutes.}
-\label{t:kmeans_centers_1200}
-\end{table*}
+\caption{Cluster sizes and centres for $\Tau = 5$ minutes.}
+\label{t:kmeans_centers_300}
+\end{table}

### 2018-02-21 dataset.tex
Final version finished. Finished results description and fixed bad boxes in bibliography. Finally managed to fit everything with a few vspaces here and there
Changed structure of the paper. Added main Methodology section with subsections
A few corrections and changes. Changed the name of length to degree, removed most references to future work. Rephrased a few sentences.

@@
 % !TEX root = paper.tex
-\section{Tyne and Wear ANPR Data}\label{s.ncl}
+\section{Methodology}
+\subsection{Tyne and Wear ANPR Data}\label{s.ncl}
 
-Automatic number-plate recognition (ANPR) cameras are actively employed in urban traffic environments and play an important role in day-to-day intelligent transportation systems.
-They can be used by government subsidised entities in urban traffic management and control; by commissioned highway agencies in electronic toll collection; or by law enforcement organisations in detecting speeding vehicles and validating number plate registrations.
-The wide diversity of applications, paired with the large improvements in price-to-performance ratios of ANPR hardware and software systems, has resulted in increased investments of ANPR cameras for urban environments~\cite{EvolutionUTMC2013, SurveyITS2011}.
-
-\begin{figure*}[t]
-\centering
-\begin{subfigure}[t]{.48\textwidth}
-  \centering
-  \includegraphics[width=.9\linewidth]{observations_per_day.pdf}
-  \caption{Total number of scans recorded per day in Tyne and Wear. There is a clear seasonal effect caused by decreasing traffic demands at weekends and increasing traffic volume during weekdays.}
-  \label{fig:observations-per-day}
-\end{subfigure}\hfill
-\begin{subfigure}[t]{.48\textwidth}
-  \centering
-  \includegraphics[width=.9\linewidth]{observations_per_day_camera.pdf}
-  \caption{Number of scans recorded per ANPR camera and day in Tyne and Wear.
-Inter-camera variability is observed, as some cameras are located in more traffic intensive road sections than others.
-Decommissioned or temporarily unavailable cameras (due to loss of power, faulty camera, road closed, etc) are depicted at the bottom of the graph.}
-  \label{fig:observations-per-camera-day}
-\end{subfigure}
-\caption{License plate scans recorded by ANPR cameras during February 2017, in the region of Tyne and Wear, United Kingdom.}
-\label{fig:time-series}
-\end{figure*}
-
-In the region of Tyne and Wear, United Kingdom, there are over 250 active ANPR cameras.
-Over 1 million license plate detections are recorded by these cameras every day.
-Figure~\ref{fig:time-series} shows the number of daily scans recorded over a month (February, 2017).
-Furthermore, every scan is stored in a central database managed by Urban Traffic Management and Control (UTMC) Tyne and Wear, and used to compute travel times across particular links of interest in the road network.
-These are usually major roads that see high volumes of traffic, or road segments more prone to traffic jams.
-Average journey times can then be conveyed back to the drivers by the way of Variable Message Signs (VMS) or web based applications.
-Figure~\ref{fig:anpr-overview} represents this interaction.
-
-Number plate data, in its essence, is a stream of events, each representing a vehicle observed by one camera at a specific point in time.
-An excerpt of the data can be found in table~\ref{t:np_data_example}.
-All number-plates were anonymised by the UTMC through a hashing algorithm before the data was shared.
-Cameras are uniquely identified by an integer and timestamps are relative to each camera's clock.
-However, the cameras are connected through a private network which provides clock synchronisation using the Network Time Protocol (NTP).
-Therefore, the timestamps can be used directly if the synchronisation error is negligible.
-The following additional information is also captured and provided by each camera:
-\begin{enumerate*}[label=(\roman*)]
-  \item the clock synchronisation error (milliseconds);
-  \item the camera's confidence that the identified number plate is the true number plate (percentage);
-  \item the direction of travel, away or towards the camera.
-\end{enumerate*}
-The confidence in the observation is especially useful as it helps diagnosing license plate recognition errors.
-On the other hand, the direction of travel is dependent upon the orientation of the camera, which is not provided.
-Hence, we chose to ignore the latter in this work, and aim to explore this information in future works.
-
-\begin{table}[ht]
+\begin{table}[!ht]
 \centering
@@
 \end{table}
+
+Automatic num\-ber plate recognition (ANPR) cameras are actively employed in urban traffic environments and play an important role in day-to-day intelligent transportation systems.
+They can be used by government subsidised entities in urban traffic management and control; by commissioned highway agencies in electronic toll collection; or by law enforcement organisations in detecting speeding vehicles and validating number plate registrations.
+The wide diversity of applications, paired with the large improvements in price-to-performance ratios of ANPR hardware and software systems, has resulted in increased investments of ANPR cameras for urban environments~\cite{EvolutionUTMC2013, SurveyITS2011}.
+
+\begin{figure*}[t]
+\centering
+\begin{subfigure}[t]{.48\textwidth}
+  \centering
+  \includegraphics[width=.9\linewidth]{observations_per_day.pdf}
+  \caption{Total number of scans recorded per day in Tyne and Wear. There is a clear seasonal effect caused by decreasing traffic demands at weekends and increasing traffic volume during weekdays.}
+  \label{fig:observations-per-day}
+\end{subfigure}\hfill
+\begin{subfigure}[t]{.48\textwidth}
+  \centering
+  \includegraphics[width=.9\linewidth]{observations_per_day_camera.pdf}
+  \caption{Number of scans recorded per ANPR camera and day in Tyne and Wear.
+Inter-camera variability is observed, as some cameras are located in more traffic intensive road sections than others.
+Decommissioned or temporarily unavailable cameras (due to loss of power, faulty camera, road closed, etc) can be identified at the bottom.}
+  \label{fig:observations-per-camera-day}
+\end{subfigure}
+\caption{License plate scans recorded by ANPR cameras during February 2017, in the region of Tyne and Wear, United Kingdom.}
+\label{fig:time-series}
+\end{figure*}
+
+In the region of Tyne and Wear, United Kingdom, there are over 250 active ANPR cameras.
+Over 1 million license plate detections are recorded by these cameras every day.
+Figure~\ref{fig:time-series} shows the number of daily scans recorded over a month (February, 2017).
+Furthermore, every scan is stored in a central database managed by Urban Traffic Management and Control (UTMC) Tyne and Wear, and used to compute travel times across particular links of interest in the road network.
+These are usually major roads that see high volumes of traffic, or road segments more prone to traffic jams.
+Average journey times can then be conveyed back to the drivers by the way of Variable Message Signs (VMS) or web based applications.
+Figure~\ref{fig:anpr-overview} represents this interaction.
+
+Number plate data, in its essence, is a stream of events, each representing a vehicle observed by one camera at a specific point in time.
+An excerpt of the data can be found in table~\ref{t:np_data_example}.
+All number plate were anonymised by the UTMC through a hashing algorithm before the data was shared.
+Cameras are uniquely identified by an integer and timestamps are relative to each camera's clock.
+Clock synchronisation is performed using the Network Time Protocol (NTP), as the cameras are connected, through a private network, to a central server.
+Therefore, the recorded timestamps can be used directly if the synchronisation error is negligible.
+The following additional information is also captured and provided by each camera:
+\begin{enumerate*}[label=(\roman*)]
+  \item the clock synchronisation error (milliseconds);
+  \item the camera's confidence that the identified number plate is the true number plate (percentage);
+  \item the direction of travel, away or towards the camera.
+\end{enumerate*}
+The confidence in the observation is especially useful as it helps diagnosing license plate recognition errors.
+On the other hand, the direction of travel is dependent upon the orientation of the camera, which is not provided.
+Hence, we chose to ignore the latter in this work.

### 2018-02-21 discussion.tex
A few corrections and changes. Changed the name of length to degree, removed most references to future work. Rephrased a few sentences.

@@
 
-Most urban cities in the world possess a network of ANPR cameras that are used for law enforcement as well as traffic monitoring and control.
+Most urban cities in the world employ a network of ANPR cameras that are used for law enforcement as well as traffic monitoring and control.
 Number plate data collected in the Tyne and Wear area is stored and leveraged by the UTMC for computing average journey times across a selection of sensitive roads.
@@
 \end{enumerate*}
+
+Hence, we chose to ignore camera direction in this work, and aim to explore this information in future works.
+
+On the other hand, a consistent set of models and methodologies for estimating $\Tau$ and $\tau$, under a variety of circumstances and road anomalies, remains an open research problem.
+Nevertheless, the importance of estimating these parameters should not be understated.
+Any errors in trip identification that occur due to poor estimation and filtering methods will propagate and will be amplified in posterior analysis done using trip data.
+Therefore, $\Tau$ and $\tau$ estimation is something that needs to be carefully addressed and studied in the future in order to fully understand its impact in the outcome of the research.

### 2018-02-21 intro.tex
Fixed the following issues: Font color red to black, formula 4.3 fixed, re-written paragraph about constraint 4.4
A few corrections and changes. Changed the name of length to degree, removed most references to future work. Rephrased a few sentences.

@@
 Alternatively one can view the problem the other way around and rather than tracking individual vehicles look at collecting information by observing vehicles passing points within the road networks.
-A prime example of this approach are Automatic Number-plate Recognition (ANPR) cameras.
-These cameras are a combination of digital camera coupled with Artificial Intelligence to identify number-plates within the image and convert these into strings of characters.
+A prime example of this approach are Automatic Number Plate Recognition (ANPR) cameras.
+These cameras are a combination of digital camera coupled with Artificial Intelligence to identify number plates within the image and convert these into strings of characters.
 ANPR cameras are normally fixed in location\footnote{Although cameras can be in a vehicle and moved from location to location.} able to view all vehicles passing that location.
@@
 It should be noted that we cannot determine the actual start and end of the journey as these will happen in areas not covered by ANPR.
-It should also be noted that for ethical reasons it is not normal to obtain actual number-plates, but rather the hash of these.
+It should also be noted that for ethical reasons it is not normal to obtain actual number plates, but rather the hash of these.
 Though, for most situations this will suffice.
@@
 \begin{itemize}
-	\item {\bf Number-plate miss-reads:} Although ANPR cameras have accuracies of around {\color{red}99\%}, miss-reads are possible.
+	\item {\bf Number plate miss-reads:} Although ANPR cameras have accuracies of around 99, miss-reads are possible.
 This can lead to sightings being missed or vehicles being wrongly sighted in locations.
@@
 The minor side of this is implausible journey times, though, more seriously, this can lead to reordering the set of cameras on a particular journey.
-	\item {\bf Clones number-plates:} For various reasons a number-plate may be cloned and used on a different vehicle.
+	\item {\bf Clones number plates:} For various reasons a number plate may be cloned and used on a different vehicle.
 This can lead to impossible journeys and journeys that the real vehicle did not make.
@@
 The main use of ANPR data for the UTMC is estimating average journey times for selected or sensitive links in the road network.
-Furthermore, several authors have extensively researched how to use number-plate data as an extension to link counts for estimating origin-destination matrices and link flows~\cite{Castillo2010, Castillo2008, Hazelton2012}.
-However, very few works have focused on analysing individual or collective travel patterns from number-plate data, particularly across extended periods of time.
+Furthermore, several authors have extensively researched how to use number plate data as an extension to link counts for estimating origin-destination matrices and link flows~\cite{Castillo2010, Castillo2008, Hazelton2012}.
+However, very few works have focused on analysing individual or collective travel patterns from number plate data, particularly across extended periods of time.
 Moreover, there is no consistent conceptual and analytical framework for transforming number plate data into a historical sequence of trips for each vehicle.
-Finally, we believe that trip data, properly identified from number-plate data, has the potential to unlock a number of new applications for urban traffic control and law enforcement.
+Finally, we believe that trip data, properly identified from number plate data, has the potential to unlock a number of new applications for urban traffic control and law enforcement.
 Thus, in section~\ref{s.trips} we present a conceptual methodology for grouping multiple camera observations of the same vehicle into one or several trips of that vehicle.
@@
 More notably, survey data, floating car data and mobile phone data have been used~\cite{ODMobileData, ClusteringGPS}.
-Although number plate data has been used in~\cite{Clustering}, to identify different categories of trips, the authors do not differentiate between private or public travel modes and focus instead on categorising trips by time of occurrence.
+Although number plate data has been used in~\cite{Clustering} to identify different categories of trips, the authors do not differentiate between private or public travel modes and focus instead on categorising trips by time of occurrence.
 Hence, in section~\ref{s.classification} we apply the \emph{k-means} clustering algorithm to derived trip data and based on the results, we discuss the limitations of proposed methods and that of ANPR data.

### 2018-02-21 trips.tex
Changed structure of the paper. Added main Methodology section with subsections
A few corrections and changes. Changed the name of length to degree, removed most references to future work. Rephrased a few sentences.

@@
 % !TEX root = paper.tex
-\section{Trip Identification}\label{s.trips}
-
-In this section we explain our technique for identifying which sets of sightings can be combined in order to make a trip.
+\subsection{Trip Identification}\label{s.trips}
 
@@
 \end{equation}
-where \( n \) is the length of the trip, i.e.\ the number of sightings.
-Moreover, let the corresponding journey time sequence, of length \(n-1\), be defined as the time difference of consecutive sightings:
+where \( n \) is the degree of the trip, i.e.\ the number of sightings.
+Moreover, let the corresponding journey time sequence, of degree \(n-1\), be defined as the time difference of consecutive sightings:
 \begin{equation} \label{e.journeytime}
@@
 \end{equation}
-We consider a trip of \emph{$w^k_u$} valid if the following conditions hold:
+We consider a trip of \emph{$w^k_u$} valid under the following conditions:
 \begin{align}
@@
 \begin{enumerate*}[label=(\roman*)]
-  \item first, to allow separate trips made by the same vehicle to be differentiated.
-For instance, given two consecutive sightings of \emph{k} three hours apart, we would want to interpret them as belonging to different trips of \emph{k};
+  \item first, to allow distinct trips made by the same vehicle to be differentiated.
+For instance, given two consecutive sightings of \emph{k} three hours apart, we likely want to interpret them as belonging to different trips of \emph{k};
   \item second, it allows implausible trips to be identified.
@@
 Two explanations are common, either one of the cameras made a detection error, or there is another vehicle with a cloned number-plate travelling on the road network.
-\end{enumerate*} Evidently, condition~\ref{e.trip.constraints.2} is only valid for trips of length two sightings or greater.
+\end{enumerate*} Evidently, condition~\ref{e.trip.constraints.2} is only valid for trips of degree two sightings or greater.
 Nevertheless, trips can easily be differentiated by first sorting sightings by time of occurrence, then calculating the journey time sequence for the entire sequence and finally comparing each element against $\Tau$.
 An example of a trip identified this way can be seen in Figure~\ref{fig:trip-example}.
+
+The simplest approach to choosing the value of $\Tau$ is to pick a fixed empirical value, such as 5 or 10 minutes.
+However, if the distance between two cameras is greater than another pair of cameras, then it makes sense that $\Tau$ is relaxed.
+Similarly, if there is an anomaly in the road network, such as a traffic jam, and the routes connecting the two cameras are affected, then the value of $\Tau$ should also be adapted.
+Hence, $\Tau$ should be a function of the distance between the two cameras (or, more accurately, of the top n-routes between these) and the distribution of observed journey times.
+The same rationale can be applied to $\tau$.
 
@@
   \includegraphics[width=1\linewidth]{length-dist.pdf}
-  \caption{Distribution of trips per length of trip.}
+  \caption{Distribution of trips per degree of trip.}
   \label{fig:length-dist}
@@
   \end{subfigure}\hfill
-  \caption{Example of a trip of length 4.
+  \caption{Example of a trip of degree 4.
 On the left side the corresponding data table is shown.
@@
 On the right side, the same trip is plotted on a map.
-However, the lines do not represent the true route taken by the vehicle, but instead the fastest driving route between sightings.
+The circles represent camera locations whereas the lines represent the fastest driving routes between sightings rather than the true route taken by the vehicle.
 Even though no routing information is available for each consecutive pair of sightings, the observed journey times can be compared against the distribution of collective journey times to rank the set of most likely routes chosen (which can be determined for instance by Stochastic User Equilibrium~\cite{Castillo2008}).}%
@@
 
-The simplest approach to choosing the value of $\Tau$ is to pick a fixed empirical value, such as 5 or 10 minutes.
-However, if the distance between two cameras is greater than another pair of cameras, then it makes sense that $\Tau$ is relaxed.
-Similarly, if there is an anomaly in the road network, such as a traffic jam, and the routes connecting the two cameras are affected, then the value of $\Tau$ should also be adapted.
-Hence, $\Tau$ should be a function of the distance between the two cameras (or, more accurately, of the top n-routes between these) and the distribution of observed journey times.
-The same rationale can be applied to $\tau$.
-
 On the other hand, a consistent set of models and methodologies for estimating $\Tau$ and $\tau$, under a variety of circumstances and road anomalies, remains an open research problem.
-Thus, for the purposes of this work, we will consider these parameters to be fixed.
+Thus, for the purposes of this work, we consider these parameters to be fixed.
 Nevertheless, the importance of estimating these parameters should not be understated.
 Any errors in trip identification that occur due to poor estimation and filtering methods will propagate and will be amplified in posterior analysis done using trip data.
-Therefore, $\Tau$ and $\tau$ estimation is something that needs to be carefully addressed and studied in the future in order to fully understand its impact in the outcome of the research.
-To briefly exemplify this, figure~\ref{fig:length-dist} shows how the number of trips per length of trip varies by fixing $\Tau$ at different empirical values ($\tau$ is ignored).
+To briefly exemplify this, figure~\ref{fig:length-dist} shows how the number of trips per degree of trip varies by fixing $\Tau$ at different empirical values ($\tau$ is ignored).
 
-\subsection{Duplicate scannings}
+\subsubsection*{Duplicate scannings}
 
-We need to ensure that every trip of vehicle \emph{k} is unique from all other trips of vehicle \emph{k}. I.e. given $W^k$ the set of all valid trips of $k$:
+We need to ensure that every trip of vehicle \emph{k} is unique from all other trips of vehicle \emph{k}. That is, given $W^k$ the set of all valid trips of $k$:
 
@@
 
+\vspace{-0.5cm}
 \begin{align}
 s^{k}_{(u,i)} \neq s^{k}_{(v,j)}, \forall u,v &= 1, 2, \ldots, N \ , \ u \neq v,  \label{e.trip.history.constraint} \\
-\forall i,j &= 1, 2, \ldots, n^u \ ,\nonumber \\
-\forall i,j &= 1, 2, \ldots, n^v \ , \nonumber
+\forall i &= 1, 2, \ldots, n^u \ , \\
+\forall j &= 1, 2, \ldots, n^v \  \nonumber
 \end{align}
-where $s_{(u,i)}^k$ is the $i^{th}$ sighting of vehicle $k$ during trip $u$, and $n^u$ is the number of sightings in trip $u$.
+where $s_{(u,i)}^k$ is the $i^{th}$ sighting of the $u^{th}$ trip of vehicle $k$, and $n^u$ is the degree of $u$.
 
-However, ANPR can identify the same vehicle multiple times in the same run, if for instance the vehicle is stopped at a junction or traffic light.
+In fact, ANPR cameras can identify the same vehicle multiple times in the same pass, if for instance the vehicle is stopped at a junction or traffic light.
 Hence, if two sightings occurred at the same location in a very short period of time, then there is a strong possibility that these are duplicate observations.
 As a simplification, we can assume that a trip should not contain cycles and that no camera should appear twice in the same trip.
-Yet, this assumption ignores cases where a vehicle is required to correct its route by passing through the same location as one previously observed in the same trip.
-Thus, we affirm that two sightings of vehicle \emph{k} are different if they were observed at two different points in time at different locations, or, if observed at the same location, then the time interval must be greater than a parameter $\gamma$, otherwise the two sightings are deemed as duplicates:
+Yet, this assumption ignores cases where a vehicle is required to correct its route by passing through cameras that have already been registered in that trip.
+Thus, two sightings of vehicle \emph{k} are different if they were observed:
+\begin{enumerate*}[label=(\roman*)]
+  \item at two different points in time at different locations;
+  \item or at the same location with a time interval greater than $\gamma$.
+\end{enumerate*}
+Otherwise the two sightings are deemed as duplicates:
 
+\vspace{-0.5cm}
 \begin{align} \label{e.sighting.different.2}
@@
 Even though a poor estimation of $\gamma$ also has an impact on error propagation, this decreases substantially after filtering duplicates according to the heuristic above, due to the low occurrence of cycles in trips.
-Nevertheless, this issue needs to be fully investigated in future works.
 
-\subsection{Errors in plate scanning}
+\subsubsection*{Errors in plate scanning}
 
@@
 Even so, in this work we do not present a complex solution for this issue.
-Methods to detect and address these two types of errors need to be designed and implemented in the future.

### 2018-02-22 classification.tex
Made some text changes and some re-placement of tables/figures: discussion/errors in plate scanning

@@
 \subsection{Clustering trips}\label{s.classification}
-
-One direct application of trip data is unsupervised learning.
-In this section we identify groups of vehicles with similar trip patterns based on frequency and diversity of travel.
-Clusters can represent private vehicles, such as work-home commuters, transit vehicles like buses and taxis, or other types of vehicles such as delivery trucks.
-%This in turn, may enable trip modes to be estimated
-%However, due to its limited coverage of the road network, it is not clear how well ANPR data can capture much the travel characteristics necessary to accurately describe the different groups.
-
-Due to distinctly different traffic behaviour during weekends, we chose to consider only trips occurring during weekdays.
-We thus used all number-plate data collected between the 6th and 24th of February and exclude data from the 2 weekends in-between.
-Furthermore, as mentioned in section~\ref{s.trips}, we used fixed empirical values for $\Tau$.
-Table~\ref{t:trips-tau} displays the number of trips, average trip degree and the proportion of trips of degree one, for varying values of $\Tau$.
-The effect of incrementing $\Tau$ on resulting trips is clear: increased trip degrees and decreased trips containing just a single sighting.
-On the other hand, we did not set a value for $\tau$, but instead handled implausible trips by filtering all sightings with confidence below 85\%.
-Duplicates were filtered after identifying consecutive sightings by the same camera occurring within the same trip.
-Clock synchronisation errors were provided in milliseconds with none exceeding 5 seconds.
-These were therefore ignored.
-
-\begin{table}[t]
-\centering
-\tabcolsep=0.25cm
-\begin{tabular}{c c c c}
-  \hline
-Tau & Trips & \thead{Average\\Degree} & \thead{Proportion of Trips \\ Degree 1} \\
-  \hline
-5 min & 13,603,759 & 1.46 & 0.70 \\
-7.5 min & 12,394,709 & 1.60 & 0.64 \\
-10 min & 11,690,791 & 1.69 & 0.61 \\
-15 min & 10,823,333 & 1.83 & 0.57 \\
-20 min & 10,305,860 & 1.92 & 0.54 \\
-30 min &  9,653,499 & 2.04 & 0.52 \\
-   \hline
-\end{tabular}
-\caption{Overview of trip data for varying values of $\Tau$.}
-\label{t:trips-tau}
-\vspace{-0.5cm}
-\end{table}
-
-Transforming trips into features which can be used in clustering algorithms was a 3-step process:
-\begin{enumerate*}[label=(\roman*)]
-  \item First, every trip was summarised as a single row of data. The following information was extracted: degree of trip, origin, destination, route, start and end times.
-  \item Second, daily trip information was obtained for each vehicle: number of trips, median of trip degrees, number of sightings, distinct number of origins destinations, and routes, hour of first sighting, hour of last sighting and total rest time between trips.
-  \item Finally, daily information per vehicle was collapsed into a single row by averaging this information across the 15 days.
-\end{enumerate*}
-
-Table~\ref{t:features} depicts a sample of the resulting features vector.
-A total of 1034107 distinct vehicles were detected.
-However, because there is a high percentage of trips containing a single sighting, some of these features were highly correlated.
-We therefore, chose to remove three of the features represented in~\ref{t:features}: \emph{Average Sightings}, \emph{Average Distinct Routes} and \emph{Average Hour Difference}, to avoid the obfuscation of the natural clustering ~\cite{Kmeans}.
-Furthermore we considered that a trip of degree one has no destination (which explains values of average distinct destination below zero) and we filtered all instances of vehicles where the total number of trips is lower than 3, resulting in 642006 unique vehicles.
 
@@
 
+One direct application of trip data is unsupervised learning.
+In this section we identify groups of vehicles with similar trip patterns based on frequency and diversity of travel.
+Clusters can represent private vehicles, such as work-home commuters, transit vehicles like buses and taxis, or other types of vehicles such as delivery trucks.
+%This in turn, may enable trip modes to be estimated
+%However, due to its limited coverage of the road network, it is not clear how well ANPR data can capture much the travel characteristics necessary to accurately describe the different groups.
+
+Due to distinctly different traffic behaviour during weekends, we chose to consider only trips occurring during weekdays.
+Therefore, we used all number plate data collected between the 6th and 24th of February and excluded data from the 2 weekends in-between.
+Furthermore, as mentioned in section~\ref{s.trips}, we used fixed empirical values for $\Tau$.
+Table~\ref{t:trips-tau} displays the number of trips, average trip degree and the proportion of trips of degree one, for varying values of $\Tau$.
+The effect of incrementing $\Tau$ on resulting trips is clear: increased trip degrees and decreased trips containing just a single sighting.
+On the other hand, we did not set a value for $\tau$, but instead handled implausible trips by filtering all sightings with confidence below 85\%.
+Duplicates were filtered after identifying consecutive sightings by the same camera occurring within the same trip.
+Clock synchronisation errors were provided in milliseconds with none exceeding 5 seconds.
+These were hence ignored.
+
+Transforming trips into features which can be used in clustering algorithms was a 3-step process:
+\begin{enumerate*}[label=(\roman*)]
+  \item First, every trip was summarised as a single row of data. The following information was extracted: degree of trip, origin, destination, route, start and end times.
+  \item Second, daily trip information was obtained for each vehicle: number of trips, median of trip degrees, number of sightings, distinct number of origins destinations, and routes, hour of first sighting, hour of last sighting and total rest time between trips.
+  \item Finally, daily information per vehicle was collapsed into a single row by averaging this information across the 15 days.
+\end{enumerate*}
+
+Table~\ref{t:features} depicts a sample of the resulting features vector.
+A total of 1,034,107 distinct vehicles were detected.
+However, because there is a high percentage of trips containing a single sighting, some of these features were highly correlated.
+We therefore, chose to remove three of the features represented in table~\ref{t:features}: \emph{Average Sightings}, \emph{Average Distinct Routes} and \emph{Average Hour Difference}, to avoid the obfuscation of the natural clustering ~\cite{Kmeans}.
+Furthermore we considered that a trip of degree one has no destination (which explains values of average distinct destination below one) and we filtered all instances of vehicles where the total number of trips is lower than 3, resulting in 642,006 unique vehicles.
+
 Clustering of vehicles was performed using the Hartigan and Wong \emph{k-means} algorithm, for each value of $\Tau$.
 The number of clusters \emph{k} was varied between 2 and 8 and executed with a maximum of 200 iterations and 100 different starting states of the algorithm.
-The Calinski-Harabasz criterion is used to pick the best value of $k$, the one that minimises the within-cluster and between-cluster errors and provides the more natural clusters ~\cite{Kmeans}.
-The results of vehicle clustering are presented and discussed in the section below.
+The Calinski-Harabasz criterion is used to determine the best value of $k$, the one that minimises the within-cluster and between-cluster errors and provides the more natural clusters ~\cite{Kmeans}.
+The results of trip clustering are presented and discussed in the next section.
 

### 2018-02-22 discussion.tex
Made some text changes and some re-placement of tables/figures: discussion/errors in plate scanning

--- 2018-02-01 discussion.tex
@@
 In this paper, we have presented a set of definitions and constraints that establish a conceptual foundation for identifying vehicle trips from number plate detections.
-We also identify two parameters, $\tau$ and $\Tau$ as critical in the discrimination of plausible and implausible trips.
+
+We have also identified two parameters, $\tau$ and $\Tau$ as critical in the discrimination of plausible and implausible trips.
 Hence, future work should first and foremost focus on developing formal methods to estimate these parameters from observed distributions of travel times and by applying knowledge about the structure of the road network.
+Any errors in trip identification that occur due to poor estimation and filtering methods will propagate and be amplified in posterior analysis done using trip data.
 Moreover, methods for addressing issues concerning camera performance, namely wrong and duplicate scans, should be further developed and researched.

### 2018-02-22 trips.tex
Made some text changes and some re-placement of tables/figures: discussion/errors in plate scanning

@@
 The same rationale can be applied to $\tau$.
+However, due to time constraints, in this work we simply used several fixed values of $\Tau$ and set $\tau$ to zero.
+Figure~\ref{fig:length-dist} shows how the number of trips per degree of trip varies by fixing $\Tau$ at different empirical values.
 
@@
   \label{fig:trip-example}%
+  \vspace{-0.40cm}
 \end{figure*}
+\vspace{-0.25cm}
 
-On the other hand, a consistent set of models and methodologies for estimating $\Tau$ and $\tau$, under a variety of circumstances and road anomalies, remains an open research problem.
-Thus, for the purposes of this work, we consider these parameters to be fixed.
-Nevertheless, the importance of estimating these parameters should not be understated.
-Any errors in trip identification that occur due to poor estimation and filtering methods will propagate and will be amplified in posterior analysis done using trip data.
-To briefly exemplify this, figure~\ref{fig:length-dist} shows how the number of trips per degree of trip varies by fixing $\Tau$ at different empirical values ($\tau$ is ignored).
-
-\subsubsection*{Duplicate scannings}
+\subsection{Duplicate scannings}
 
@@
 Even though a poor estimation of $\gamma$ also has an impact on error propagation, this decreases substantially after filtering duplicates according to the heuristic above, due to the low occurrence of cycles in trips.
+\vspace{-0.25cm}
 
-\subsubsection*{Errors in plate scanning}
+\subsection{Errors in plate scanning}
 
-Number plate recognition cameras have accuracy rates of 99.9\% or higher.
-If we consider that on average 1 to 10 out of 10000 scans are misclassified number-plates then approximately 200 to 2000 scans everyday are incorrect.
-We categorise the errors made by ANPR cameras into two types:
-\begin{enumerate*}[label=(\roman*)]
-  \item a passing vehicle is not detected by the camera;
-  \item a different vehicle is detected instead.
-\end{enumerate*}
-In the first case, there is no observed data.
-The corresponding trip sequence for the affected vehicle will be missing a sighting but we will have no indication of this.
-It should be noted that this could also happen in the case where a vehicle does not take the `normal' route and pass this intermediary camera.
-For the second case, there is a recorded sighting, but this will be assigned to the incorrect vehicle.
-Even though, the camera provides a value of \emph{confidence} which helps in diagnosing incorrect scans, errors still occur for high values of confidence.
-Even so, in this work we do not present a complex solution for this issue.
+ANPR cameras have average accuracy rates of 99.9\% and higher.
+If we consider that on average 1 to 10 out of 10000 number plate scans are misclassified number-plates then approximately between 200 to 2000 scans everyday are incorrect.
+These errors propagate and lead to innacuracies in the identification of trips.
+In fact, misclassifications affect the trip sequences of two vehicles: the true passing vehicle and the vehicle erroneously detected instead.
+The true vehicle will be missing a sighting in the correspoding trip sequence vector whereas the other vehicle's trip sequence will contain an extra invalid sighting.
+Moreover, ANPR cameras may exceptionally fail to detect passing vehicles.
+Even though only the passing vehicle is affected in this case, it's trip sequence vector is nonetheless affected.
+Therefore, due to their impact in trip identification, it is important to detect and address missing and misclassified scans.
+
+We filer invalid sightings by removing all sightings containing a \emph{confidence} value below $85\%$. However, we do not address missing scans in this work.
+
+\begin{table}[!ht]
+\centering
+\tabcolsep=0.25cm
+\begin{tabular}{c c c c}
+  \hline
+Tau & Trips & \thead{Average\\Degree} & \thead{Proportion of Trips \\ Degree 1} \\
+  \hline
+5 min & 13,603,759 & 1.46 & 0.70 \\
+7.5 min & 12,394,709 & 1.60 & 0.64 \\
+10 min & 11,690,791 & 1.69 & 0.61 \\
+15 min & 10,823,333 & 1.83 & 0.57 \\
+20 min & 10,305,860 & 1.92 & 0.54 \\
+30 min &  9,653,499 & 2.04 & 0.52 \\
+   \hline
+\end{tabular}
+\caption{Overview of trip data for varying values of $\Tau$.}
+\label{t:trips-tau}
+\vspace{-0.5cm}
+\end{table}

### 2018-02-23 classification.tex
A few changes - submitted to supervisor for review

@@
 % !TEX root = paper.tex
-\subsection{Clustering trips}\label{s.classification}
+\subsection{Clustering vehicles}\label{s.classification}
 
@@
 
-One direct application of trip data is unsupervised learning.
-In this section we identify groups of vehicles with similar trip patterns based on frequency and diversity of travel.
-Clusters can represent private vehicles, such as work-home commuters, transit vehicles like buses and taxis, or other types of vehicles such as delivery trucks.
-%This in turn, may enable trip modes to be estimated
-%However, due to its limited coverage of the road network, it is not clear how well ANPR data can capture much the travel characteristics necessary to accurately describe the different groups.
+One application of trip data is to group vehicles together according to similar trip patterns, namely frequency and diversity of travel.
+We can distinguish vehicles based on private and transit modes of transportation and classes within these, such as everyday commuters versus sporadic drivers, buses versus taxis, and other types of vehicles such as lorries and delivery trucks.
+Ideally, we would like to build a model, through supervised learning, that is able to classify a vehicle from unseen trip data.
+However, due to privacy laws, there is no publicly available database that maps a plate number into one or several of the categories above.
+Therefore, we've decided to group vehicles by performing unsupervised learning using the \emph{k-means} clustering algorithm.
+Despite not being able to perform a rigorous validation or interpretation of the resulting clusters, we can begin to understand how useful trip data is to describe vehicles' travel patterns, given the limited coverage of the cameras in the road network.
 
-Due to distinctly different traffic behaviour during weekends, we chose to consider only trips occurring during weekdays.
+Due to distinctly different traffic behaviour during weekends, we only consider trips occurring during weekdays.
 Therefore, we used all number plate data collected between the 6th and 24th of February and excluded data from the 2 weekends in-between.
@@
 It is noteworthy that the best value of $k$ increases inversely to $\Tau$.
-Although a higher value of $k$ seems to suggest that trips with smaller values of $\Tau$ are able capture the variance in the data better, we have to consider that varying $\Tau$ affects the average trip degree in the same direction whilst affecting the total number of trips in the opposite direction (table~\ref{t:trips-tau}).
+Although a higher value of $k$ seems to suggest that trips with smaller values of $\Tau$ are better able to capture the variance of trip data, we have to consider that varying $\Tau$ affects the average trip degree in the same direction whilst affecting the total number of trips in the opposite direction (table~\ref{t:trips-tau}).
 
-Tables~\ref{t:kmeans_centers_450} and~\ref{t:kmeans_centers_1200} depict the cluster centres for $\Tau$ equal to 7.5 minutes and 20 minutes respectively.
-These partially meet our expectations.
+%This in turn, may enable trip modes to be estimated
+%However, due to its limited coverage of the road network, it is not clear how well ANPR data can capture much the travel characteristics necessary to accurately describe the different groups.
+
+Table~\ref{t:kmeans_centers_300} depicts the cluster centres for the combination of $\Tau$ and $k$ that maximises the Calinski-Harabasz criterion.
+These results partially meet our expectations.
 For instance, we were expecting to find a relatively small cluster representing taxis with a high average number of trips per day, occurring over a variety of origins and destinations and over a large time frame.
-Clusters 2 and 7 for $\Tau = 7.5$ and cluster 4 for $\Tau = 20$ do indeed fit this profile.
-What differentiates cluster 2 from 7 in the first case is essentially a lower mean number of trips per day and smaller average time at which the first and last trips of the day occur.
-Additionally, buses to some extent fit this category as well, however we expected that buses could be separated from taxis by showing less diversity in the number of origins and destinations as these essentially do multiple runs of the same trip throughout the day.
+Clusters 5, 6 and 7 do indeed fit this profile.
+What differentiates cluster 5 from 6 is the particularly high number of average trips per day.
+When $k$ is equal to 7 these two clusters merge into one.
+To some extent, other categories fit this profile as well, namely buses, lorries and delivery trucks.
+However, we could expect buses to show less diversity in the number of origins and destination as these essentially do multiple runs of the same route throughout the day.
 Still, this can be explained by the fact that buses take routes through main and secondary roads.
 As most cameras are placed in main roads, the one long bus trip can be perceived as multiple small trips as the bus alternates between arterial and main roads.
-This is in fact, one of the downsides of ANPR data and is something that the methodology presented in section~\ref{s.trips} needs to develop in the future.
+This is in fact, one of the downsides of ANPR data and the methodology presented in section~\ref{s.trips}.
 
@@
 However, we observe at least one big group of trips occurring mostly during lunch hour.
-Several interpretations are possible but further work is needed to provide a more consistent interpretation and validation of these results.
 A big contributing factor however is the fact that a high proportion of trips contains only a single sighting (table~\ref{t:trips-tau}).
-It may be the case that commuters choose other routes than those going through ANPR cameras.
-Still, we need to devise further models and methods that are able to capture this uncertainty.
+It may be the case that many of these drivers choose routes other than those going through ANPR cameras.
+
+To improve the interpretability of these results, we've gain access to a database that outputs the vehicle's weight, wheelplane, make and category from the corresponding plate number.
+Broadly speaking, vehicle categories are defined relative to the transport of persons or goods, and sub-categories are given according to maximum allowed mass.
+With this information we sorted vehicles in the ANPR database into one of the following classes: \textit{Bus, Car, Motorcycle, Truck, Van}.
+Unfortunately, we were unable to identify taxis and therefore were unable to separate private cars from taxis.
+Table~\ref{t:} shows the mean relative frequency of these classes, and the corresponding deviation for the clusters shown in table~\ref{t:kmeans_centers_300}.
+Only the clusters showing more notable deviations from the mean are shown.
+Cluster 5, in particular is mainly composed by trucks.
+Cluster 6, sees an increase of frequency of cars which may represent taxis.
+Vans are seen in larger proportions in Cluster 7 and 3.
+These results provide some evidence in favor of some of our assumptions above and demonstrate that whilst ANPR data is able to capture some of the travel patterns, a clear and robust assessment of travel patterns is still not possible.
+Furthermore, it is not clear whether this is due to limited coverage of the road network or to a lack of methods that model, quantify and address the uncertainty of trips identified from ANPR data.
+
+%Several interpretations are possible but further work is needed to provide a more consistent interpretation and validation of these results.
+The lack of baselines and validation data is hence one of the shortcomings of this study.
+An additional pitfall of this methodology is that it is unable to evaluate the reliability of extracted trips.
+
+%Nevertheless, trip data derived from ANPR number plate data is able to capture . This opens up room for the development of further methods that model and quantify this uncertainty.
 
@@
   \hline
-1 & 157962 & 2.45 & 2.27 & 11.38 & 14.93 \\
-  2 & 303513 & 2.18 & 2.09 & 12.53 & 14.47 \\
-  3 & 21549 & 6.02 & 5.00 & 8.68 & 17.23 \\
-  4 & 108094 & 2.99 & 2.73 & 9.99 & 15.86 \\
+1 & 157,962 & 2.45 & 2.27 & 11.38 & 14.93 \\
+  2 & 303,513 & 2.18 & 2.09 & 12.53 & 14.47 \\
+  3 & 21,549 & 6.02 & 5.00 & 8.68 & 17.23 \\
+  4 & 108,094 & 2.99 & 2.73 & 9.99 & 15.86 \\
   5 & 509 & 33.62 & 10.42 & 5.81 & 19.67 \\
-  6 & 1993 & 17.99 & 11.24 & 6.45 & 19.09 \\
-  7 & 4971 & 10.41 & 7.64 & 7.97 & 18.04 \\
-  8 & 58059 & 4.09 & 3.62 & 9.22 & 16.60 \\
+  6 & 1,993 & 17.99 & 11.24 & 6.45 & 19.09 \\
+  7 & 4,971 & 10.41 & 7.64 & 7.97 & 18.04 \\
+  8 & 58,059 & 4.09 & 3.62 & 9.22 & 16.60 \\
    \hline

### 2018-02-23 discussion.tex
A few changes - submitted to supervisor for review

@@
 However, the results presented here could benefit from extra work and further validation.
+Namely, gaining access to a database that maps plate numbers to vehicle types would not only provide means to better validate the proposed methodology but also enable the application of supervised learning instead.
+Furthermore, this work can be improved upon by using other data sources as a baseline for evaluating the reliability of extracted trips, such as GPS taxi traces.
 Finally, future work can focus on using trip data to solve interesting research problems such as:
@@
 \end{enumerate*}
+
+\subsection*{Acknowledgments}
+
+The authors would like to thank Phil Blythe, Professor of Intelligent Transport Systems at Newcastle University, and Ray King from the Tyne and Wear Urban Traffic and Management Centre for providing guidance and access to the Automatic Number Plate Recognition database.

### 2018-02-23 paper.tex
A few corrections and changes. Changed the name of length to degree, removed most references to future work. Rephrased a few sentences.
A few changes - submitted to supervisor for review

@@
 
-\title{\Large Clustering Trips Identified from Automatic Number \\ Plate Recognition Camera Scans}
+\title{\Large Clustering Vehicles based on Trips Identified from Automatic \\ Number Plate Recognition Camera Scans}
 \author{Pedro M. Pinto Silva \thanks{School of Computing, Newcastle University, United Kingdom}
@@
 In the Tyne and Wear region (UK) the network of ANPR cameras is used to monitor travel times across sensitive roads.
-However, few works have explored the full potential of number-plate scans for analysing individual and collective travel patterns.
+So far, few works have explored the full potential of number-plate scans for analysing individual and collective travel patterns.
 In this work we present a methodology for deriving trips from vehicle sightings at fixed camera locations.

### 2018-02-23 dataset.tex
Made some text changes and some re-placement of tables/figures: discussion/errors in plate scanning
Latest changes with suggested fixes by supervisors.

@@
 
-\begin{table}[!ht]
+\begin{figure}[t]
 \centering
-\tabcolsep=0.17cm
-\small
-\begin{tabular}{c c c c c}
-  \hline
-  Vehicle & Camera & Timestamp & \thead{Clock \\Error} & Confidence \\
-  \hline
-  169239 & 1031 & 1454284800.26 &   0 & 100 \\
-  12862943 & 18 & 1454284800.97 &   8 &  61 \\
-  16243894 & 22 & 1454284801.46 &   6 &  86 \\
-  4817789 & 52 & 1454284803.43 &  13 &  94 \\
-  5503486 & 110 & 1454284802.19 &  22 &  91 \\
-  15244177 & 115 & 1454284802.83 &  18 &  87 \\
-   \hline
-\end{tabular}
-\caption{Sample of number plate data. Clock error is given in milliseconds and confidence as a percentile value.}
-\label{t:np_data_example}
-%\vspace{-2cm}
-\end{table}
+\includegraphics[width=.90\linewidth]{ANPR-overview.pdf}
+\caption{Overview of a ANPR-based system for traffic monitoring and control.}
+\label{fig:anpr-overview}
+\vspace{-0.25cm}
+\end{figure}
 
-Automatic num\-ber plate recognition (ANPR) cameras are actively employed in urban traffic environments and play an important role in day-to-day intelligent transportation systems.
-They can be used by government subsidised entities in urban traffic management and control; by commissioned highway agencies in electronic toll collection; or by law enforcement organisations in detecting speeding vehicles and validating number plate registrations.
-The wide diversity of applications, paired with the large improvements in price-to-performance ratios of ANPR hardware and software systems, has resulted in increased investments of ANPR cameras for urban environments~\cite{EvolutionUTMC2013, SurveyITS2011}.
-
-\begin{figure*}[t]
+\begin{figure*}[!ht]
 \centering
@@
 \label{fig:time-series}
+\vspace{-0.4cm}
 \end{figure*}
+
+Automatic num\-ber plate recognition (ANPR) cameras are actively employed in urban traffic environments and play an important role in day-to-day intelligent transportation systems.
+They can be used by government subsidised entities in urban traffic management and control; by commissioned highway agencies in electronic toll collection; or by law enforcement organisations in detecting speeding vehicles and validating number plate registrations.
+The wide diversity of applications, paired with the large improvements in price-to-performance ratios of ANPR hardware and software systems, has resulted in increased investments of ANPR cameras for urban environments~\cite{EvolutionUTMC2013, SurveyITS2011}.
 
@@
 Figure~\ref{fig:time-series} shows the number of daily scans recorded over a month (February, 2017).
-Furthermore, every scan is stored in a central database managed by Urban Traffic Management and Control (UTMC) Tyne and Wear, and used to compute travel times across particular links of interest in the road network.
+Furthermore, every scan is stored in a central database managed by the Tyne and Wear UTMC, and used to compute travel times across particular links of interest in the road network.
 These are usually major roads that see high volumes of traffic, or road segments more prone to traffic jams.
@@
 Number plate data, in its essence, is a stream of events, each representing a vehicle observed by one camera at a specific point in time.
-An excerpt of the data can be found in table~\ref{t:np_data_example}.
+An excerpt of the data can be found in Table~\ref{t:np_data_example}.
 All number plate were anonymised by the UTMC through a hashing algorithm before the data was shared.
@@
 Hence, we chose to ignore the latter in this work.
+
+\begin{table}[t]
+\centering
+\tabcolsep=0.12cm
+\begin{tabular}{c c c c c}
+  \hline
+  Vehicle & Camera & Timestamp & \thead{Clock \\Error} & Confidence \\
+  \hline
+  169239 & 1031 & 1454284800.26 &   0 & 100 \\
+  12862943 & 18 & 1454284800.97 &   8 &  61 \\
+  16243894 & 22 & 1454284801.46 &   6 &  86 \\
+  4817789 & 52 & 1454284803.43 &  13 &  94 \\
+  5503486 & 110 & 1454284802.19 &  22 &  91 \\
+  15244177 & 115 & 1454284802.83 &  18 &  87 \\
+   \hline
+\end{tabular}
+\caption{Sample of number plate data. Clock error is given in milliseconds and confidence as a percentile value.}
+\label{t:np_data_example}
+\vspace{-0.3cm}
+\end{table}

### 2018-02-23 discussion.tex
Latest changes with suggested fixes by supervisors.

@@
 Most urban cities in the world employ a network of ANPR cameras that are used for law enforcement as well as traffic monitoring and control.
-Number plate data collected in the Tyne and Wear area is stored and leveraged by the UTMC for computing average journey times across a selection of sensitive roads.
+Number plate data collected in the Tyne and Wear area is stored and leveraged by the Urban Traffic and Management Centre (UTMC) for computing average journey times across a selection of sensitive roads.
 However, number plate data could be used more extensively to identify and study individual and collective travel patterns.
@@
 Furthermore, this work can be improved upon by using other data sources as a baseline for evaluating the reliability of extracted trips, such as GPS taxi traces.
+
 Finally, future work can focus on using trip data to solve interesting research problems such as:
@@
 
+\vspace{-0.3cm}
 \subsection*{Acknowledgments}

### 2018-02-23 paper.tex
Latest changes with suggested fixes by supervisors.

@@
 \usepackage{microtype}
+\usepackage{enumitem}
 
@@
 \title{\Large Clustering Vehicles based on Trips Identified from Automatic \\ Number Plate Recognition Camera Scans}
-\author{Pedro M. Pinto Silva \thanks{School of Computing, Newcastle University, United Kingdom}
+\author{Pedro M. Pinto Silva \thanks{School of Computing, Newcastle University, United Kingdom\break \{p.pinto-da-silva2, matthew.forshaw, stephen.mcgough\}@newcas\-tle.ac.uk}
 \and
@@
 % Default Copyright Statement
-\fancyfoot[R]{\footnotesize{\textbf{Copyright \textcopyright\ 2018 by BigTraffic\\
+\fancyfoot[R]{\footnotesize{\textbf{Copyright \textcopyright\ 2018\\
 Unauthorized reproduction of this article is prohibited}}}
@@
 In this work we present a methodology for deriving trips from vehicle sightings at fixed camera locations.
-We illustrate the effect of parameters $\tau$ and $\Tau$ on trip discrimination and on the detection of implausible trips.
-To demonstrate the potential of trip data we apply \emph{k-means} clustering to trips identified from over 40 million plate scans recorded over fifteen weekdays.
+We identify two parameters $\tau$ and $\Tau$ as essential for identifying implausible trips and differentiating between multiple trips of the same vehicle.
+To demonstrate the applicability of trip data we apply \emph{k-means} clustering to trips identified from over 40 million plate scans recorded over fifteen weekdays.
 Results show that whilst private and transit travel modes can begin to be inferred from the resulting clusters, further work needs to be put into developing a more consistent and integrated framework for trip identification in ANPR data.\end{abstract}
@@
 \bibitem{ODMobileData}
-Alexander, Lauren, et al. "Origin–-destination trips by purpose and time of day inferred from mobile phone data." Transportation research part c: emerging technologies 58 (2015): 240-250.
+Alexander, L., Jiang, S., Murga, M., \& González, M.
+C.
+"Origin–-destination trips by purpose and time of day inferred from mobile phone data." Transportation research part c: emerging technologies 58 (2015): 240-250.
 
 \bibitem{Castillo2010}
-Castillo, Enrique, et al. {\em Optimal use of plate-scanning resources for route flow estimation in traffic networks}. IEEE Transactions on Intelligent Transportation Systems 11.2 (2010): 380-391.
+Castillo, E.
+Gallego, I., Menéndez, J.
+M., \& Rivas, A {\em Optimal use of plate-scanning resources for route flow estimation in traffic networks}.
+IEEE Transactions on Intelligent Transportation Systems 11.2 (2010): 380-391.
 
 \bibitem{Castillo2008}
-Castillo, Enrique, José María Menéndez, and Pilar Jiménez.
+Castillo, E., Menéndez, J.
+M., \& Jiménez, P.
 {\em Trip matrix and path flow reconstruction and estimation based on plate scanning and link observations}.
@@
 \bibitem{Clustering}
-Chen, Huiyu, Chao Yang, and Xiangdong Xu. {\em Clustering Vehicle Temporal and Spatial Travel Behavior Using License Plate Recognition Data}. Journal of Advanced Transportation 2017 (2017).
+Chen, H., Yang, C., \& Xu, X. {\em Clustering Vehicle Temporal and Spatial Travel Behavior Using License Plate Recognition Data}. Journal of Advanced Transportation 2017 (2017).
 
@@
 \bibitem{EvolutionUTMC2013}
-Hamilton, Andrew, et al. {\em The evolution of urban traffic control: changing policy and technology}. Transportation planning and technology 36.1 (2013): 24-43.
+Hamilton, Andrew, Waterson, B., Cherrett, T., Robinson, A., \& Snell, I.
+{\em The evolution of urban traffic control: changing policy and technology}.
+Transportation planning and technology 36.1 (2013): 24-43.
 
@@
 \bibitem{Hazelton2012}
-Parry, Katharina, and Martin L.
-Hazelton.
-{\em Estimation of origin-–destination matrices from link counts and sporadic routing data}.
-Transportation Research Part B: Methodological 46.1 (2012): 175-188.
+Parry, K., \& Hazelton, M. L. {\em Estimation of origin-–destination matrices from link counts and sporadic routing data}. Transportation Research Part B: Methodological 46.1 (2012): 175-188.
 
 \bibitem{ClusteringGPS}
-Schüssler, Nadine, and Kay W.
-Axhausen.
-{\em Identifying trips and activities and their characteristics from GPS raw data without further information}.
-Arbeitsbericht Verkehrs-und Raumplanung 502 (2008).
+Schüssler, N., \& Axhausen, K. W. {\em Identifying trips and activities and their characteristics from GPS raw data without further information}. Arbeitsbericht Verkehrs-und Raumplanung 502 (2008).
 
 \bibitem{Kmeans}
-Steinley, Douglas, and Michael J. Brusco. {\em Choosing the number of clusters in Κ-means clustering}. Psychological Methods 16.3 (2011): 285.
+Steinley, D., \& Brusco, M. J. {\em Choosing the number of clusters in Κ-means clustering}. Psychological Methods 16.3 (2011): 285.
 
 \bibitem{SurveyITS2011}
-Zhang, Junping, et al. {\em Data-driven intelligent transportation systems: A survey}. IEEE Transactions on Intelligent Transportation Systems 12.4 (2011): 1624-1639.
+Zhang, J., Wang, F.
+Y., Wang, K., Lin, W.
+H., Xu, X., \& Chen, C.
+{\em Data-driven intelligent transportation systems: A survey}.
+IEEE Transactions on Intelligent Transportation Systems 12.4 (2011): 1624-1639.
 

### 2018-02-23 trips.tex
A few changes - submitted to supervisor for review
Latest changes with suggested fixes by supervisors.

@@
 % !TEX root = paper.tex
-\subsection{Trip Identification}\label{s.trips}
+\subsection{Trip identification}\label{s.trips}
 
@@
 n &\ge 1 , \label{e.trip.constraints.1} \\
-\tau_{(i)} &< jt^{k}_{u(i)} < \Tau_{(i)} \ , \ \forall i \in jt^{k}_{u(i)} \label{e.trip.constraints.2},
+\tau &< jt^{k}_{u(i)} < \Tau \ , \ \forall i \in \{0, 1 ,\ldots, n-1\} \label{e.trip.constraints.2},
 \end{align}
-where $\tau_{(i)}$ and $\Tau_{(i)}$ are the lower and upper bound of the $i_{th}$ element of the journey time sequence.
+where $\tau$ and $\Tau$ are the lower and upper bound of the $i_{th}$ element of the journey time sequence.
 
-The first condition~\ref{e.trip.constraints.1} is straightforward and specifies that every identified trip should have at least one sighting.
+\textit{Condition~\ref{e.trip.constraints.1}} is straightforward and specifies that every identified trip should have at least one sighting.
 Obviously, vehicles can make trips that do not pass through any ANPR cameras and thus have no associated sightings: $n = 0$.
 However, this work focus on trips that we can observe and hence we consider that $n > 0$.
-The second condition~\ref{e.trip.constraints.2} defines a minimum and maximum travel times between consecutive observations.
-Its purpose is twofold:
+
+\textit{Condition~\ref{e.trip.constraints.2}} defines a minimum and maximum travel times between consecutive observations. Its purpose is twofold:
 \begin{enumerate*}[label=(\roman*)]
   \item first, to allow distinct trips made by the same vehicle to be differentiated.
-For instance, given two consecutive sightings of \emph{k} three hours apart, we likely want to interpret them as belonging to different trips of \emph{k};
+For instance, given two consecutive sightings of \emph{k} three hours apart, we are likely to want to interpret them as belonging to different trips of \emph{k};
   \item second, it allows implausible trips to be identified.
@@
 Two explanations are common, either one of the cameras made a detection error, or there is another vehicle with a cloned number-plate travelling on the road network.
-\end{enumerate*} Evidently, condition~\ref{e.trip.constraints.2} is only valid for trips of degree two sightings or greater.
+\end{enumerate*} Evidently, Condition~\ref{e.trip.constraints.2} is only valid for trips of degree two sightings or greater.
 Nevertheless, trips can easily be differentiated by first sorting sightings by time of occurrence, then calculating the journey time sequence for the entire sequence and finally comparing each element against $\Tau$.
@@
 The same rationale can be applied to $\tau$.
-However, due to time constraints, in this work we simply used several fixed values of $\Tau$ and set $\tau$ to zero.
+However, we focus on $\Tau$ and leave $\tau$ for analysis at a later stage.
 Figure~\ref{fig:length-dist} shows how the number of trips per degree of trip varies by fixing $\Tau$ at different empirical values.
@@
 \end{figure*}
-\vspace{-0.25cm}
 
-\subsection{Duplicate scannings}
+\subsection{Duplicate scannings}\label{s.duplicates}
 
@@
 Even though a poor estimation of $\gamma$ also has an impact on error propagation, this decreases substantially after filtering duplicates according to the heuristic above, due to the low occurrence of cycles in trips.
-\vspace{-0.25cm}
 
-\subsection{Errors in plate scanning}
+\subsection{Errors in plate scanning}\label{s.errors}
 
-ANPR cameras have average accuracy rates of 99.9\% and higher.
+ANPR cameras have an average accuracy rate of 99.9\% or higher.
 If we consider that on average 1 to 10 out of 10000 number plate scans are misclassified number-plates then approximately between 200 to 2000 scans everyday are incorrect.
@@
 The true vehicle will be missing a sighting in the correspoding trip sequence vector whereas the other vehicle's trip sequence will contain an extra invalid sighting.
+The later may be more easily detected and removed than the former as it may generate a sighting for a vehicle that is normally seen in a different part of the country.
+
+\begin{table*}[t]
+\centering
+\begin{tabular}{c c c c c c c c c c c}
+  \hline
+ \thead{Total\\Trips} & \thead{Average\\Trips} & \thead{Average\\Degree} & \thead{Average\\Sightings} & \thead{Average\\Distinct\\Origins} & \thead{Average\\Distinct\\Destinations} & \thead{Average\\Distinct\\Routes} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} & \thead{Average\\Hour\\Difference} & \thead{Average\\Rest\\Time} \\
+  \hline
+41 & 3.42 & 1.25 & 5.75 & 2.75 & 1.00 & 3.00 & 15.33 & 19.25 & 3.80 & 3.70 \\
+3 & 1.50 & 1.00 & 1.50 & 1.50 & 0.00 & 1.50 & 14.00 & 14.50 & 0.73 & 0.73 \\
+7 & 2.33 & 1.33 & 3.33 & 2.33 & 0.67 & 2.33 & 11.00 & 13.00 & 2.44 & 2.41 \\
+12 & 2.40 & 1.10 & 3.60 & 2.40 & 0.60 & 2.40 & 14.40 & 16.40 & 2.00 & 1.95 \\
+   \hline
+\end{tabular}
+\caption{Sample of extracted features from trips taken from 15 weekdays of number plate data.}
+\vspace{-0.2cm}
+\label{t:features}
+\end{table*}
+
 Moreover, ANPR cameras may exceptionally fail to detect passing vehicles.
@@
 Therefore, due to their impact in trip identification, it is important to detect and address missing and misclassified scans.
-
-We filer invalid sightings by removing all sightings containing a \emph{confidence} value below $85\%$. However, we do not address missing scans in this work.
-
-\begin{table}[!ht]
-\centering
-\tabcolsep=0.25cm
-\begin{tabular}{c c c c}
-  \hline
-Tau & Trips & \thead{Average\\Degree} & \thead{Proportion of Trips \\ Degree 1} \\
-  \hline
-5 min & 13,603,759 & 1.46 & 0.70 \\
-7.5 min & 12,394,709 & 1.60 & 0.64 \\
-10 min & 11,690,791 & 1.69 & 0.61 \\
-15 min & 10,823,333 & 1.83 & 0.57 \\
-20 min & 10,305,860 & 1.92 & 0.54 \\
-30 min &  9,653,499 & 2.04 & 0.52 \\
-   \hline
-\end{tabular}
-\caption{Overview of trip data for varying values of $\Tau$.}
-\label{t:trips-tau}
-\vspace{-0.5cm}
-\end{table}
+We filter invalid sightings by removing all sightings containing a \emph{confidence} value below $85\%$.
+However, we do not address missing scans in this work.

### 2018-02-26 classification.tex

@@
 \subsection{Clustering vehicles}\label{s.classification}
-
-\begin{table*}[t]
-\centering
-\begin{tabular}{c c c c c c c c c c c}
-  \hline
- \thead{Total\\Trips} & \thead{Average\\Trips} & \thead{Average\\Degree} & \thead{Average\\Sightings} & \thead{Average\\Distinct\\Origins} & \thead{Average\\Distinct\\Destinations} & \thead{Average\\Distinct\\Routes} & \thead{Average\\First\\Hour} & \thead{Average\\Last\\Hour} & \thead{Average\\Hour\\Difference} & \thead{Average\\Rest\\Time} \\
-  \hline
-41 & 3.42 & 1.25 & 5.75 & 2.75 & 1.00 & 3.00 & 15.33 & 19.25 & 3.80 & 3.70 \\
-3 & 1.50 & 1.00 & 1.50 & 1.50 & 0.00 & 1.50 & 14.00 & 14.50 & 0.73 & 0.73 \\
-7 & 2.33 & 1.33 & 3.33 & 2.33 & 0.67 & 2.33 & 11.00 & 13.00 & 2.44 & 2.41 \\
-12 & 2.40 & 1.10 & 3.60 & 2.40 & 0.60 & 2.40 & 14.40 & 16.40 & 2.00 & 1.95 \\
-   \hline
-\end{tabular}
-\caption{Sample of extracted features from trips taken from 15 weekdays of number plate data.}
-\vspace{-0.2cm}
-\label{t:features}
-\end{table*}
 
@@
 The effect of incrementing $\Tau$ on resulting trips is clear: increased trip degrees and decreased trips containing just a single sighting.
-On the other hand, we did not set a value for $\tau$, but instead handled implausible trips by filtering all sightings with confidence below 85\%.
+On the other hand, we did not set a value for $\tau$, but instead handled implausible trips by filtering out all sightings with confidence below 85\%.
 Duplicates were filtered after identifying consecutive sightings by the same camera occurring within the same trip.
@@
 These were hence ignored.
+
+\begin{table}[t]
+\centering
+\tabcolsep=0.25cm
+\begin{tabular}{c c c c}
+  \hline
+$\Tau$ & Trips & \thead{Average\\Degree} & \thead{Proportion of Trips \\ Degree 1} \\
+  \hline
+5 min & 13,603,759 & 1.46 & 0.70 \\
+7.5 min & 12,394,709 & 1.60 & 0.64 \\
+10 min & 11,690,791 & 1.69 & 0.61 \\
+15 min & 10,823,333 & 1.83 & 0.57 \\
+20 min & 10,305,860 & 1.92 & 0.54 \\
+30 min &  9,653,499 & 2.04 & 0.52 \\
+   \hline
+\end{tabular}
+\caption{Overview of trip data for varying values of $\Tau$.}
+\label{t:trips-tau}
+\vspace{-0.5cm}
+\end{table}
 
@@
 However, because there is a high percentage of trips containing a single sighting, some of these features were highly correlated.
-We therefore, chose to remove three of the features represented in table~\ref{t:features}: \emph{Average Sightings}, \emph{Average Distinct Routes} and \emph{Average Hour Difference}, to avoid the obfuscation of the natural clustering ~\cite{Kmeans}.
+We therefore, chose to remove three of the features represented in Table~\ref{t:features}: \emph{Average Sightings}, \emph{Average Distinct Routes} and \emph{Average Hour Difference}, to avoid the obfuscation of the natural clustering ~\cite{Kmeans}.
 Furthermore we considered that a trip of degree one has no destination (which explains values of average distinct destination below one) and we filtered all instances of vehicles where the total number of trips is lower than 3, resulting in 642,006 unique vehicles.
@@
 The number of clusters \emph{k} was varied between 2 and 8 and executed with a maximum of 200 iterations and 100 different starting states of the algorithm.
-The Calinski-Harabasz criterion is used to determine the best value of $k$, the one that minimises the within-cluster and between-cluster errors and provides the more natural clusters ~\cite{Kmeans}.
+The Calinski-Harabasz criterion is used to determine the best value of $k$.
+It minimises and maximises the within-cluster and between-cluster sum of squares respectively, resulting in more natural clusters ~\cite{Kmeans}.
 The results of trip clustering are presented and discussed in the next section.
@@
 It is noteworthy that the best value of $k$ increases inversely to $\Tau$.
-Although a higher value of $k$ seems to suggest that trips with smaller values of $\Tau$ are better able to capture the variance of trip data, we have to consider that varying $\Tau$ affects the average trip degree in the same direction whilst affecting the total number of trips in the opposite direction (table~\ref{t:trips-tau}).
+Although a higher value of $k$ seems to suggest that trips with smaller values of $\Tau$ are better able to capture the variance of trip data, we have to consider that varying $\Tau$ affects the average trip degree in the same direction whilst affecting the total number of trips in the opposite direction (Table~\ref{t:trips-tau}).
 
@@
 However, we observe at least one big group of trips occurring mostly during lunch hour.
-A big contributing factor however is the fact that a high proportion of trips contains only a single sighting (table~\ref{t:trips-tau}).
+A big contributing factor however is the fact that a high proportion of trips contains only a single sighting (Table~\ref{t:trips-tau}).
 It may be the case that many of these drivers choose routes other than those going through ANPR cameras.
 
-To improve the interpretability of these results, we've gain access to a database that outputs the vehicle's weight, wheelplane, make and category from the corresponding plate number.
+To improve the interpretability of these results, we would like to gain access to a governmental database that provides vehicles' weight, wheelplane, make and category from the corresponding plate number.
 Broadly speaking, vehicle categories are defined relative to the transport of persons or goods, and sub-categories are given according to maximum allowed mass.
-With this information we sorted vehicles in the ANPR database into one of the following classes: \textit{Bus, Car, Motorcycle, Truck, Van}.
-Unfortunately, we were unable to identify taxis and therefore were unable to separate private cars from taxis.
-Table~\ref{t:} shows the mean relative frequency of these classes, and the corresponding deviation for the clusters shown in table~\ref{t:kmeans_centers_300}.
-Only the clusters showing more notable deviations from the mean are shown.
-Cluster 5, in particular is mainly composed by trucks.
-Cluster 6, sees an increase of frequency of cars which may represent taxis.
-Vans are seen in larger proportions in Cluster 7 and 3.
-These results provide some evidence in favor of some of our assumptions above and demonstrate that whilst ANPR data is able to capture some of the travel patterns, a clear and robust assessment of travel patterns is still not possible.
-Furthermore, it is not clear whether this is due to limited coverage of the road network or to a lack of methods that model, quantify and address the uncertainty of trips identified from ANPR data.
+With this information it would be possible to sort vehicles in the ANPR database into one of the following classes: \textit{Bus, Car, Motorcycle, Truck, Van}.
+Even though taxis cannot be identified in this way, taxi companies could be contacted directly to obtain this data.
+However, such databases are available for law enforcement, but not to the public for ethical and privacy reasons.
+Additionally, this would require us to de-anonymize the hashed number plates.
+Nevertheless, these results demonstrate that whilst ANPR data is able to capture some of the travel patterns, a clearer and more robust assessment of travel patterns is needed.
+Furthermore, it is not clear whether this is due to limited coverage of the ANPR cameras in the road network or of the methodology used.
 
@@
   \hline
-Tau & Best $k$ & Betweenss & \thead{Average\\Withinss} & \thead{Calinski-\\Harabasz} \\
+$\Tau$ & Best $k$ & Betweenss & \thead{Average\\Withinss} & \thead{Calinski-\\Harabasz} \\
   \hline
-5 min &   8 & 0.923 & 0.125 & 1,116,962 \\
+5 min &   8 & 0.923 & 0.125 & \textbf{1,116,962} \\
   7.5 min &   7 & 0.904 & 0.143 & 1,003,744 \\

### 2018-02-26 intro.tex
Made some text changes and some re-placement of tables/figures: discussion/errors in plate scanning

@@
 However, this has numerous ethical and technical issues.
-From an ethical standpoint should we be allowed to know where all vehicles are at any given point in time.
+From an ethical standpoint should we be allowed to know where all vehicles are at any given point in time?
 From a technical point of view, although every vehicle could be fitted with a GPS tracker -- costly in its own right -- there would still exist the issue of how we would collect and stream all of this data for future processing.
-Alternatively one can view the problem the other way around and rather than tracking individual vehicles look at collecting information by observing vehicles passing points within the road networks.
+Alternatively one can view the problem the other way around and rather than tracking individual vehicles look at collecting information by observing vehicles passing points within the road network.
 A prime example of this approach are Automatic Number Plate Recognition (ANPR) cameras.
@@
 
-For ANPR the problem now becomes that of recovering as much information about vehicle's journey as possible from the limited number of observations.
+For ANPR the problem now becomes that of recovering as much information about a vehicle's journey as possible from the limited number of observations.
 ANPR cameras are normally located on major roads and interchanges, however, this only covers a tiny fraction of the road network.
-We can, though, estimate routes between cameras by understanding the distances between cameras and the most ``sensible'' routes between them.
+We can, however, estimate routes between cameras by understanding the distances between cameras and the most ``sensible'' routes between them.
 This allows us, given a set of ANPR sightings of the same vehicle, to produce a ``most likely'' route for that journey.
@@
 Once we have a set of sightings of a vehicle using ANPR, we now need to convert these into actual journeys.
-The first requirement is to identify individual journeys.
+The first requirement is to separate the stream of sightings of a vehicle into individual journeys.
 Although this can't be done with certainty we can apply general rules to distinguish one journey from the next.
@@
 However, if the timings between two sightings is significantly longer than what would be expected then this would imply that the vehicle stopped between these two cameras and that the later sighting is part of a new journey.
+The process of journey identification needs to be performed on dirty data which contains numerous impurities which need to be handled.
+These include:
 
-The process of journey identification needs to be performed on dirty data which contains numerous impurities which need to be handled. These include:
-
+\setlist{nosep}
 \begin{itemize}
-	\item {\bf Number plate miss-reads:} Although ANPR cameras have accuracies of around 99, miss-reads are possible.
+	\item {\bf Number plate miss-reads:} Although ANPR cameras have accuracies of around 99\%, miss-reads are possible.
 This can lead to sightings being missed or vehicles being wrongly sighted in locations.
+	\vspace{0.1cm}
 	\item {\bf Timing errors:} The time-stamps of sightings could be erroneous.
 The minor side of this is implausible journey times, though, more seriously, this can lead to reordering the set of cameras on a particular journey.
-	\item {\bf Clones number plates:} For various reasons a number plate may be cloned and used on a different vehicle.
+	\vspace{0.1cm}
+	\item {\bf Cloned number plates:} For various reasons a number plate may be cloned and used on a different vehicle.
 This can lead to impossible journeys and journeys that the real vehicle did not make.
@@
 The rest of this paper is presented as follows.
-In Section \ref{s.related} we discuss related work.
-Section \ref{s.ncl} we presents the ANPR data for the Newcastle area.
-Our process for identifying individual journeys is presented in Section \ref{s.trips} while Section \ref{s.classification} presents our classification approach.
-We present results in Section \ref{s.results} before offering conclusions and future directions in Section \ref{s.conclusions}.
+In Section~\ref{s.related} we discuss related work.
+Section~\ref{s.ncl} presents the ANPR data for the Newcastle area.
+Our process for identifying individual journeys is defined in Section~\ref{s.trips}, while methods for addressing issues resulting from poor camera performance are described in Sections~\ref{s.duplicates} and~\ref{s.errors}.
+Section~\ref{s.classification} outlines our clustering approach.
+We report results in Section~\ref{s.results} before offering conclusions and future directions in Section~\ref{s.conclusions}.
 
@@
 
-\begin{figure}[t]
-\centering
-\includegraphics[width=.90\linewidth]{ANPR-overview.pdf}
-\caption{Overview of an ANPR system.}
-\label{fig:anpr-overview}
-\vspace{-0.6cm}
-\end{figure}
-
-The main use of ANPR data for the UTMC is estimating average journey times for selected or sensitive links in the road network.
-Furthermore, several authors have extensively researched how to use number plate data as an extension to link counts for estimating origin-destination matrices and link flows~\cite{Castillo2010, Castillo2008, Hazelton2012}.
+The main use of ANPR data for an Urban Traffic Management and Control Centre (UTMC) is estimating average journey times for selected or sensitive links in the road network.
+Furthermore, several authors, notably Enrique Castillo et al.\ and Andrew Hazelton et al.\, have extensively researched how to use number plate data as an extension to link counts for estimating origin-destination matrices and link flows~\cite{Castillo2010, Castillo2008, Hazelton2012}.
 However, very few works have focused on analysing individual or collective travel patterns from number plate data, particularly across extended periods of time.
@@
 
-Determining the distribution of travel modes is one of the fundamental steps in the four-stage model: an essential traffic modelling methodology for transportation planning~\cite{FourStepModel}.
+Determining the distribution of travel modes is third of the four fundamental steps in the four-stage model: trip generation, trip distribution, modal split and traffic assignment.
+The four-stage model is the most widely used traffic modelling methodology for transportation planning~\cite{FourStepModel}.
 Previous works have used trip information derived from different sources of data to identify travel mode or purpose of trip.
 More notably, survey data, floating car data and mobile phone data have been used~\cite{ODMobileData, ClusteringGPS}.
-Although number plate data has been used in~\cite{Clustering} to identify different categories of trips, the authors do not differentiate between private or public travel modes and focus instead on categorising trips by time of occurrence.
+Although number plate data has been used by Chen et al.\ (2017)~\cite{Clustering} to identify different categories of trips, the authors do not differentiate between private or public travel modes and focus instead on categorising trips by time of occurrence.
 Hence, in section~\ref{s.classification} we apply the \emph{k-means} clustering algorithm to derived trip data and based on the results, we discuss the limitations of proposed methods and that of ANPR data.
